### 2015-10-17 report/paper.tex
Add report

\documentclass{article} % For LaTeX2e
% We will use NIPS submission format
\usepackage{nips13submit_e,times}
% for hyperlinks
\usepackage{hyperref}
\usepackage{url}
% For figures
\usepackage{graphicx} 
\usepackage{subfigure} 
% math packages
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsopn}
\usepackage{ifthen}
\usepackage{natbib}

\title{Project-I by Group Sydney}

\author{
Diego Antognin and Jason Racine \\
EPFL \\
\texttt{diego.antognini@epfl.ch}, \texttt{jason.racine@epfl.ch} \\
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\nipsfinalcopy

\begin{document}

\maketitle

\begin{abstract}

\end{abstract}

\section{Regression}

\section{Data Description}

The train-data for regression consists of $N = 2800$ input ($\mathbf{X}$) and output ($\mathbf{y}$) data samples. Each input sample is a vector $\mathbf{x}_n$ with dimension $D = 76$. Out of these $76$ variables, $63$ are real, $3$ binary, $4$ categorical with 3 categories, $6$ are categorical with 4 categories.

We also have test-data of size $N=1200$ without their corresponding output. Our goal is to produce predictions for those data, as well as an approximation of the test-error.

\section{Data visualization and cleaning}

NORMALIZED ALL DATA OR NORMALIZATION PER CLUSTER ?

We first have plot the distribution of our features (plot not shown because too big for the $76$ features). As expected, they are not center and we should normalized them. The Figure \ref{fig:histogram} shows an histogram of the output ($\mathbf{y}$) and we can conclude our data  seem to be a combination of three Gaussian distributions. It will be used later in order to separate the data in three sets and apply different regression models on them. We can also observe that each cluster have different sizes : 1946, 576 and 278. Moreover, on the right, we can see some data points (there are $2$) which have a higher values than the others. We consider them as outliers and we will remove.

To separate the data, we have observed that the feature $2$ and $16$ could help us. Figure \ref{fig:feature2}. We can observe $11$ misclassified data (green points), which we will remove them in order to not corrupt our model.  So, this feature can allow us to find the first cluster. For the two others clusters, we need to oberve the feature $16$. Figure \ref{fig:feature16}. We can observe $14$ misclassified data (green points and one blue). They also will be consider as outliers and remove. We set the threshold in order to minimize the number of misclassified data samples. Those thresholds are $0.42$ for the feature $2$ and $1.17$ for the feature $16$.

We are also interested about the correlation between the input and output variables. We have observed the correlation for each cluster and conclude that for the first cluster, they are mainly in $[-0.1,0.1]$, except two features which are highly correlated. For the second and third cluster, also mainly in $[-0.1,0.1]$ but this time, there are more correlated features (~15). Moreover, the features seem not have correlation between them.

ENCODING CATEGORICAL FEATURES

We can note that the rank of our input matrix $\mathbf{X}$ is rank-deficient with a rank of 66 instead of 76.

\begin{figure}[!t]
\center
\subfigure[Histogram of $\mathbf{y}$. We can see three Gaussian distributions and also some outliers on the right.]{\includegraphics[width=2.5in]{figures/histogram.jpg} \label{fig:histogram}}
\hfill
\subfigure[Feature 2, because we have assumed that the data was generated from three Gaussian, this feature can help us to separate the data. Green data points are misclassified data. The separation is at $x=0.42$.]{\includegraphics[width=2.5in]{figures/feature2.jpg}\label{fig:feature2}}
\hfill
\subfigure[Feature 16, because we have assumed that the data was generated from three Gaussian, this feature can help us to separate the data. Green data points are misclassified data. We can see see a blue data point which is misclassified, it will be an outlier. The separation is at $x=1.17$.]{\includegraphics[width=2.5in]{figures/feature16.jpg}\label{fig:feature16}}
\hfill
\caption{}
\end{figure}

\end{document}

### 2015-10-17 report/paper.tex
Add final preprocessing steps

@@
 
-NORMALIZED ALL DATA OR NORMALIZATION PER CLUSTER ?
-
 We first have plot the distribution of our features (plot not shown because too big for the $76$ features).
-As expected, they are not center and we should normalized them.
+As expected, they are not center and we should normalized them (each cluster will be normalized independently).
 The Figure \ref{fig:histogram} shows an histogram of the output ($\mathbf{y}$) and we can conclude our data  seem to be a combination of three Gaussian distributions.
@@
 
-ENCODING CATEGORICAL FEATURES
+We use dummy encoding for the categorical variables (for a categorical variable of size $k$, we need $k-1$ features), which gives us a total of $93$ input variable.
 
-We can note that the rank of our input matrix $\mathbf{X}$ is rank-deficient with a rank of 66 instead of 76.
+We can note that the rank of our input matrix $\mathbf{X}$ is rank-deficient with a rank of 66 instead of 76, and is rank-deficient of 65 for each cluster.
 

### 2015-10-17 main.m

@@ -8,12 +8,5 @@ load('Sydney_regression.mat');
 y = y_train;
 X = X_train;
 
-yN = normalizedData(y);
-XN = normalizedData(X);
 [y_cls1, X_cls1, y_cls2, X_cls2, y_cls3, X_cls3] = preprocess(y, X);
 
-res = [0 0 0];
-for i = 1:1:length(XN)
-    idx_cls = findCorrespondingCluster(XN(i,:));
-    res(idx_cls) = res(idx_cls)+1;
-end

### 2015-10-17 preprocess.m

@@ -4,14 +4,76 @@
 function [y_cls1, X_cls1, y_cls2, X_cls2, y_cls3, X_cls3, idx_cls1, idx_cls2, idx_cls3] = preprocess(y, X)
     [idx_cls1, idx_cls2, idx_cls3] = findClusters(y, X);
     
+    % Find the discrete features
+    [idx_feature_2, idx_feature_3, idx_feature_4] = findDiscreteFeatures(X);
+    idx_feature_real = setdiff(setdiff(setdiff(1:1:size(X,2), idx_feature_2), idx_feature_3), idx_feature_4);
+
+    % We shouldn't normalize y
+    % Normalize each cluster independantely withou categorical features
     y_cls1 = y(idx_cls1);
-    X_cls1 = X(idx_cls1,:);
+    X_cls1 = normalizedData(X(idx_cls1,idx_feature_real));
     
     y_cls2 = y(idx_cls2);
-    X_cls2 = X(idx_cls2,:);
+    X_cls2 = normalizedData(X(idx_cls2,idx_feature_real));
     
     y_cls3 = y(idx_cls3);
-    X_cls3 = X(idx_cls3,:);
+    X_cls3 = normalizedData(X(idx_cls3,idx_feature_real));
+    
+    % We encode categorical features and add the to the normalized X
+    X_cls1 = dummyFeatureEncoding(X_cls1, X(idx_cls1,:), idx_feature_2, idx_feature_3, idx_feature_4);
+    X_cls2 = dummyFeatureEncoding(X_cls2, X(idx_cls2,:), idx_feature_2, idx_feature_3, idx_feature_4);
+    X_cls3 = dummyFeatureEncoding(X_cls3, X(idx_cls3,:), idx_feature_2, idx_feature_3, idx_feature_4);
+end
+
+function [XX] = dummyFeatureEncoding(X,X_original, idx_feature_2, idx_feature_3, idx_feature_4)
+    XX = zeros(length(X_original),1);
+    % The index to copy the real values (non discrete) of X
+    i_real_X = 1;
+    for i = 1:1:size(X_original,2)
+        % If categorical of 2, just copy the cols
+        if any(idx_feature_2 == i)
+          XX(:,end+1) = X_original(:,i);
+        % If categorical of 3 or 4, create (k-1) cols and put 0-0-0 if 0,
+        % 1-0-0 if 1, 0-1-0 if 2, 0-0-1 if 4 etc.
[…]

### 2015-10-23 report/paper.tex
Advance report

@@
 
-\section{Data Description}
+\subsection{Data Description}
 
@@
 
-\section{Data visualization and cleaning}
+\subsection{Data visualization and cleaning}
 
@@
 
+\subsection{Ridge Regression}
+
+Since our matrix is ill-conditioned, using the least squares with normal equation wasn't suitable.
+Moreover, the results obtained were less good (not so much) compared to the ridge regression method, and so we will not report the results obtained with this method.
+The method using the least squares with gradient descent was appropriated, but compared to the ridge regression, it was a little bit less efficient and was also slower to tune correctly the learning step.
+This is the reason why we will not report the results using this method.
+Therefore, we have so used the ridge regression method, which is suitable because we should lift the eigenvalues, faster and a little bit more efficient than least squares using gradient descent.
+Only results obtained with this latter will be reported.
+
+\subsubsection{Evaluation methods}
+
+In order to evaluate our different models, we first have split our data ($\mathbf{X}$) in two sets of $80\%$ and $20\%$, for the training and the test sets.
+We learn with our models on the training set and use the test set in order to estimate the error using the \textit{Root Mean Square Error}.
+Moreover, we have also used k-cross validation (see next section for the values of k).
+We repeated the experiment $100$ times with different seed in order to split the data in a unique random way for each trial.
+Finally, we made varied the lambda from $10^{-5}$ to $10^{5}$ with $500$ values between.
+
+\subsubsection{Model comparison}
+
+The figure XXX represents the \textit{RMSE} for each models.
+The first model used in the constant one, which will allow us to compare improvements with others model.
+It simply returns the value of the mean of the outputs and doesn't depend on the inputs.
+Our second model consists of using also a constant model as in first, but this time separating the data in the three clusters identified during the exploratory data analysis.
+We can observed a big improvement, justifying our choice to split the data in clusters.
+
+The third model does a ridge regression on the overall data, in order to check that the separation of the data into clusters is better, even using only constant model per cluster.
+Hopefully, this is the case.
+The next model add the normalization on the input variables, another way to check if the separation of the data into cluster was a good things.
+Normalization is generally a good thing and often a necessity depending on the algorithms used (e.g.
+gradient descent ones).
+It improves only slightly but because normalization is a good practice, we keep it.
+
+The fifth model consists of the separation of the data, using the normalization per cluster and finally does a ridge regression per cluster.
+With this manner, we can tune each cluster independently.
+We can see an important improvement because of the separation of the data.
+The next model adds the dummy encoding for the categorical variables.
+We can observe that it seems to be equivalent.
+
+The seventh model is about using polynomial basis functions for each cluster.
+We have to be careful about the polynomial to avoid the case where $D > N$ (especially for the cluster two and three, because there are much smaller than the first one).
+This model doesn't use dummy encoding and so can add higher polynomial.
+The degrees of the polynomial for the clusters are $3,6$ and $3$.
+When we add higher degrees, we can directly observe overfitting because suddenly the training error decreases to a value near zero and the test errors grows exponentially, and so we know that our degrees are not to high.
+The eight model is similar to the seventh one, except that this time it uses dummy encoding and have lower degrees : $3,5$ and $2$, due to the number of input variables ($92$ vs $76$).
+We can see that these feature transformations significantly improve the prediction accuracy compared to others model.
+
+\subsubsection{Feature transformations}
+
+In hope to minimize the error, we tried different feature transformations.
+The first try was add polynomial of them self as $X = [X X^2 X^3 ...]$ for which each cluster has is own degree.
+We made varied the different degrees and also the lambdas in order to keep the lambda which minimize the test error.
+
+We also thought about removing some uncorrelated features, we tried to remove the most uncorrelated then remove also the second one etc, but unfortunately we didn't get improvements.
+The reason is that maybe a feature may not be correlated directly with the input, however a combination of uncorrelated feature might be highly correlated.
+
 \end{document}

### 2015-10-23 testModels.m

@@ -22,7 +22,7 @@ N = length(y);
 numberOfExperiments = 100;
 proportionOfTraining = 0.8;
 
-%
+%%
 % **********************************
 %             TEST 0
 % **********************************
@@ -105,7 +105,7 @@ K = 10;
 for i = 1:1:numberOfExperiments
     setSeed(28111993*i);
     [XTr, yTr, XTe, yTe] = splitProp(proportionOfTraining, y, normalizedData(X));
-    
+
     % Learning
     lambda = findLambda(K, yTr, XTr, 0, 1);
     
@@ -131,7 +131,7 @@ k3 = 4;
 for i = 1:1:numberOfExperiments
     setSeed(28111993*i);
     [XTr, yTr, XTe, yTe] = splitProp(proportionOfTraining, y, X);
-    
+    fprintf('%d\n', i);    
     [y_cls1, X_cls1, y_cls2, X_cls2, y_cls3, X_cls3, idx_cls1, idx_cls2, idx_cls3] = preprocess(yTr, XTr);
 
     % Learning
@@ -273,7 +273,7 @@ d3 = 2;
 for i = 1:1:numberOfExperiments
     setSeed(28111993*i);
     [XTr, yTr, XTe, yTe] = splitProp(proportionOfTraining, y, X);
-    fprintf('%d \n', i);
+    
     % Learning
      lambda_cls1 = findLambda(k1, yTr, XTr, 1, d1, d2, d3);
      lambda_cls2 = findLambda(k2, yTr, XTr, 2, d1, d2, d3);

### 2015-10-24 report/paper.tex
Finish testModels

@@
 Those thresholds are $0.42$ for the feature $2$ and $1.17$ for the feature $16$.
+So the final size of our clusters are $1937, 563$ and $273$.
 
@@
 Moreover, the results obtained were less good (not so much) compared to the ridge regression method, and so we will not report the results obtained with this method.
-The method using the least squares with gradient descent was appropriated, but compared to the ridge regression, it was a little bit less efficient and was also slower to tune correctly the learning step.
+The method using the least squares with gradient descent was appropriated, but compared to the ridge regression, it was a little bit less efficient and was also much slower to tune correctly the learning step.
 This is the reason why we will not report the results using this method.
@@
 In order to evaluate our different models, we first have split our data ($\mathbf{X}$) in two sets of $80\%$ and $20\%$, for the training and the test sets.
-We learn with our models on the training set and use the test set in order to estimate the error using the \textit{Root Mean Square Error}.
-Moreover, we have also used k-cross validation (see next section for the values of k).
+We choose this percentage in order to have still enough data for the training set and also enough in order to estimate correctly our models on the testing set with a concrete size.
+We learn with our models only on the training set and use the test set in order to estimate the error using the \textit{Root Mean Square Error}.
+Moreover, we have also used k-cross validation (see next section for the values of k) to tune the parameters like lambda of the different polynomial degrees.
 We repeated the experiment $100$ times with different seed in order to split the data in a unique random way for each trial.
@@
 The first model used in the constant one, which will allow us to compare improvements with others model.
-It simply returns the value of the mean of the outputs and doesn't depend on the inputs.
-Our second model consists of using also a constant model as in first, but this time separating the data in the three clusters identified during the exploratory data analysis.
-We can observed a big improvement, justifying our choice to split the data in clusters.
-
-The third model does a ridge regression on the overall data, in order to check that the separation of the data into clusters is better, even using only constant model per cluster.
-Hopefully, this is the case.
-The next model add the normalization on the input variables, another way to check if the separation of the data into cluster was a good things.
+It simply returns the mean of the outputs and doesn't depend on the inputs.
+Our second model does a ridge regression on the overall data using $10$ fold cross validation, because we have a lot of data, splitting them into $10$ fold seems reasonable.
+As expected, this is much better than previous model because we take into account the input features.
+Next model is similar to the second one except that we add normalization on the real inputs (it doesn't make sense to also normalize categorical variables).
 Normalization is generally a good thing and often a necessity depending on the algorithms used (e.g.
@@
 
-The fifth model consists of the separation of the data, using the normalization per cluster and finally does a ridge regression per cluster.
-With this manner, we can tune each cluster independently.
-We can see an important improvement because of the separation of the data.
-The next model adds the dummy encoding for the categorical variables.
-We can observe that it seems to be equivalent.
+At this stage, we want to try to separate our data into cluster.
+The fourth model consists of using a constant model as in first but this time separating the data into the three clusters identified during the exploratory data analysis.
+We can observe a big improvement, justifying our choice to split the data in clusters.
+The next model is about doing a ridge regression per cluster, each one having its own lambda and its own normalization.
+We also use k-cross validation where the first cluster has a $k_1=10$ ($1937$) because we have a lot of data, the second $k_2=6$ ($563$ samples) and the last $k_3=4$ ($273$) because it doesn't have so much data and so we have to choose a smaller k.
+As expected we obtained another big improvement justifying again the choice of splitting the data.
 
-The seventh model is about using polynomial basis functions for each cluster.
+The sixth model consists of using dummy encoding for the categorical variables.
+As explained in the exploratory data analysis, we use $k-1$ features to represent a $k$ categorical variable.
+This is a common practice.
+Unfortunately, we didn't get major improvements.
+
+The two next models is about using polynomial basis functions for each cluster.
+Categorical variables aren't concerned, they are just replicated for each degree, because it doesn't make any sense to build polynomial for them.
+We normalize only when we did our feature transformations, in order to avoid loosing information.
 We have to be careful about the polynomial to avoid the case where $D > N$ (especially for the cluster two and three, because there are much smaller than the first one).
-This model doesn't use dummy encoding and so can add higher polynomial.
-The degrees of the polynomial for the clusters are $3,6$ and $3$.
-When we add higher degrees, we can directly observe overfitting because suddenly the training error decreases to a value near zero and the test errors grows exponentially, and so we know that our degrees are not to high.
-The eight model is similar to the seventh one, except that this time it uses dummy encoding and have lower degrees : $3,5$ and $2$, due to the number of input variables ($92$ vs $76$).
+The seventh model doesn't use dummy encoding and the eight yes.
+
+For the seventh model, the degrees of the polynomial for the clusters are $3,6$ and $3$.
+When we add higher degrees, we can directly observe overfitting because suddenly the training error decreases to a value near zero and the test errors grows exponentially, so with this manner we know that our degrees are not to high.
+The eighth model is similar to the seventh one, except that this time it uses dummy encoding and has lower degrees : $3,5$ and $2$, due to the number of input variables ($92$ vs $76$).
 We can see that these feature transformations significantly improve the prediction accuracy compared to others model.
+However, dummy encoding seems to bit a bit more efficient.
 
@@
 We made varied the different degrees and also the lambdas in order to keep the lambda which minimize the test error.
+We also try feature transformations where the degree is a root, but unfortunately it was worst.
+We also wanted to try to just pow features with different degrees (without add polynomial of themselves) but we couldn't do it due to a lack of time.
 

### 2015-10-24 preprocess.m

@@ -64,46 +64,57 @@ function [y, X] = preprocessCluster(y_, X_, idx, degree, idx_feature_2, idx_feat
     
     % Normalize
     y = y_(idx);
+    % WITH NORMALIZATION
     X = normalizedData(X(:,real_features_idx_to_normalize));
-
+    % WITHOUT NORMALIZATION
+    % X = X(:,real_features_idx_to_normalize);
+    
+    % WITH DUMMY ENCODING
     % We encode categorical features and add them to the normalized X
     X = dummyFeatureEncoding(X, X__, idx_feature_2_poly, idx_feature_3_poly, idx_feature_4_poly); 
-    
+    % WITHOUT DUMMY ENCODING
+    % X = addCategoricalVariableWithoutEncoding(X, X__, idx_feature_2_poly, idx_feature_3_poly, idx_feature_4_poly);
+
 end
 
 function [XX] = dummyFeatureEncoding(X,X_original, idx_feature_2, idx_feature_3, idx_feature_4)
     % WITH DUMMY ENCODING
     
-%     XX = zeros(size(X_original, 1),1);
-%     % The index to copy the real values (non discrete) of X
-%     i_real_X = 1;
-%     for i = 1:1:size(X_original,2)
-%         % If categorical of 2, just copy the cols
-%         if any(idx_feature_2 == i)
-%           XX(:,end+1) = X_original(:,i);
-%         % If categorical of 3 or 4, create (k-1) cols and put 0-0-0 if 0,
-%         % 1-0-0 if 1, 0-1-0 if 2, 0-0-1 if 4 etc.
-%         elseif any(idx_feature_3 == i)
-%             for j = 1:1:(3-1)
-%                 cols = zeros(size(X, 1),1);
-%                 cols(find(X_original(:,i) == j),1) = 1;
-%                 XX(:,end+1) = cols;
-%             end
-%         elseif any(idx_feature_4 == i)
-%             for j = 1:1:(4-1)
-%                 cols = zeros(size(X,1),1);
-%                 cols(find(X_original(:,i) == j),1) = 1;
-%                 XX(:,end+1) = cols;
-%             end
-%         % Copy the real values (non discrete)
-%         else
-%             XX(:,end+1) = X(:,i_real_X);
-%             i_real_X = i_real_X + 1;
-%         end
-%     end
+    XX = zeros(size(X_original, 1),1);
+    % The index to copy the real values (non discrete) of X
+    i_real_X = 1;
[…]

### 2015-10-24 report/paper.tex
Fix bug RMSE

@@
 Moreover, we have also used k-cross validation (see next section for the values of k) to tune the parameters like lambda of the different polynomial degrees.
-We repeated the experiment $100$ times with different seed in order to split the data in a unique random way for each trial.
+We repeated the experiment $30$ times with different seed in order to split the data in a unique random way for each trial.
 Finally, we made varied the lambda from $10^{-5}$ to $10^{5}$ with $500$ values between.
@@
 
+\begin{figure}[!t]
+\center
+\includegraphics[width=6in]{figures/models.jpg} \label{fig:models}
+\caption{Boxplot of our models, using the RMSE. Each model is represented with a box where we can see their mean and their standard deviation.}
+\end{figure}
+
 \end{document}

### 2015-10-24 learningCurve.m

@@ -10,28 +10,30 @@ y = y_train;
 X = X_train;
 N = length(y);
 
-[y_cls1, X_cls1, y_cls2, X_cls2, y_cls3, X_cls3] = preprocess(y, X);
+numberOfExperiments = 2;
+proportionOfTraining = 0.8;
+
 d1 = 3;
 d2 = 5;
 d3 = 2;
 
-l1 = 0.035;
-l2 = 0.1298;
-l3 = 0.7480;
-
-k1 = 5;
-k2 = 5;
+k1 = 10;
+k2 = 6;
 k3 = 4;
 
-K = 0:5:95;
+K = 10:10:90;
 K = K(2:end);
 K = K./100;
-for i = 1:1:length(K)
-    k = K(i);
-    
-    [err_tr_cls1(:,i), err_te_cls1(:,i)] = modelLearningCurve(k, k1, d1, l1, y_cls1, X_cls1);
-    [err_tr_cls2(:,i), err_te_cls2(:,i)] = modelLearningCurve(k, k2, d2, l2, y_cls2, X_cls2);
-    [err_tr_cls3(:,i), err_te_cls3(:,i)] = modelLearningCurve(k, k3, d3, l3, y_cls3, X_cls3);
+
+for j = 1:1:numberOfExperiments
+    setSeed(28111993*j);
+    [XTr, yTr, XTe, yTe] = splitProp(proportionOfTraining, y, X);
+    fprintf('%d\n', j); 
+
+    for i = 1:1:length(K)
+        k = K(i);
+        [err_tr_cls1(j,i), err_te_cls1(j,i), err_tr_cls2(j,i), err_te_cls2(j,i), err_tr_cls3(j,i), err_te_cls3(j,i)] = modelLearningCurve(k, k1, k2, k3, yTr, XTr, d1, d2, d3, yTe, XTe);
+    end
 end
 
 figure

### 2015-10-24 modelLearningCurve.m

@@ -1,55 +1,54 @@
 % Written by Diego Antognini & Jason Racine, EPFL 2015
 % all rights reserved
 
-function [err_tr, err_te_cv] = modelLearningCurve(prop, K, d, lambda, y, X)
-    for i = 1:1:10
-        setSeed(28111993*i);
-        [XTr, yTr, XTe, yTe] = splitProp(prop, y, X);
+function [err_tr_cls1, err_te_cls1, err_tr_cls2, err_te_cls2, err_tr_cls3, err_te_cls3] = modelLearningCurve(prop, k1, k2, k3, y, X, d1, d2, d3, globalYTe, globalXTe)
+    [XTr, yTr, XTe, yTe] = splitProp(prop, y, X);
+    
+     lambda_cls1 = findLambda(k1, yTr, XTr, 1, d1, d2, d3);
+     lambda_cls2 = findLambda(k2, yTr, XTr, 2, d1, d2, d3);
+     lambda_cls3 = findLambda(k3, yTr, XTr, 3, d1, d2, d3);
+    
+     [y_cls1, X_cls1, y_cls2, X_cls2, y_cls3, X_cls3, idx_cls1, idx_cls2, idx_cls3] = preprocess(yTr, XTr, d1, d2, d3);
 
-        % CV Error
-        N = length(yTr);
-        idxCV = splitGetCV(K, N);
-        for k=1:1:K
-            [XXTr, yyTr, XXTe, yyTe] = splitGetTrTe(yTr, XTr, idxCV, k);
-
-            tXTr = [ones(length(yyTr),1) buildPoly(XXTr, d)];
-            tXTe = [ones(length(yyTe),1) buildPoly(XXTe, d)];
-
-            beta_rr = ridgeRegression(yyTr, tXTr, lambda);
-            err_te_rr_cv(k) = RMSE(yyTe, tXTe*beta_rr);
-        end
-        err_te_cv(i) = mean(err_te_rr_cv);
-        
-        % Training Error
-        tXTr = [ones(length(yTr),1) buildPoly(XTr, d)];
-        beta_rr = ridgeRegression(yTr, tXTr, lambda);
-        err_tr(i) = RMSE(yTr, tXTr*beta_rr);
-    end
+     tX_cls1 = [ones(length(y_cls1),1) X_cls1];
+     tX_cls2 = [ones(length(y_cls2),1) X_cls2];
[…]

### 2015-10-24 testModels.m

@@ -317,7 +317,7 @@ saveFile(err8_cls3, 'err8_cls3');
 %%
 s = [1 numberOfExperiments];
 
-[y_cls1, X_cls1, y_cls2, X_cls2, y_cls3, X_cls3, idx_cls1, idx_cls2, idx_cls3] = preprocess(y, X, d1, d2, d3);
+[y_cls1, X_cls1, y_cls2, X_cls2, y_cls3, X_cls3, idx_cls1, idx_cls2, idx_cls3] = preprocess(y, X);
 
 N = length(y_cls1) + length(y_cls2) + length(y_cls3);
 ratio_cls1 = size(X_cls1, 1)/N;
@@ -327,12 +327,39 @@ ratio_cls3 = size(X_cls3, 1)/N;
 err1 = openFile('err1', s);
 err2 = openFile('err2',s);
 err3 = openFile('err3', s);
-err4 = ratio_cls1*openFile('err4_cls1', s) + ratio_cls2*openFile('err4_cls2', s) + ratio_cls3*openFile('err4_cls3', s);
-err5 = ratio_cls1*openFile('err5_cls1', s) + ratio_cls2*openFile('err5_cls2',s) + ratio_cls3*openFile('err5_cls3',s);
-err6 = ratio_cls1*openFile('err6_cls1', s) + ratio_cls2*openFile('err6_cls2',s) + ratio_cls3*openFile('err6_cls3',s);
-err7 = ratio_cls1*openFile('err7_cls1', s) + ratio_cls2*openFile('err7_cls2',s) + ratio_cls3*openFile('err7_cls3',s);
-err8 = ratio_cls1*openFile('err8_cls1', s) + ratio_cls2*openFile('err8_cls2',s) + ratio_cls3*openFile('err8_cls3',s);
 
+err4 = sqrt((1/N)*( ...
+            (size(X_cls1, 1)*(openFile('err4_cls1', s).^2)) ...
+          + (size(X_cls2, 1)*(openFile('err4_cls2', s).^2)) ...
+          + (size(X_cls3, 1)*(openFile('err4_cls3', s).^2)) ...
+     ));
+     
+ 
+err5 = sqrt((1/N)*( ...
+            (size(X_cls1, 1)*(openFile('err5_cls1', s).^2)) ...
+          + (size(X_cls2, 1)*(openFile('err5_cls2', s).^2)) ...
+          + (size(X_cls3, 1)*(openFile('err5_cls3', s).^2)) ...
+     ));
+ 
[…]

### 2015-10-24 report/paper.tex
pretty print for figures

@@
 To separate the data, we have observed that the feature $2$ and $16$ could help us.
-Figure \ref{fig:feature2}.
-We can observe $11$ misclassified data (green points), which we will remove them in order to not corrupt our model.
+With Figure \ref{fig:feature2} we can observe $11$ misclassified data, which we will remove them in order to not corrupt our model.
 So, this feature can allow us to find the first cluster.
 For the two others clusters, we need to oberve the feature $16$.
-Figure \ref{fig:feature16}.
-We can observe $14$ misclassified data (green points and one blue).
+With Figure \ref{fig:feature16}, we can observe $14$ misclassified data.
 They also will be consider as outliers and remove.
@@
 We have observed the correlation for each cluster and conclude that for the first cluster, they are mainly in $[-0.1,0.1]$, except two features which are highly correlated.
-For the second and third cluster, also mainly in $[-0.1,0.1]$ but this time, there are more correlated features (~15).
+For the second and third cluster, also mainly in $[-0.1,0.1]$ but this time, there are more correlated features (around 15).
 Moreover, the features seem not have correlation between them.
@@
 \center
-\subfigure[Histogram of $\mathbf{y}$. We can see three Gaussian distributions and also some outliers on the right.]{\includegraphics[width=2.5in]{figures/histogram.jpg} \label{fig:histogram}}
+\subfigure[Histogram of $\mathbf{y}$. We can see three Gaussian distributions and also some outliers on the right.]{\includegraphics[width=2.7in]{figures/histogram.jpg} \label{fig:histogram}}
 \hfill
-\subfigure[Feature 2, because we have assumed that the data was generated from three Gaussian, this feature can help us to separate the data.
+\subfigure[Feature 2 can help us to separate the data.
 Green data points are misclassified data.
-The separation is at $x=0.42$.]{\includegraphics[width=2.5in]{figures/feature2.jpg}\label{fig:feature2}}
+The separation is at $x=0.42$.]{\includegraphics[width=2.7in]{figures/feature2.jpg}\label{fig:feature2}}
 \hfill
-\subfigure[Feature 16, because we have assumed that the data was generated from three Gaussian, this feature can help us to separate the data.
+\subfigure[Feature 16 can help us to separate the data.
 Green data points are misclassified data.
-We can see see a blue data point which is misclassified, it will be an outlier.
-The separation is at $x=1.17$.]{\includegraphics[width=2.5in]{figures/feature16.jpg}\label{fig:feature16}}
+The separation is at $x=1.17$.]{\includegraphics[width=2.7in]{figures/feature16.jpg}\label{fig:feature16}}
 \hfill
+\includegraphics[width=6in]{figures/models.jpg} \label{fig:models}
+\caption{Boxplot of our models, using the RMSE. Each model is represented with a box where we can see their mean and their standard deviation.}
 \caption{}
@@
 Since our matrix is ill-conditioned, using the least squares with normal equation wasn't suitable.
-Moreover, the results obtained were less good (not so much) compared to the ridge regression method, and so we will not report the results obtained with this method.
+Moreover, the results obtained were less good (not so much) compared to the ridge regression method, and so we will not report those.
 The method using the least squares with gradient descent was appropriated, but compared to the ridge regression, it was a little bit less efficient and was also much slower to tune correctly the learning step.
 This is the reason why we will not report the results using this method.
-Therefore, we have so used the ridge regression method, which is suitable because we should lift the eigenvalues, faster and a little bit more efficient than least squares using gradient descent.
+Therefore, we have so used the ridge regression method, which is suitable because we could lift the eigenvalues in order to avoid ill-conditioned matrix, faster and a little bit more efficient than least squares using gradient descent.
 Only results obtained with this latter will be reported.
@@
 We choose this percentage in order to have still enough data for the training set and also enough in order to estimate correctly our models on the testing set with a concrete size.
-We learn with our models only on the training set and use the test set in order to estimate the error using the \textit{Root Mean Square Error}.
-Moreover, we have also used k-cross validation (see next section for the values of k) to tune the parameters like lambda of the different polynomial degrees.
-We repeated the experiment $30$ times with different seed in order to split the data in a unique random way for each trial.
-Finally, we made varied the lambda from $10^{-5}$ to $10^{5}$ with $500$ values between.
+We learn only on the training set and then use the test set in order to estimate the error using the \textit{Root Mean Square Error}.
+Moreover, we have also used k-cross validation (see next section for the values of k) to tune the parameters up like lambda of the different polynomial degrees.
+We repeated the experiment $30$ times with different seed in order to split the data in a unique random way for each trial and so .
+Finally, we made vary the lambda from $10^{-5}$ to $10^{5}$ with $500$ values in-between.
 
@@
 
-The figure XXX represents the \textit{RMSE} for each models.
+The figure \ref{fig:models} represents the \textit{RMSE} for each models.
 The first model used in the constant one, which will allow us to compare improvements with others model.
@@
 
-\begin{figure}[!t]
-\center
-\includegraphics[width=6in]{figures/models.jpg} \label{fig:models}
-\caption{Boxplot of our models, using the RMSE. Each model is represented with a box where we can see their mean and their standard deviation.}
-\end{figure}
-
 \end{document}

### 2015-10-24 report/paper.tex
Advance report

@@
 \begin{abstract}
-
+This report provides a summary of the project one of the PCML class.
+The project consists of doing regression and the classification on data.
+For the regression data set, we have observed that it was essential to separate the data into three clusters and doing a ridge regression on each of them due to the ill-condition matrix and the rapidity of tuning the algorithm.
+Moreover feature transformations like using polynomial basis was a key to decrease the prediction accuracy.
 \end{abstract}
@@
 We are also interested about the correlation between the input and output variables.
-We have observed the correlation for each cluster and conclude that for the first cluster, they are mainly in $[-0.1,0.1]$, except two features which are highly correlated.
+We have observed the correlation for each cluster and concfigurelude that for the first cluster, they are mainly in $[-0.1,0.1]$, except two features which are highly correlated.
 For the second and third cluster, also mainly in $[-0.1,0.1]$ but this time, there are more correlated features (around 15).
@@
 
-\begin{figure}[!t]
+\begin{figure}[!ht]
 \center
@@
 The separation is at $x=1.17$.]{\includegraphics[width=2.7in]{figures/feature16.jpg}\label{fig:feature16}}
+\hfill
+\subfigure[Different degree for the polynomial basis of the cluster 1. We can observe a low variance.]{\includegraphics[width=2.7in]{figures/degrees_cls1.jpg}\label{fig:degrees_cls1}}
 \hfill
@@
 
-The figure \ref{fig:models} represents the \textit{RMSE} for each models.
-The first model used in the constant one, which will allow us to compare improvements with others model.
+The figure \ref{fig:models} represents the RMSE for each models.
+The first model used is the constant one, which will allow us to compare improvements with other models.
 It simply returns the mean of the outputs and doesn't depend on the inputs.
@@
 
-At this stage, we want to try to separate our data into cluster.
-The fourth model consists of using a constant model as in first but this time separating the data into the three clusters identified during the exploratory data analysis.
-We can observe a big improvement, justifying our choice to split the data in clusters.
+At this stage, we want to try to separate our data into clusters.
+We must be careful because it increases the ill-conditioning !
+The fourth model consists of using a constant model but this time separating the data into the three clusters identified during the exploratory data analysis and computing the mean for each cluster.
+We can observe a big improvement, justifying our choice to split the data into clusters.
 The next model is about doing a ridge regression per cluster, each one having its own lambda and its own normalization.
-We also use k-cross validation where the first cluster has a $k_1=10$ ($1937$) because we have a lot of data, the second $k_2=6$ ($563$ samples) and the last $k_3=4$ ($273$) because it doesn't have so much data and so we have to choose a smaller k.
+We also use k-cross validation where the first cluster has a $k_1=10$ because we have a lot of data ($1937$), the second $k_2=6$, because we don't have a lot of data ($563$) and the last $k_3=4$ because we really a small number of data ($273$) and so we have to choose a smaller k.
 As expected we obtained another big improvement justifying again the choice of splitting the data.
@@
 Unfortunately, we didn't get major improvements.
+We think that the reason why dummy encoding doesn't help is because we have a low bias and so, having more features doesn't help.
 
@@
 We have to be careful about the polynomial to avoid the case where $D > N$ (especially for the cluster two and three, because there are much smaller than the first one).
-The seventh model doesn't use dummy encoding and the eight yes.
+Moreover, if we don't observe significant improvement between degrees, we keep the lowest one.
+The Figure \ref{fig:degrees_cls1}, shows the case for the first cluster, we can also see that the variance is very low (same observation for the third one, however the second one has higher variance).
+The seventh model doesn't use dummy encoding and the eighth does.
 
@@
 The eighth model is similar to the seventh one, except that this time it uses dummy encoding and has lower degrees : $3,5$ and $2$, due to the number of input variables ($92$ vs $76$).
-We can see that these feature transformations significantly improve the prediction accuracy compared to others model.
-However, dummy encoding seems to bit a bit more efficient.
+We can see that these feature transformations significantly improve the prediction accuracy compared to other models.
+However, dummy encoding seems to more efficient, and so our final model will be the seventh.
 
@@
 The first try was add polynomial of them self as $X = [X X^2 X^3 ...]$ for which each cluster has is own degree.
-We made varied the different degrees and also the lambdas in order to keep the lambda which minimize the test error.
+We made vary the different degrees and also the lambdas in order to keep the lambda which minimize the test error.
 We also try feature transformations where the degree is a root, but unfortunately it was worst.
@@
 
-We also thought about removing some uncorrelated features, we tried to remove the most uncorrelated then remove also the second one etc, but unfortunately we didn't get improvements.
-The reason is that maybe a feature may not be correlated directly with the input, however a combination of uncorrelated feature might be highly correlated.
+We also thought about removing some uncorrelated features, we tried to remove the most uncorrelated ones and then remove also the second one etc, but unfortunately we didn't get improvements.
+We think that the reason is that a feature may not be correlated directly with the output however a combination of uncorrelated features might be highly correlated.
+
+\subsubsection{Summary}
+
+We analyzed a regression data set with different methods and concluded that least squares wasn't a good approach due to the ill-conditioning and the least squares gradient descent was to slow and not more efficient than ridge regression.
+Moreover, the separation of the data into clusters was essential in order to decrease the prediction accuracy.
+Unfortunately, dummy encoding didn't help however.
+Using polynomial basis help to decrease the variance.
+Finally, the variance of cluster 1 and 3 are very low, the one for cluster 2 is a bit higher.
+We think that we have bias because we use ridge regression and we don't think it is too high because we have seen that adding features helps and at a certain points it wasn't efficient.
+The best model we have is the seventh one, and we estimate the RMSE of the prediction to be around 97 with a standard deviation of 35.
+Finally, we think that our model is improvable however, we cannot try other things due to a lack of time.
 

### 2015-10-24 report/paper.tex
Fix typo reports

@@
 This report provides a summary of the project one of the PCML class.
-The project consists of doing regression and the classification on data.
-For the regression data set, we have observed that it was essential to separate the data into three clusters and doing a ridge regression on each of them due to the ill-condition matrix and the rapidity of tuning the algorithm.
+The project consists of doing regression and classification on data.
+For the regression data set we have observed that it was essential to separate the data into three clusters and doing a ridge regression on each of them due to the ill-condition matrix and the rapidity of tuning the algorithm.
 Moreover feature transformations like using polynomial basis was a key to decrease the prediction accuracy.
@@
 We first have plot the distribution of our features (plot not shown because too big for the $76$ features).
-As expected, they are not center and we should normalized them (each cluster will be normalized independently).
+As expected, they are not center and we should normalized them.
 The Figure \ref{fig:histogram} shows an histogram of the output ($\mathbf{y}$) and we can conclude our data  seem to be a combination of three Gaussian distributions.
-It will be used later in order to separate the data in three sets and apply different regression models on them.
+It will be used later in order to separate the data in three sets and apply different regression models on them (each cluster will be normalized independently).
 We can also observe that each cluster have different sizes : 1946, 576 and 278.
 Moreover, on the right, we can see some data points (there are $2$) which have a higher values than the others.
-We consider them as outliers and we will remove.
+We consider them as outliers and will remove them.
 
 To separate the data, we have observed that the feature $2$ and $16$ could help us.
-With Figure \ref{fig:feature2} we can observe $11$ misclassified data, which we will remove them in order to not corrupt our model.
-So, this feature can allow us to find the first cluster.
-For the two others clusters, we need to oberve the feature $16$.
+With Figure \ref{fig:feature2} we can see how we can use the second feature to separate cluster one to the others and observe $11$ misclassified data, which we will remove them in order to not corrupt our model.
+For the two other clusters, we need to observe the feature $16$.
 With Figure \ref{fig:feature16}, we can observe $14$ misclassified data.
-They also will be consider as outliers and remove.
+They also will be consider as outliers and removed.
 We set the threshold in order to minimize the number of misclassified data samples.
@@
 We are also interested about the correlation between the input and output variables.
-We have observed the correlation for each cluster and concfigurelude that for the first cluster, they are mainly in $[-0.1,0.1]$, except two features which are highly correlated.
-For the second and third cluster, also mainly in $[-0.1,0.1]$ but this time, there are more correlated features (around 15).
+We have observed the correlation for each cluster and conclude that for the first cluster, they are mainly in the range $[-0.1,0.1]$, except two features which are highly correlated.
+For the second and third cluster, also mainly in range $[-0.1,0.1]$ but this time, there are more correlated features (around 15).
 Moreover, the features seem not have correlation between them.
@@
 \hfill
-\subfigure[Different degree for the polynomial basis of the cluster 1. We can observe a low variance.]{\includegraphics[width=2.7in]{figures/degrees_cls1.jpg}\label{fig:degrees_cls1}}
+\subfigure[Different degree for the polynomial basis of the cluster 1.
+We can observe a low variance.
+The third degree seems the best model to choose.]{\includegraphics[width=2.7in]{figures/degrees_cls1.jpg}\label{fig:degrees_cls1}}
 \hfill
-\includegraphics[width=6in]{figures/models.jpg} \label{fig:models}
-\caption{Boxplot of our models, using the RMSE. Each model is represented with a box where we can see their mean and their standard deviation.}
+\subfigure[Boxplot of our models, using the RMSE.
+Each model is represented with a box where we can see their mean and their standard deviation.]{\includegraphics[width=5in]{figures/models.jpg}} \label{fig:models}
 \caption{}
@@
 Since our matrix is ill-conditioned, using the least squares with normal equation wasn't suitable.
-Moreover, the results obtained were less good (not so much) compared to the ridge regression method, and so we will not report those.
-The method using the least squares with gradient descent was appropriated, but compared to the ridge regression, it was a little bit less efficient and was also much slower to tune correctly the learning step.
+Moreover, the results obtained were less good compared to the ridge regression method, and so we will not report those.
+The method using the least squares with gradient descent was appropriated, but compared to the ridge regression, it was less efficient and was also much slower to tune correctly the learning step.
 This is the reason why we will not report the results using this method.
-Therefore, we have so used the ridge regression method, which is suitable because we could lift the eigenvalues in order to avoid ill-conditioned matrix, faster and a little bit more efficient than least squares using gradient descent.
+Therefore, we have so used the ridge regression method, which is suitable because we could lift the eigenvalues in order to avoid ill-conditioned matrix.
+It is also faster and a little bit more efficient than least squares using gradient descent.
 Only results obtained with this latter will be reported.
@@
 In order to evaluate our different models, we first have split our data ($\mathbf{X}$) in two sets of $80\%$ and $20\%$, for the training and the test sets.
-We choose this percentage in order to have still enough data for the training set and also enough in order to estimate correctly our models on the testing set with a concrete size.
+We choose this percentage in order to have still enough data for the training set and also enough to estimate correctly our models on the testing set with a concrete size.
 We learn only on the training set and then use the test set in order to estimate the error using the \textit{Root Mean Square Error}.
-Moreover, we have also used k-cross validation (see next section for the values of k) to tune the parameters up like lambda of the different polynomial degrees.
-We repeated the experiment $30$ times with different seed in order to split the data in a unique random way for each trial and so .
+Moreover, we have also used k-cross validation (see next section for the values of k) to tune the parameters up like lambda or the different polynomial degrees.
+We repeated the experiment $30$ times with different seeds in order to split the data in a unique random way for each trial and so try to have an unbiased estimation error.
 Finally, we made vary the lambda from $10^{-5}$ to $10^{5}$ with $500$ values in-between.
@@
 
-The figure \ref{fig:models} represents the RMSE for each models.
+The Figure \ref{fig:models} represents the RMSE for each models.
 The first model used is the constant one, which will allow us to compare improvements with other models.
@@
 As expected, this is much better than previous model because we take into account the input features.
-Next model is similar to the second one except that we add normalization on the real inputs (it doesn't make sense to also normalize categorical variables).
+Next model (third) is similar to the second one except that we add normalization on the real inputs (it doesn't make sense to also normalize categorical variables).
 Normalization is generally a good thing and often a necessity depending on the algorithms used (e.g.
 gradient descent ones).
-It improves only slightly but because normalization is a good practice, we keep it.
+It improves only slightly but because normalization is a good practice, we have decided to keep it.
 
@@
 We must be careful because it increases the ill-conditioning !
-The fourth model consists of using a constant model but this time separating the data into the three clusters identified during the exploratory data analysis and computing the mean for each cluster.
+The fourth model consists of using a constant model but this time separating the data into the three clusters identified during the exploratory data analysis and then computing the mean for each cluster.
 We can observe a big improvement, justifying our choice to split the data into clusters.
-The next model is about doing a ridge regression per cluster, each one having its own lambda and its own normalization.
+The next model (fifth) is about doing a ridge regression per cluster, each one having its own lambda and its own normalization.
 We also use k-cross validation where the first cluster has a $k_1=10$ because we have a lot of data ($1937$), the second $k_2=6$, because we don't have a lot of data ($563$) and the last $k_3=4$ because we really a small number of data ($273$) and so we have to choose a smaller k.
@@
 Unfortunately, we didn't get major improvements.
-We think that the reason why dummy encoding doesn't help is because we have a low bias and so, having more features doesn't help.
+We think that the reason why dummy encoding doesn't help is because we don't have a high bias and so, having more features doesn't help.
 
@@
 Categorical variables aren't concerned, they are just replicated for each degree, because it doesn't make any sense to build polynomial for them.
-We normalize only when we did our feature transformations, in order to avoid loosing information.
-We have to be careful about the polynomial to avoid the case where $D > N$ (especially for the cluster two and three, because there are much smaller than the first one).
+We normalize after the feature transformations, in order to avoid loosing information.
+We have to be careful about the polynomial degrees to avoid the case where $D > N$ (especially for the cluster two and three, because there are much smaller than the first one).
 Moreover, if we don't observe significant improvement between degrees, we keep the lowest one.
@@
 We can see that these feature transformations significantly improve the prediction accuracy compared to other models.
-However, dummy encoding seems to more efficient, and so our final model will be the seventh.
+However, dummy encoding doesn't seem to more efficient, and so our final model will be the seventh.
 
@@
 In hope to minimize the error, we tried different feature transformations.
-The first try was add polynomial of them self as $X = [X X^2 X^3 ...]$ for which each cluster has is own degree.
-We made vary the different degrees and also the lambdas in order to keep the lambda which minimize the test error.
+The first try was add polynomial of them self as $X = [X X^2 X^3 ...]$ for which each cluster has its own degree.
+We made vary the different degrees and also the lambdas in order to keep the lambda which minimizes the test error.
 We also try feature transformations where the degree is a root, but unfortunately it was worst.
@@
 
-We analyzed a regression data set with different methods and concluded that least squares wasn't a good approach due to the ill-conditioning and the least squares gradient descent was to slow and not more efficient than ridge regression.
+We analyzed a regression data set with different methods and concluded that least squares wasn't a good approach due to the ill-conditioning and the least squares gradient descent was to slow and was less efficient than ridge regression.
 Moreover, the separation of the data into clusters was essential in order to decrease the prediction accuracy.
-Unfortunately, dummy encoding didn't help however.
-Using polynomial basis help to decrease the variance.
+Unfortunately, dummy encoding didn't help.
+However using polynomial basis help to decrease the variance.
 Finally, the variance of cluster 1 and 3 are very low, the one for cluster 2 is a bit higher.
-We think that we have bias because we use ridge regression and we don't think it is too high because we have seen that adding features helps and at a certain points it wasn't efficient.
+We think that we have bias because we use ridge regression but we don't think it is too high because we have seen that adding features helps and at a certain points it wasn't efficient.
 The best model we have is the seventh one, and we estimate the RMSE of the prediction to be around 97 with a standard deviation of 35.
-Finally, we think that our model is improvable however, we cannot try other things due to a lack of time.
+Finally, we think that our model is improvable, however we couldn't try other things in oorder to improve the estimation error, due to a lack of time.
 

### 2015-10-26 report/paper.tex
Advance report*

@@
 We first have plot the distribution of our features (plot not shown because too big for the $76$ features).
-As expected, they are not center and we should normalized them.
+As expected, they are not center and we should normalize them.
 The Figure \ref{fig:histogram} shows an histogram of the output ($\mathbf{y}$) and we can conclude our data  seem to be a combination of three Gaussian distributions.
@@
 The best model we have is the seventh one, and we estimate the RMSE of the prediction to be around 97 with a standard deviation of 35.
-Finally, we think that our model is improvable, however we couldn't try other things in oorder to improve the estimation error, due to a lack of time.
+Finally, we think that our model is improvable, however we couldn't try other things in order to improve the estimation error, due to a lack of time.
+
+\section{Classification}
+
+\subsection{Data Description}
+
+The train-data for classification consists of $N = 1500$ input ($\mathbf{X}$) and output ($\mathbf{y}$) data samples.
+Each input sample is a vector $\mathbf{x}_n$ with dimension $D = 25$.
+Out of these $25$ variables, $20$ are real, $3$ categorical with 2 categories, $1$ is categorical with 4 categories and $1$ is categorical with 5 categories.
+
+We also have test-data of size $N=1500$ without their corresponding output.
+Our goal is to produce predictions for those data, as well as an approximation of the test-error (using RMSE, 0-1 loss and logarithmic loss) using logistic regression and penalized logistic regression methods.
+
+\subsection{Data visualization and cleaning}
+
+We first have plot the distribution of our features.
+As expected, they are not center and we should normalize them (good practice because we also use methods using gradient descents).
+Then we plot the histogram of the output which shows that approximately $64.8\%$ of the data are in a class and the other $35.2\%$ in another.
+We also plot the histogram of the categorical variables, which shows us that mostly their values are uniform.
+However, for the real features, we have distinguish 3 types : 1) Look likes a Poisson distribution where the peak is near 0 and the values are mainly positive (like feature 17) 2) Same as one but the values are mainly negative (like feature 5) and so, maybe they have been transformed with a negative sign from a Poisson distribution.
+3) Look likes Gaussian distribution which is not centered (like feature 10).
+The Figure \ref{fig:featuresType} shows their histogram, with their values normalized.
+
+We use dummy encoding for categorical variables, as we did for the regression.
+
+We are also interested about the correlation between the input and output variables.
+The Figure \ref{fig:correlation} shows the correlation.
+We have observed that the features are correlated mainly in the range $[-0.3,0.3]$ and some have higher correlation near $\pm0.4$ and $0.5$ like feature $7$ or $24$.We have decided to not remove features with low correlation because we don't have any pertinent reason to do so.
+
+\begin{figure}[!ht]
+\center
+\subfigure[Histogram of different feature types.
+We can see that the first and second looks like a Poisson distribution (the second has maybe be transformed with negative sign) and the third one a Gaussian distribution.]{\includegraphics[width=2.6in]{figures/featuresType.jpg} \label{fig:featuresType}}
+\hfill
+\subfigure[Plot of the correlation of the features and the outputs.
+Some feature are highly correlated  whereas some have really low correlation.]{\includegraphics[width=2.6in]{figures/correlation.jpg} \label{fig:correlation}}
+\hfill
+\caption{}
+\end{figure}
 

### 2015-10-26 classification/exploratoryDataAnalysis.m

@@ -46,7 +46,7 @@ for feature=1:size(X,2)
     plot(X(idx,feature),y(idx),'ob');
     ylabel('y');
     xlabel(sprintf('Values of feature %d', feature));
-    saveas(f, sprintf('plots/features/featuresVSOuputTo1/feature%d.jpg', feature));
+    % saveas(f, sprintf('plots/features/featuresVSOuputTo1/feature%d.jpg', feature));
 end
 
 idx = find(y == -1);
@@ -55,8 +55,9 @@ for feature=1:size(X,2)
     plot(X(idx,feature),y(idx),'ob');
     ylabel('y');
     xlabel(sprintf('Values of feature %d', feature));
-    saveas(f, sprintf('plots/features/featuresVSOuputTo0/feature%d.jpg', feature));
+    % saveas(f, sprintf('plots/features/featuresVSOuputTo0/feature%d.jpg', feature));
 end
+
 %% Plot histogram of each feature
 for feature=1:size(X,2)
     f = figure;
@@ -72,7 +73,7 @@ for feature=1:size(X,2)
     histogram(X(idx,feature));
     xlabel(sprintf('Feature %d', feature));
     ylabel('Count');
-    saveas(f, sprintf('plots/features/featuresHistogramWhereOutputTo1/feature%d.jpg', feature));
+    % saveas(f, sprintf('plots/features/featuresHistogramWhereOutputTo1/feature%d.jpg', feature));
 end
 
 idx = find(y == -1);
@@ -81,8 +82,32 @@ for feature=1:size(X,2)
     histogram(X(idx,feature));
     xlabel(sprintf('Feature %d', feature));
     ylabel('Count');
-    saveas(f, sprintf('plots/features/featuresHistogramWhereOutputTo0/feature%d.jpg', feature));
+    % saveas(f, sprintf('plots/features/featuresHistogramWhereOutputTo0/feature%d.jpg', feature));
 end
+
+% For the report
+f = figure;
+X_ = normalizedData(X);
+subplot(1, 3, 1);
+histogram(X_(:,17));
+xlabel('First type');
+ylabel('Count');
+xlim([-5 5]);
+
+subplot(1, 3, 2);
+histogram(X_(:,5));
+xlabel('Second type');
+xlim([-5 5]);
+ylabel('Count');
+
+subplot(1, 3, 3);
+histogram(X_(:,10));
+xlabel('Third type');
+ylabel('Count');
+xlim([-5 5]);
+
[…]

### 2015-10-26 classification/main.m
add some models to do

@@ -11,6 +11,10 @@ X = X_train;
 N = length(y);
 
 % 1) Constant baseline
+% 2) Logistic regression
+% 3) + Output either 0 or 1
+% 4) + without categorical variables
+% 5) + Removing outliers
 
 numberOfExperiments = 30;
 proportionOfTraining = 0.8;

### 2015-10-31 report/paper.tex
add some models to do
Advance report

@@
 \author{
-Diego Antognin and Jason Racine \\
+Diego Antognini \& Jason Racine\\
 EPFL \\
@@
 For the regression data set we have observed that it was essential to separate the data into three clusters and doing a ridge regression on each of them due to the ill-condition matrix and the rapidity of tuning the algorithm.
-Moreover feature transformations like using polynomial basis was a key to decrease the prediction accuracy.
+Moreover feature transformations like using polynomial basis was a key to decrease the prediction error.
+For the classification data set we have observed using logistic regression was as good as penalized logistic regression.
+Transformation of features in order to make them look like Gaussian helps a lot as well as adding a polynomial basis help a lot to improve prediction accuracy.
 \end{abstract}
@@
 We first have plot the distribution of our features (plot not shown because too big for the $76$ features).
-As expected, they are not center and we should normalize them.
+As expected, they are not centered and we should normalize them.
 The Figure \ref{fig:histogram} shows an histogram of the output ($\mathbf{y}$) and we can conclude our data  seem to be a combination of three Gaussian distributions.
@@
 
-\begin{figure}[!ht]
+\begin{figure}
 \center
@@
 We learn only on the training set and then use the test set in order to estimate the error using the \textit{Root Mean Square Error}.
-Moreover, we have also used k-cross validation (see next section for the values of k) to tune the parameters up like lambda or the different polynomial degrees.
+Moreover, we have also used $k$-fold cross validation (see next section for the values of $k$) to tune the parameters up like lambda or the different polynomial degrees.
 We repeated the experiment $30$ times with different seeds in order to split the data in a unique random way for each trial and so try to have an unbiased estimation error.
@@
 In hope to minimize the error, we tried different feature transformations.
-The first try was add polynomial of them self as $X = [X X^2 X^3 ...]$ for which each cluster has its own degree.
+The first try was adding polynomial of them self as $X = [X X^2 X^3 ...]$ for which each cluster has its own degree.
 We made vary the different degrees and also the lambdas in order to keep the lambda which minimizes the test error.
 We also try feature transformations where the degree is a root, but unfortunately it was worst.
-We also wanted to try to just pow features with different degrees (without add polynomial of themselves) but we couldn't do it due to a lack of time.
+We also wanted to try to just pow features with different degrees (without adding polynomial of themselves) but we couldn't do it due to a lack of time.
 
@@
 We analyzed a regression data set with different methods and concluded that least squares wasn't a good approach due to the ill-conditioning and the least squares gradient descent was to slow and was less efficient than ridge regression.
-Moreover, the separation of the data into clusters was essential in order to decrease the prediction accuracy.
+Separation of the data into clusters was essential in order to decrease the prediction accuracy.
 Unfortunately, dummy encoding didn't help.
-However using polynomial basis help to decrease the variance.
+However using polynomial basis helped to decrease the bias.
 Finally, the variance of cluster 1 and 3 are very low, the one for cluster 2 is a bit higher.
 We think that we have bias because we use ridge regression but we don't think it is too high because we have seen that adding features helps and at a certain points it wasn't efficient.
-The best model we have is the seventh one, and we estimate the RMSE of the prediction to be around 97 with a standard deviation of 35.
+The best model we have is the seventh one, and we estimate the \textit{RMSE} of the prediction to be around $97$ with a standard deviation of $35$.
 Finally, we think that our model is improvable, however we couldn't try other things in order to improve the estimation error, due to a lack of time.
@@
 We also have test-data of size $N=1500$ without their corresponding output.
-Our goal is to produce predictions for those data, as well as an approximation of the test-error (using RMSE, 0-1 loss and logarithmic loss) using logistic regression and penalized logistic regression methods.
+Our goal is to produce predictions for those data, as well as an approximation of the test-error (using \textit{RMSE}, \textit{0-1 loss} and \textit{logarithmic loss}) using logistic regression and penalized logistic regression methods.
 
@@
 We also plot the histogram of the categorical variables, which shows us that mostly their values are uniform.
-However, for the real features, we have distinguish 3 types : 1) Look likes a Poisson distribution where the peak is near 0 and the values are mainly positive (like feature 17) 2) Same as one but the values are mainly negative (like feature 5) and so, maybe they have been transformed with a negative sign from a Poisson distribution.
+However, for the real features, we have distinguished 3 types : 1) Look likes a Poisson distribution where the peak is near 0 and the values are mainly positive (like feature 17) 2) Same as one but the values are mainly negative (like feature 5) and so, maybe they have been transformed with a negative sign from a Poisson distribution.
 3) Look likes Gaussian distribution which is not centered (like feature 10).
-The Figure \ref{fig:featuresType} shows their histogram, with their values normalized.
+The Figure \ref{fig:featuresType} shows their histogram with their values normalized.
 
@@
 
-\begin{figure}[!ht]
+Finally, our matrix ($\mathbf{X}$) is a full-rank matrix, which is a good news for logistic regression.
+
+\begin{figure}
 \center
 \subfigure[Histogram of different feature types.
-We can see that the first and second looks like a Poisson distribution (the second has maybe be transformed with negative sign) and the third one a Gaussian distribution.]{\includegraphics[width=2.6in]{figures/featuresType.jpg} \label{fig:featuresType}}
+We can see that the first and second looks like a Poisson distribution (the second has maybe be transformed with negative sign) and the third one a Gaussian distribution.]{\includegraphics[width=2.65in]{figures/featuresType.jpg} \label{fig:featuresType}}
 \hfill
 \subfigure[Plot of the correlation of the features and the outputs.
-Some feature are highly correlated  whereas some have really low correlation.]{\includegraphics[width=2.6in]{figures/correlation.jpg} \label{fig:correlation}}
+Some feature are highly correlated  whereas some have really low correlation.]{\includegraphics[width=2.65in]{figures/correlation.jpg} \label{fig:correlation}}
 \hfill
+\subfigure[Boxplot of our models, using the 0-1 Loss.
+Each model is represented with a box where we can see their mean and their standard deviation.]{\includegraphics[width=5in]{figures/models_classification.jpg}} \label{fig:models_classification}
 \caption{}
@@
 
+\subsection{Logistic Regression}
+
+Since our matrix is full-rank using logistic regression was appropriate.
+Because logistic regression is more efficient in term of complexity and because we didn't need to penalize some betas, we report only the results using this method.
+However, using penalized logistic regression gave us similar results although it was much slower.
+
+\subsection{Evaluation methods}
+
+In order to evaluate our different models, the method is similar that the one used in regression.
+We split our data ($\mathbf{X}$) in two sets of $80\%$ and $20\%$, we compute the test-error using the \textit{Root Mean Square Error}, \textit{0-1 loss} and \textit{logarithmic loss}.
+Because all the errors grows similarly and because we think that \textit{0-1 loss} is the easiest error to visualize, we have decided to report only this error.
+To find the learning step $\alpha$, we have used $10$ fold cross validation because we have enough data to use this $k$.
+We repeated the experiment $30$ times with different seeds in order to split the data in a unique random way for each trial and so try to have an unbiased estimation error.
+
+\subsection{Model comparison}
+
+The Figure \ref{fig:models_classification} represents the \textit{0-1 Loss} for each models.
+The first model used is the constant one, which will allow us to compare improvements with other models.
+It simply returns $1$ and doesn't depend on the inputs.
+Our second model does a logistic regression, using normalization, transforming the output is the range $[0,1]$ and using only real features.
+We can that it clearly outperforms the constant model.
+In the next model (third) we added the categorical variables which didn't improve the error and increase a little bit the variance.
+However, because it could be important features, we have decided to keep them.
+
+The fourth model consists in using dummy encoding for the categorical variable.
+This times, the error was improved and the variance has diminished.
+The next model (fifth) adds a polynomial basis of degree 2.
+We can observe the the variance as increased, we have some outliers (five in totel, there might be hidden due to the legend) and the error has increased.
+We think that the bias is better but due to its high variance, we consider as a bad model.
+
+In order to try to improve the error, we have transformed the feature (more information in the next section) with the formula using the absolute values (see next section) in order to have more Gaussian-like distribution.
+The sixth model represents this transformation.
+We can observe that the error is a little bit higher than model 4 but the mean is more stable and the variance lower.
+Because we assume that data sample sare generated from a Gaussian distribution, we decided to continue this way.
+The seventh one adds a polynomial basis of degree 2, which improves the error and still decrease the variance.
+Moreover, because we use more features, we think that our model has also a lower bias.
+
+The two last models (eighth and ninth) use another transformation where negative values weren't transformed into positive one, holding them with the formula $-\sqrt[4]{-x}$ and try using also a polynomial basis.
+However, the error wasn't as good as the seventh model, surely because the transformed feature look like less Gaussian than the transformation using the absolute values.
+We can also see that the variance seems also higher.
+
+Our final model will be the seventh.
+
+\subsection{Feature transformations}
+
+In hope to minimize the error, we tried different feature transformations.
+The first try was add polynomial of them self as $X = [X X^2 X^3 ...]$.
+With the data not transformed into Gaussian like distribution, we didn't get an improvement (even with degree 3).
+We then try to transform the features which look like a Poisson distribution in order to transform them into a Gaussian-like distribution.
+We first use the formula $\sqrt[4]{|x|}$ which gave us distribution which was similar to Gaussian even if we transformed the negative values into positive.
+Mixed with polynomial of features at degree 2 allowed us to have an important improvement for the error.
+We also try to keep the negative values, this time using the formula $-\sqrt[4]{-x}$ but unfortunately, it didn't help.
+We think the reason is first that distribution was more a superposition of two Gaussian-like, and secondly  because we try to do an operation to data which are not applicable.
+Using a degree greater than 2 didn't improve the error, surely because our bias is not too high.
+
+We also try to remove some uncorrelated features, it didn't improve surely for the same reasons as specified in regression section. We didn't have time to try to remove combination of features.
+
+\subsection{Summary}
+
+We analyzed a classification data set with different methods and concluded that logistic regression was appropriate because the input matrix was full-rank.
+Moreover, using the penalized logistic regression gave us similar results.
+Using normalization was essential, and also transform the output in the range $[0,1]$.
+Adding categorical variables with dummy encoding helps.
+The key of our model was to transform features which look like Poisson distribution into a Gaussian like distribution using $\sqrt[4]{|x|}$ and using a polynomial basis of degree 2.
+All together, they give us our best model which is the seventh one.
+Finally, we think that our model is improvable because we didn't remove outliers because we couldn't find such outliers in all dimensions (or most of the dimensions).
+We estimate the \textit{RMSE} of the prediction to be around $0.329$ with a standard deviation of $0.028$.
+For the \textit{0-1 loss} around $0.109$ with standard deviation of $0.018$.
+Finally for the \textit{logarithmic loss} around $0.509$ with standard deviation $0.018$.
+
+\subsubsection*{Acknowledgments}
+All the code, as well as the report were written by both members.
+We worked together in the regression and classification, exchanged our ideas in order to test models and avoiding loosing our time.
+We would like to thank the teaching assistants who were very helpful during the office hours.
+
 \end{document}

### 2015-10-31 classification/exploratoryDataAnalysis.m

@@ -59,31 +59,31 @@ for feature=1:size(X,2)
 end
 
 %% Plot histogram of each feature
-for feature=1:size(X,2)
-    f = figure;
-    histogram(X(:,feature));
-    xlabel(sprintf('Feature %d', feature));
-    ylabel('Count');
-    % saveas(f, sprintf('plots/features/featuresHistogram/feature%d.jpg', feature));
-end
-
-idx = find(y == 1);
-for feature=1:size(X,2)
-    f = figure;
-    histogram(X(idx,feature));
-    xlabel(sprintf('Feature %d', feature));
-    ylabel('Count');
-    % saveas(f, sprintf('plots/features/featuresHistogramWhereOutputTo1/feature%d.jpg', feature));
-end
-
-idx = find(y == -1);
-for feature=1:size(X,2)
-    f = figure;
-    histogram(X(idx,feature));
-    xlabel(sprintf('Feature %d', feature));
-    ylabel('Count');
-    % saveas(f, sprintf('plots/features/featuresHistogramWhereOutputTo0/feature%d.jpg', feature));
-end
+% for feature=1:size(X,2)
+%     f = figure;
+%     histogram(X(:,feature));
+%     xlabel(sprintf('Feature %d', feature));
+%     ylabel('Count');
+%     % saveas(f, sprintf('plots/features/featuresHistogram/feature%d.jpg', feature));
+% end
+% 
+% idx = find(y == 1);
+% for feature=1:size(X,2)
+%     f = figure;
+%     histogram(X(idx,feature));
+%     xlabel(sprintf('Feature %d', feature));
+%     ylabel('Count');
+%     % saveas(f, sprintf('plots/features/featuresHistogramWhereOutputTo1/feature%d.jpg', feature));
+% end
+% 
+% idx = find(y == -1);
+% for feature=1:size(X,2)
+%     f = figure;
+%     histogram(X(idx,feature));
+%     xlabel(sprintf('Feature %d', feature));
+%     ylabel('Count');
+%     % saveas(f, sprintf('plots/features/featuresHistogramWhereOutputTo0/feature%d.jpg', feature));
+% end
 
 % For the report
 f = figure;
@@ -93,20 +93,26 @@ histogram(X_(:,17));
 xlabel('First type');
 ylabel('Count');
 xlim([-5 5]);
[…]

### 2015-10-31 classification/main.m

@@ -589,8 +589,9 @@ h_legend = legend(findobj(gca,'Tag','Box'), ...
 set(h_legend,'FontSize',12);
 set(gca, 'XGrid','on')
 set(gca, 'YGrid','on')
-%ylim([0 3000])
-%set(gca,'YTick',0:100:3000)
+set(gca,'LineWidth',1.5);
+ylim([0 0.45])
+set(gca,'YTick',0:0.05:0.45)
 xlabel('Model');
 ylabel('0-1 Loss');
 print('../report/figures/models_classification','-djpeg','-noui')

### 2015-10-31 regression/main.m

@@ -367,9 +367,9 @@ h_legend = legend(findobj(gca,'Tag','Box'), ...
 '6 + Dummy encoding', ...
 '7 Same as 5 + polynomial basis', ...
 '8 Same as 6 + polynomial basis');
-set(h_legend,'FontSize',12);
 set(gca, 'XGrid','on')
 set(gca, 'YGrid','on')
+set(gca,'LineWidth',1.5);
 ylim([0 3000])
 set(gca,'YTick',0:100:3000)
 xlabel('Model');

### 2015-10-31 regression/old/runTestPoly.m

@@ -47,7 +47,10 @@ function [mean_err_tr, mean_err_te] = runTestPoly(K, y_, X_, lambda, idxCls)
     legend('Train error', 'Test error', 'Location', 'northwest');
     xlabel('Degree');
     ylabel('RMSE');
-    print('../report/figures/degrees_cls1','-djpeg','-noui')
+    set(gca, 'XGrid','on');
+    set(gca,'fontsize', 18);
+    set(gca,'LineWidth',1.5);
+    print('../../report/figures/degrees_cls1','-djpeg','-noui')
     
     mean_err_tr = mean_err_tr(end);
     mean_err_te = mean_err_te(end);

### 2015-10-31 report/paper.tex
Nearly finish report

@@
 This report provides a summary of the project one of the PCML class.
-The project consists of doing regression and classification on data.
+The project consists of doing regression and classification.
 For the regression data set we have observed that it was essential to separate the data into three clusters and doing a ridge regression on each of them due to the ill-condition matrix and the rapidity of tuning the algorithm.
@@
 For the classification data set we have observed using logistic regression was as good as penalized logistic regression.
-Transformation of features in order to make them look like Gaussian helps a lot as well as adding a polynomial basis help a lot to improve prediction accuracy.
+Transformation of features in order to make them look like Gaussian as well as adding a polynomial basis help a lot to improve prediction accuracy.
 \end{abstract}
@@
 
-We first have plot the distribution of our features (plot not shown because too big for the $76$ features).
+We first have plot the distribution of our features.
 As expected, they are not centered and we should normalize them.
@@
 
-To separate the data, we have observed that the feature $2$ and $16$ could help us.
-With Figure \ref{fig:feature2} we can see how we can use the second feature to separate cluster one to the others and observe $11$ misclassified data, which we will remove them in order to not corrupt our model.
-For the two other clusters, we need to observe the feature $16$.
-With Figure \ref{fig:feature16}, we can observe $14$ misclassified data.
-They also will be consider as outliers and removed.
-We set the threshold in order to minimize the number of misclassified data samples.
-Those thresholds are $0.42$ for the feature $2$ and $1.17$ for the feature $16$.
-So the final size of our clusters are $1937, 563$ and $273$.
-
-We are also interested about the correlation between the input and output variables.
-We have observed the correlation for each cluster and conclude that for the first cluster, they are mainly in the range $[-0.1,0.1]$, except two features which are highly correlated.
-For the second and third cluster, also mainly in range $[-0.1,0.1]$ but this time, there are more correlated features (around 15).
-Moreover, the features seem not have correlation between them.
-
-We use dummy encoding for the categorical variables (for a categorical variable of size $k$, we need $k-1$ features), which gives us a total of $93$ input variable.
-
-We can note that the rank of our input matrix $\mathbf{X}$ is rank-deficient with a rank of 66 instead of 76, and is rank-deficient of 65 for each cluster.
-
-\begin{figure}
+\begin{figure}[!ht]
+\vspace{-2em}
 \center
@@
 The third degree seems the best model to choose.]{\includegraphics[width=2.7in]{figures/degrees_cls1.jpg}\label{fig:degrees_cls1}}
-\hfill
-\subfigure[Boxplot of our models, using the RMSE.
-Each model is represented with a box where we can see their mean and their standard deviation.]{\includegraphics[width=5in]{figures/models.jpg}} \label{fig:models}
+\vspace{-1em}
 \caption{}
@@
 
+To separate the data, we have observed that the feature $2$ and $16$ could help us.
+With Figure \ref{fig:feature2} we can see how we can use the second feature to separate cluster one to the others and observe $11$ misclassified data, which we will remove them in order to not corrupt our model.
+For the two other clusters, we need to observe the feature $16$.
+With Figure \ref{fig:feature16}, we can observe $14$ misclassified data.
+They also will be considered as outliers and removed.
+We set the threshold in order to minimize the number of misclassified data samples.
+Those thresholds are $0.42$ for the feature $2$ and $1.17$ for the feature $16$.
+So the final size of our clusters are $1937, 563$ and $273$.
+
+We are also interested about the correlation between the input and output variables.
+We have observed the correlation for each cluster and conclude that for the first cluster, they are mainly in the range $[-0.1,0.1]$, except two features which are highly correlated.
+For the second and third cluster, also mainly in range $[-0.1,0.1]$ but this time, there are more correlated features (around 15).
+However, the features seem not have correlation between them.
+
+We use dummy encoding for the categorical variables (for a categorical variable of size $k$, we need $k-1$ features), which gives us a total of $93$ input variables.
+
+We can note that the rank of our input matrix $\mathbf{X}$ is rank-deficient with a rank of 66 instead of 76, and is rank-deficient of 65 for each cluster.
+\vspace{-0.5em}
 \subsection{Ridge Regression}
@@
 Moreover, the results obtained were less good compared to the ridge regression method, and so we will not report those.
-The method using the least squares with gradient descent was appropriated, but compared to the ridge regression, it was less efficient and was also much slower to tune correctly the learning step.
-This is the reason why we will not report the results using this method.
-Therefore, we have so used the ridge regression method, which is suitable because we could lift the eigenvalues in order to avoid ill-conditioned matrix.
-It is also faster and a little bit more efficient than least squares using gradient descent.
-Only results obtained with this latter will be reported.
+The method using the least squares with gradient descent is appropriated as well as ridge regression method, which is suitable because we could lift the eigenvalues in order to avoid ill-conditioned matrix.
+It is also faster and a little bit more efficient than least squares using gradient descent and so, only results obtained with this latter will be reported.
 
-\subsubsection{Evaluation methods}
+\subsection{Evaluation methods}
 
-In order to evaluate our different models, we first have split our data ($\mathbf{X}$) in two sets of $80\%$ and $20\%$, for the training and the test sets.
-We choose this percentage in order to have still enough data for the training set and also enough to estimate correctly our models on the testing set with a concrete size.
-We learn only on the training set and then use the test set in order to estimate the error using the \textit{Root Mean Square Error}.
-Moreover, we have also used $k$-fold cross validation (see next section for the values of $k$) to tune the parameters up like lambda or the different polynomial degrees.
+In order to evaluate our different models, we have split our data ($\mathbf{X}$) in two sets of $80\%$ and $20\%$, for the training and the test sets.
+We choose this percentage in order to have still enough data for the training set and also enough to estimate correctly our models on the testing set.
+We learn only on the training set and then use the test set in order to estimate the error using the \textit{RMSE}.
+Moreover, we have also used $k$ fold cross validation (see next section for the values of $k$) to tune the parameters up like lambda or the different polynomial degrees.
 We repeated the experiment $30$ times with different seeds in order to split the data in a unique random way for each trial and so try to have an unbiased estimation error.
@@
 
-\subsubsection{Model comparison}
+\subsection{Model comparison}
 
-The Figure \ref{fig:models} represents the RMSE for each models.
+The Figure \ref{fig:models} represents the \textit{RMSE} for each models.
 The first model used is the constant one, which will allow us to compare improvements with other models.
@@
 Next model (third) is similar to the second one except that we add normalization on the real inputs (it doesn't make sense to also normalize categorical variables).
-Normalization is generally a good thing and often a necessity depending on the algorithms used (e.g.
+Normalization is often a necessity depending on the algorithms used (e.g.
 gradient descent ones).
-It improves only slightly but because normalization is a good practice, we have decided to keep it.
+It improves only slightly but because this is a good practice, we decided to keep it.
+
+\begin{figure}
+\vspace{-1em}
+\center
+\includegraphics[width=4in]{figures/models.jpg} 
+\vspace{-1em}
+\caption{Boxplot of our models, using the RMSE. Each model is represented with a box where we can see their mean and their standard deviation.}
+\label{fig:models}
+\end{figure}
 
 At this stage, we want to try to separate our data into clusters.
-We must be careful because it increases the ill-conditioning !
-The fourth model consists of using a constant model but this time separating the data into the three clusters identified during the exploratory data analysis and then computing the mean for each cluster.
+The fourth model consists of using a constant model but this time separating the data into three clusters and then computing the mean for each cluster.
 We can observe a big improvement, justifying our choice to split the data into clusters.
-The next model (fifth) is about doing a ridge regression per cluster, each one having its own lambda and its own normalization.
-We also use k-cross validation where the first cluster has a $k_1=10$ because we have a lot of data ($1937$), the second $k_2=6$, because we don't have a lot of data ($563$) and the last $k_3=4$ because we really a small number of data ($273$) and so we have to choose a smaller k.
+The next model (fifth) is about doing a ridge regression per cluster, each one having its own lambda.
+We also use $k$ fold cross validation where the first cluster has a $k_1=10$ because we have a lot of data ($1937$), the second $k_2=6$, because we don't have a lot of data ($563$) and the last $k_3=4$ because we really a small number of data ($273$) and so we have to choose a smaller k.
 As expected we obtained another big improvement justifying again the choice of splitting the data.
@@
 The sixth model consists of using dummy encoding for the categorical variables.
-As explained in the exploratory data analysis, we use $k-1$ features to represent a $k$ categorical variable.
-This is a common practice.
 Unfortunately, we didn't get major improvements.
@@
 Categorical variables aren't concerned, they are just replicated for each degree, because it doesn't make any sense to build polynomial for them.
-We normalize after the feature transformations, in order to avoid loosing information.
-We have to be careful about the polynomial degrees to avoid the case where $D > N$ (especially for the cluster two and three, because there are much smaller than the first one).
+We have to be careful about the polynomial degrees to avoid the case where $D > N$.
 Moreover, if we don't observe significant improvement between degrees, we keep the lowest one.
@@
 When we add higher degrees, we can directly observe overfitting because suddenly the training error decreases to a value near zero and the test errors grows exponentially, so with this manner we know that our degrees are not to high.
-The eighth model is similar to the seventh one, except that this time it uses dummy encoding and has lower degrees : $3,5$ and $2$, due to the number of input variables ($92$ vs $76$).
-We can see that these feature transformations significantly improve the prediction accuracy compared to other models.
+The eighth model is similar to the seventh one, except that this time it uses dummy encoding and has lower degrees : $3,5$ and $2$, due to the number of input variables.
+We can see that feature transformations significantly improve the test error compared to other models.
 However, dummy encoding doesn't seem to more efficient, and so our final model will be the seventh.
 
-\subsubsection{Feature transformations}
+\subsection{Feature transformations}
 
@@
 We also try feature transformations where the degree is a root, but unfortunately it was worst.
-We also wanted to try to just pow features with different degrees (without adding polynomial of themselves) but we couldn't do it due to a lack of time.
 
@@
 
-\subsubsection{Summary}
+\subsection{Summary}
 
 We analyzed a regression data set with different methods and concluded that least squares wasn't a good approach due to the ill-conditioning and the least squares gradient descent was to slow and was less efficient than ridge regression.
-Separation of the data into clusters was essential in order to decrease the prediction accuracy.
+Separation of the data into clusters was essential in order to decrease the test error.
 Unfortunately, dummy encoding didn't help.
@@
 \subsection{Data visualization and cleaning}
+\begin{figure}
+\center
+\subfigure[Histogram of different feature types.
+We can see that the first and second look like a Poisson distribution (the second has maybe be transformed with negative sign) and the third one a Gaussian distribution.]{\includegraphics[width=2.65in]{figures/featuresType.jpg} \label{fig:featuresType}}
+\hfill
+\subfigure[Plot of the correlation of the features and the outputs.
+Some feature are highly correlated  whereas some have really low correlation.]{\includegraphics[width=2.65in]{figures/correlation.jpg} \label{fig:correlation}}
+\caption{}
+\end{figure}
 
 We first have plot the distribution of our features.
-As expected, they are not center and we should normalize them (good practice because we also use methods using gradient descents).
+As expected, they are not centered and we should normalize them (good practice because we also use methods using gradient descents).
 Then we plot the histogram of the output which shows that approximately $64.8\%$ of the data are in a class and the other $35.2\%$ in another.
 We also plot the histogram of the categorical variables, which shows us that mostly their values are uniform.
-However, for the real features, we have distinguished 3 types : 1) Look likes a Poisson distribution where the peak is near 0 and the values are mainly positive (like feature 17) 2) Same as one but the values are mainly negative (like feature 5) and so, maybe they have been transformed with a negative sign from a Poisson distribution.
-3) Look likes Gaussian distribution which is not centered (like feature 10).
+However, for the real features, we have distinguished 3 types : 1) Look likes a Poisson distribution where the peak is near 0 and the values are mainly positive 2) Same as one but the values are mainly negative and so, maybe they have been transformed with a negative sign from a Poisson distribution.
+3) Look likes Gaussian distribution which is not centered.
 The Figure \ref{fig:featuresType} shows their histogram with their values normalized.
 
-We use dummy encoding for categorical variables, as we did for the regression.
+We use dummy encoding for categorical variables which gives us a total of $31$ input variables.
 
@@
 
-\begin{figure}
-\center
-\subfigure[Histogram of different feature types.
-We can see that the first and second looks like a Poisson distribution (the second has maybe be transformed with negative sign) and the third one a Gaussian distribution.]{\includegraphics[width=2.65in]{figures/featuresType.jpg} \label{fig:featuresType}}
-\hfill
-\subfigure[Plot of the correlation of the features and the outputs.
-Some feature are highly correlated  whereas some have really low correlation.]{\includegraphics[width=2.65in]{figures/correlation.jpg} \label{fig:correlation}}
-\hfill
-\subfigure[Boxplot of our models, using the 0-1 Loss.
-Each model is represented with a box where we can see their mean and their standard deviation.]{\includegraphics[width=5in]{figures/models_classification.jpg}} \label{fig:models_classification}
-\caption{}
-\end{figure}
-
 \subsection{Logistic Regression}
@@
 Since our matrix is full-rank using logistic regression was appropriate.
-Because logistic regression is more efficient in term of complexity and because we didn't need to penalize some betas, we report only the results using this method.
-However, using penalized logistic regression gave us similar results although it was much slower.
+Because logistic regression is more efficient in term of complexity and that we don't need to penalize some betas, we report only the results using this method.
+However, using penalized logistic regression gave us similar results.
 
@@
 In order to evaluate our different models, the method is similar that the one used in regression.
-We split our data ($\mathbf{X}$) in two sets of $80\%$ and $20\%$, we compute the test-error using the \textit{Root Mean Square Error}, \textit{0-1 loss} and \textit{logarithmic loss}.
-Because all the errors grows similarly and because we think that \textit{0-1 loss} is the easiest error to visualize, we have decided to report only this error.
+We split our data ($\mathbf{X}$) in two sets of $80\%$ and $20\%$, we compute the test-error using the \textit{RMSE}, \textit{0-1 loss} and \textit{logarithmic loss}.
+Because all the errors grows similarly and we think that \textit{0-1 loss} is the easiest error to visualize, we have decided to report only this error.
 To find the learning step $\alpha$, we have used $10$ fold cross validation because we have enough data to use this $k$.
-We repeated the experiment $30$ times with different seeds in order to split the data in a unique random way for each trial and so try to have an unbiased estimation error.
+We repeated the experiment $30$ times with different seeds in order to split the data in a unique random way for each trial.
 
@@
 
-The Figure \ref{fig:models_classification} represents the \textit{0-1 Loss} for each models.
+The Figure \ref{fig:models_classification} represents the \textit{0-1 loss} for each models.
 The first model used is the constant one, which will allow us to compare improvements with other models.
 It simply returns $1$ and doesn't depend on the inputs.
-Our second model does a logistic regression, using normalization, transforming the output is the range $[0,1]$ and using only real features.
-We can that it clearly outperforms the constant model.
+Our second model does a logistic regression, using normalization, using only real features and transforming the output in the range $[0,1]$ because the logistic regression results are in this range.
+We can see that it clearly outperforms the constant model.
 In the next model (third) we added the categorical variables which didn't improve the error and increase a little bit the variance.
 However, because it could be important features, we have decided to keep them.
+
+\begin{figure}
+\center
+\includegraphics[width=4in]{figures/models_classification.jpg} 
+\vspace{-1em}
+\caption{Boxplot of our models, using the 0-1 Loss. Each model is represented with a box where we can see their mean and their standard deviation.}
+\label{fig:models_classification}
+\vspace{-0.5em}
+\end{figure}
 
@@
 The next model (fifth) adds a polynomial basis of degree 2.
-We can observe the the variance as increased, we have some outliers (five in totel, there might be hidden due to the legend) and the error has increased.
-We think that the bias is better but due to its high variance, we consider as a bad model.
+We can observe the the variance as increased, we have some outliers (five in total, there might be hidden due to the legend) and the error has increased.
+We think that the bias is better but due to its high variance, we consider it as a bad model.
 
-In order to try to improve the error, we have transformed the feature (more information in the next section) with the formula using the absolute values (see next section) in order to have more Gaussian-like distribution.
+In order to try to improve the error, we have transformed the feature (more information in the next section) with the formula $\sqrt[4]{|x|}$ in order to have more Gaussian-like distribution.
 The sixth model represents this transformation.
 We can observe that the error is a little bit higher than model 4 but the mean is more stable and the variance lower.
-Because we assume that data sample sare generated from a Gaussian distribution, we decided to continue this way.
-The seventh one adds a polynomial basis of degree 2, which improves the error and still decrease the variance.
+Because we assume that data samples are generated from a Gaussian distribution, we decided to continue this way.
+The seventh one adds a polynomial basis of degree 2, which improves the error and still decreases the variance.
 Moreover, because we use more features, we think that our model has also a lower bias.
@@
 We can also see that the variance seems also higher.
-
 Our final model will be the seventh.
@@
 In hope to minimize the error, we tried different feature transformations.
-The first try was add polynomial of them self as $X = [X X^2 X^3 ...]$.
+The first try was add polynomial of them self as $X = [X X^2]$.
 With the data not transformed into Gaussian like distribution, we didn't get an improvement (even with degree 3).
 We then try to transform the features which look like a Poisson distribution in order to transform them into a Gaussian-like distribution.
-We first use the formula $\sqrt[4]{|x|}$ which gave us distribution which was similar to Gaussian even if we transformed the negative values into positive.
+We first use the formula $\sqrt[4]{|x|}$ which gave us good results even if we transformed the negative values into positive.
 Mixed with polynomial of features at degree 2 allowed us to have an important improvement for the error.
-We also try to keep the negative values, this time using the formula $-\sqrt[4]{-x}$ but unfortunately, it didn't help.
-We think the reason is first that distribution was more a superposition of two Gaussian-like, and secondly  because we try to do an operation to data which are not applicable.
+We also try to keep the negative values, using the formula $-\sqrt[4]{-x}$ but unfortunately, it didn't help.
+We think the reason is first that distribution was more a superposition of two Gaussian.
 Using a degree greater than 2 didn't improve the error, surely because our bias is not too high.
-
-We also try to remove some uncorrelated features, it didn't improve surely for the same reasons as specified in regression section. We didn't have time to try to remove combination of features.
 
@@
 Moreover, using the penalized logistic regression gave us similar results.
-Using normalization was essential, and also transform the output in the range $[0,1]$.
+Using normalization was essential as well as transforming the output in the range $[0,1]$.
 Adding categorical variables with dummy encoding helps.
 The key of our model was to transform features which look like Poisson distribution into a Gaussian like distribution using $\sqrt[4]{|x|}$ and using a polynomial basis of degree 2.
-All together, they give us our best model which is the seventh one.
-Finally, we think that our model is improvable because we didn't remove outliers because we couldn't find such outliers in all dimensions (or most of the dimensions).
-We estimate the \textit{RMSE} of the prediction to be around $0.329$ with a standard deviation of $0.028$.
-For the \textit{0-1 loss} around $0.109$ with standard deviation of $0.018$.
-Finally for the \textit{logarithmic loss} around $0.509$ with standard deviation $0.018$.
-
-\subsubsection*{Acknowledgments}
-All the code, as well as the report were written by both members.
-We worked together in the regression and classification, exchanged our ideas in order to test models and avoiding loosing our time.
-We would like to thank the teaching assistants who were very helpful during the office hours.
+We think that our model is improvable due to outliers because we couldn't find such outliers in all dimensions (or most of the dimensions).
+We estimate the mean and standard deviation of the prediction, using the \textit{RMSE},\textit{0-1 loss} and \textit{logarithmic loss} respectively to be around $0.329$, $0.109$ and $0.509$ for the mean, and $0.028$, $0.018$, $0.018$for the standard deviation .
 

### 2015-10-31 report/paper.tex
Add predictions

@@
 However, dummy encoding doesn't seem to more efficient, and so our final model will be the seventh.
+Our final lambdas are $\lambda_1 = 0.0369, \lambda_2 = 1.8303e^{-4}$ and $\lambda_3 = 0.0053$.
 
@@
 We think that we have bias because we use ridge regression but we don't think it is too high because we have seen that adding features helps and at a certain points it wasn't efficient.
-The best model we have is the seventh one, and we estimate the \textit{RMSE} of the prediction to be around $97$ with a standard deviation of $35$.
+The best model we have is the seventh one, and we estimate the \textit{RMSE} of the prediction to be around $115.96$ with a standard deviation of $45.59$.
 Finally, we think that our model is improvable, however we couldn't try other things in order to improve the estimation error, due to a lack of time.

### 2015-10-31 regression/main.m

@@ -307,9 +307,6 @@ s = [1 numberOfExperiments];
 [y_cls1, X_cls1, y_cls2, X_cls2, y_cls3, X_cls3, idx_cls1, idx_cls2, idx_cls3] = preprocess(y, X);
 
 N = length(y_cls1) + length(y_cls2) + length(y_cls3);
-ratio_cls1 = size(X_cls1, 1)/N;
-ratio_cls2 = size(X_cls2, 1)/N;
-ratio_cls3 = size(X_cls3, 1)/N;
 
 err1 = openFile('results/err1', s);
 err2 = openFile('results/err2',s);

### 2015-10-31 regression/prediction.m

@@ -1,0 +1,113 @@
+% Written by Diego Antognini & Jason Racine, EPFL 2015
+% all rights reserved
+
+%clear all;
+%close all;
+
+load('Sydney_regression.mat');
+
+y = y_train;
+X = X_train;
+N = length(y);
+
+k1 = 10;
+k2 = 8;
+k3 = 6;
+
+d1 = 3;
+d2 = 6;
+d3 = 3;
+
+%% COMPUTING ESTIMATE ERROR
+K = 10;
+N = length(y);
+idxCV = splitGetCV(K, N);
+setSeed(28111993);
+shuffle = randperm(N);
+XRandom = X(shuffle,:);
+yRandom = y(shuffle,:);
+
+for k = 1:1:K;
+    [XTr, yTr, XTe, yTe] = splitGetTrTe(yRandom, XRandom, idxCV, k);
+    
+    lambda_cls1 = 0.0369;
+    lambda_cls2 = 1.8303e-4;
+    lambda_cls3 = 0.0053;
+
+    [y_cls1, X_cls1, y_cls2, X_cls2, y_cls3, X_cls3, idx_cls1, idx_cls2, idx_cls3] = preprocess(yTr, XTr, d1, d2, d3);
+
+    tX_cls1 = [ones(length(y_cls1),1) X_cls1];
+    tX_cls2 = [ones(length(y_cls2),1) X_cls2];
+    tX_cls3 = [ones(length(y_cls3),1) X_cls3];
+
+    beta_cls1 = ridgeRegression(y_cls1, tX_cls1, lambda_cls1);
+    beta_cls2 = ridgeRegression(y_cls2, tX_cls2, lambda_cls2);
+    beta_cls3 = ridgeRegression(y_cls3, tX_cls3, lambda_cls3);
+
+    [y_cls1, X_cls1, y_cls2, X_cls2, y_cls3, X_cls3, idx_cls1, idx_cls2, idx_cls3] = preprocess(yTe, XTe, d1, d2, d3, 1);
+
+    tXTe_cls1 = [ones(length(y_cls1),1) X_cls1];
+    tXTe_cls2 = [ones(length(y_cls2),1) X_cls2];
+    tXTe_cls3 = [ones(length(y_cls3),1) X_cls3];
+
+    err_cls1(k) = RMSE(y_cls1, tXTe_cls1*beta_cls1);
+    err_cls2(k) = RMSE(y_cls2, tXTe_cls2*beta_cls2);
[…]

### 2015-10-31 regression/preprocess.m

@@ -2,7 +2,7 @@
 % all rights reserved
 
 % If degrees are not specified, they will get the value 1.
-function [y_cls1, X_cls1, y_cls2, X_cls2, y_cls3, X_cls3, idx_cls1, idx_cls2, idx_cls3] = preprocess(y, X, degree1, degree2, degree3)
+function [y_cls1, X_cls1, y_cls2, X_cls2, y_cls3, X_cls3, idx_cls1, idx_cls2, idx_cls3] = preprocess(y, X, degree1, degree2, degree3, is_prediction)
     if ~exist('degree1', 'var')
         degree1 = 1;
     end
@@ -16,8 +16,28 @@ function [y_cls1, X_cls1, y_cls2, X_cls2, y_cls3, X_cls3, idx_cls1, idx_cls2, id
     end
     
     % Find idx of each cluster
-    [idx_cls1, idx_cls2, idx_cls3] = findClusters(y, X);
-    
+    if ~exist('is_prediction', 'var')
+        [idx_cls1, idx_cls2, idx_cls3] = findClusters(y, X);
+    else
+        idx_cls1 = [];
+        idx_cls2 = [];
+        idx_cls3 = [];
+        X_ = normalizedData(X);
+        for n = 1:1:size(X_,1)
+           res = findCorrespondingCluster(X_(n,:));
+           if res == 1
+               idx_cls1 = [idx_cls1 n];
+           elseif res == 2
+               idx_cls2 = [idx_cls2 n];
+           elseif res == 3
+               idx_cls3 = [idx_cls3 n];
+           end
+        end
+        idx_cls1 = idx_cls1';
+        idx_cls2 = idx_cls2';
+        idx_cls3 = idx_cls3';
+    end
+
     % Find different type of features
     [idx_feature_2, idx_feature_3, idx_feature_4] = findDiscreteFeatures(X);
     idx_feature_real = setdiff(setdiff(setdiff(1:1:size(X,2), idx_feature_2), idx_feature_3), idx_feature_4);
@@ -76,15 +96,15 @@ function [y, X] = preprocessCluster(y_, X_, idx, degree, idx_feature_2, idx_feat
     % Normalize
     y = y_(idx);
     % WITH NORMALIZATION
-    %X = normalizedData(X(:,real_features_idx_to_normalize));
+    X = normalizedData(X(:,real_features_idx_to_normalize));
     % WITHOUT NORMALIZATION
[…]

### 2015-10-31 report/paper.tex
Finish report

--- 2015-10-31 report/paper.tex
@@
 It is also faster and a little bit more efficient than least squares using gradient descent and so, only results obtained with this latter will be reported.
-
+\vspace{-0.5em}
 \subsection{Evaluation methods}
@@
 Unfortunately, we didn't get major improvements.
-We think that the reason why dummy encoding doesn't help is because we don't have a high bias and so, having more features doesn't help.
+We think that the reason why dummy encoding didn't help is because we don't have a high bias and so, having more features doesn't help.
 
@@
 The eighth model is similar to the seventh one, except that this time it uses dummy encoding and has lower degrees : $3,5$ and $2$, due to the number of input variables.
-We can see that feature transformations significantly improve the test error compared to other models.
-However, dummy encoding doesn't seem to more efficient, and so our final model will be the seventh.
+We can see that feature transformations significantly improves the test error compared to other models.
+The reason is certainly because we had a high bias.
+However, dummy encoding doesn't seem to more efficient.
+Our final model will be the seventh with lambdas $\lambda_1 = 0.0369, \lambda_2 = 1.8303e^{-4}, \lambda_3 = 0.0053$.
 
@@
 
-We analyzed a regression data set with different methods and concluded that least squares wasn't a good approach due to the ill-conditioning and the least squares gradient descent was to slow and was less efficient than ridge regression.
+We analyzed a regression data set with different methods and concluded that least squares wasn't a good approach due to the ill-conditioning and the least squares gradient descent was too slow and less efficient than ridge regression.
 Separation of the data into clusters was essential in order to decrease the test error.
@@
 Finally, the variance of cluster 1 and 3 are very low, the one for cluster 2 is a bit higher.
-We think that we have bias because we use ridge regression but we don't think it is too high because we have seen that adding features helps and at a certain points it wasn't efficient.
-The best model we have is the seventh one, and we estimate the \textit{RMSE} of the prediction to be around $97$ with a standard deviation of $35$.
-Finally, we think that our model is improvable, however we couldn't try other things in order to improve the estimation error, due to a lack of time.
+We think that we have bias because we use ridge regression but we don't think it is too high because we have seen that adding features helps until a certain point.
+The best model we have is the seventh one and we estimate the \textit{RMSE} of the prediction to be around $115.96$ with a standard deviation of $45.59$.
+However, we should take into consideration outliers, which gives us a \textit{RMSE} of $407.55$ with a standard deviation of $229.34$.
+Finally, we think that our model is improvable, however we couldn't try other things to improve the estimation error, due to a lack of time.
 
@@
 
-Since our matrix is full-rank using logistic regression was appropriate.
+Since our matrix is full-rank, using logistic regression was appropriate.
 Because logistic regression is more efficient in term of complexity and that we don't need to penalize some betas, we report only the results using this method.
@@
 
-The fourth model consists in using dummy encoding for the categorical variable.
-This times, the error was improved and the variance has diminished.
+The fourth model consists in using dummy encoding for the categorical variables.
+This time the error was improved and the variance has diminished as well as certainly the bias due to the larger set of features.
 The next model (fifth) adds a polynomial basis of degree 2.
-We can observe the the variance as increased, we have some outliers (five in total, there might be hidden due to the legend) and the error has increased.
+We can observe the the variance as increased, we have some outliers (five in total, hidden by the legend) and the error has increased.
 We think that the bias is better but due to its high variance, we consider it as a bad model.
@@
 
-The two last models (eighth and ninth) use another transformation where negative values weren't transformed into positive one, holding them with the formula $-\sqrt[4]{-x}$ and try using also a polynomial basis.
+The two last models (eighth and ninth) use another transformation where negative values weren't transformed into positive one with the formula $-\sqrt[4]{-x}$ and try using also a polynomial basis.
 However, the error wasn't as good as the seventh model, surely because the transformed feature look like less Gaussian than the transformation using the absolute values.
 We can also see that the variance seems also higher.
-Our final model will be the seventh.
+Our final model will be the seventh with a learning step $\alpha=0.001$.
 
@@
 In hope to minimize the error, we tried different feature transformations.
-The first try was add polynomial of them self as $X = [X X^2]$.
+We first added polynomial of them self as $X = [X X^2]$.
 With the data not transformed into Gaussian like distribution, we didn't get an improvement (even with degree 3).
-We then try to transform the features which look like a Poisson distribution in order to transform them into a Gaussian-like distribution.
+We then tried to transform the features which look like a Poisson distribution in order to transform them into a Gaussian-like distribution.
 We first use the formula $\sqrt[4]{|x|}$ which gave us good results even if we transformed the negative values into positive.
 Mixed with polynomial of features at degree 2 allowed us to have an important improvement for the error.
-We also try to keep the negative values, using the formula $-\sqrt[4]{-x}$ but unfortunately, it didn't help.
+We also try to keep the negative values using the formula $-\sqrt[4]{-x}$ but unfortunately, it didn't help.
 We think the reason is first that distribution was more a superposition of two Gaussian.
@@
 The key of our model was to transform features which look like Poisson distribution into a Gaussian like distribution using $\sqrt[4]{|x|}$ and using a polynomial basis of degree 2.
-We think that our model is improvable due to outliers because we couldn't find such outliers in all dimensions (or most of the dimensions).
-We estimate the mean and standard deviation of the prediction, using the \textit{RMSE},\textit{0-1 loss} and \textit{logarithmic loss} respectively to be around $0.329$, $0.109$ and $0.509$ for the mean, and $0.028$, $0.018$, $0.018$for the standard deviation .
+We think that our model is improvable due to outliers because we couldn't find such in most of the dimensions.
+We estimate the mean and standard deviation of the prediction, using the \textit{RMSE},\textit{0-1 loss} and \textit{logarithmic loss} respectively to be around $0.329$, $0.109$ and $0.509$ for the mean, and $0.028$, $0.018$, $0.018$ for the standard deviation .
 

### 2015-11-01 report/paper.tex
Add Submission

@@
 We think that we have bias because we use ridge regression but we don't think it is too high because we have seen that adding features helps until a certain point.
-The best model we have is the seventh one and we estimate the \textit{RMSE} of the prediction to be around $115.96$ with a standard deviation of $45.59$.
-However, we should take into consideration outliers, which gives us a \textit{RMSE} of $407.55$ with a standard deviation of $229.34$.
+The best model we have is the seventh one and we estimate the \textit{RMSE} of the prediction to be around $115.34$ with a standard deviation of $454.61$.
+However, we should take into consideration outliers, which gives us a \textit{RMSE} of $412.52$ with a standard deviation of $234.78$.
 Finally, we think that our model is improvable, however we couldn't try other things to improve the estimation error, due to a lack of time.

### 2015-11-01 Submission/leastSquaresGD.m

@@ -1,7 +1,7 @@
 % Written by Diego Antognini & Jason Racine, EPFL 2015
 % all rights reserved
 
-function [beta] = penLogisticRegression(y, tX, alpha, lambda)
+function [beta] = leastSquaresGD(y, tX, alpha)
     % algorithm parameters
     maxIters = 35000;
     epsilon_convergence = 1e-6;
@@ -13,7 +13,7 @@ function [beta] = penLogisticRegression(y, tX, alpha, lambda)
     
     for k = 1:maxIters
         % Compute gradient
-        g = computeGradient(y, tX, beta, lambda);
+        g = computeGradient(y, tX, beta);
         
         % Cost function
         % L(k) = computeCost(y, tX, beta);
@@ -29,21 +29,14 @@ function [beta] = penLogisticRegression(y, tX, alpha, lambda)
     end
 end
 
-function [gradient] = computeGradient(y, tX, beta, lambda)
-    penalized_term = 2*lambda*beta;
-    
-    % We don't penalized beta zero
-    penalized_term(1,1) = 0;
-    
-    gradient = tX'*(sigmoid(tX*beta) - y) + penalized_term;
+function [gradient] = computeGradient(y, tX, beta)
+    N = length(y);
+    e = y - tX*beta;
+    gradient = -(1.0/N)*tX'*e;
 end
 
-function [res] = sigmoid(x)
-    ex = exp(x);
-    res = ex./(1 + ex);
-end
-
-function [res] = computeCost(y, tX, beta)
-    tXBeta = tX*beta;
-    res = -sum(y.*tXBeta - log(1 + exp(tXBeta)));
+function [e] = computeCost(y, tX, beta)
+    N = length(y);
+    e = y - tX*beta;
+    e = e'*e/(2.0*N);
 end

### 2015-11-01 Submission/normalizedData.m

@@ -1,0 +1,8 @@
+% Written by Diego Antognini & Jason Racine, EPFL 2015
+% all rights reserved
+
+function [XNorm] = normalizedData(X)
+    XNorm = X-repmat(mean(X), size(X, 1), 1);
+    XNorm = XNorm./repmat(std(XNorm), size(X, 1), 1);
+end
+

### 2015-11-01 Submission/penLogisticRegression.m

@@ -38,11 +38,6 @@ function [gradient] = computeGradient(y, tX, beta, lambda)
     gradient = tX'*(sigmoid(tX*beta) - y) + penalized_term;
 end
 
-function [res] = sigmoid(x)
-    ex = exp(x);
-    res = ex./(1 + ex);
-end
-
 function [res] = computeCost(y, tX, beta)
     tXBeta = tX*beta;
     res = -sum(y.*tXBeta - log(1 + exp(tXBeta)));

### 2015-11-01 classification/prediction.m

@@ -1,6 +1,12 @@
 % Written by Diego Antognini & Jason Racine, EPFL 2015
 % all rights reserved
 
+% We MUST normalize and transform -1 to 0 in order to have GD working
+% Add categorical features from preprocess
+% Add dummy encoding
+% Add polynomial basis
+% Add Transformation
+
 clear all;
 close all;
 
@@ -25,7 +31,8 @@ yRandom = y(shuffle,:);
 for k = 1:1:K;
     [XTr, yTr, XTe, yTe] = splitGetTrTe(yRandom, XRandom, idxCV, k);
 
-    [yTr, XTr] = preprocess(yTr, XTr, degree);    
+    [yTr, XTr] = preprocess(yTr, XTr, degree);   
+    
     beta = logisticRegression(yTr, [ones(length(yTr), 1) XTr], alpha);
     
     [yTe, XTe] = preprocess(yTe, XTe, degree);
@@ -48,7 +55,8 @@ std_err_log = std(err_log);
 
 %% TRAINING FOR PREDICTION
 
-[yTr, XTr] = preprocess(y, X, degree);    
+[yTr, XTr] = preprocess(y, X, degree);   
+
 beta = logisticRegression(yTr, [ones(length(yTr), 1) XTr], alpha);
 
 % PREPARE THE DATA TO PREDICT
@@ -58,5 +66,5 @@ y = zeros(size(X, 1), 1);
 [y, X_] = preprocess(y, X, degree);
 tXTe = [ones(length(y),1) X_];
 y = (sigmoid(tXTe*beta) >= 0.5).*1.0;
-y(find(y == 0)) = -1;
+
 csvwrite('predictions_classification.csv', y);

### 2015-11-01 regression/prediction.m

@@ -1,8 +1,11 @@
 % Written by Diego Antognini & Jason Racine, EPFL 2015
 % all rights reserved
 
-%clear all;
-%close all;
+% Add in preprocessCluster around line 68 the normalization 
+% Remove in preprocessCluster around line 75 the dummy encoding
+
+clear all;
+close all;
 
 load('Sydney_regression.mat');
 
@@ -18,6 +21,10 @@ d1 = 3;
 d2 = 6;
 d3 = 3;
 
+lambda_cls1 = 0.0369;
+lambda_cls2 = 1.8303e-4;
+lambda_cls3 = 0.0053;
+    
 %% COMPUTING ESTIMATE ERROR
 K = 10;
 N = length(y);
@@ -29,10 +36,6 @@ yRandom = y(shuffle,:);
 
 for k = 1:1:K;
     [XTr, yTr, XTe, yTe] = splitGetTrTe(yRandom, XRandom, idxCV, k);
-    
-    lambda_cls1 = 0.0369;
-    lambda_cls2 = 1.8303e-4;
-    lambda_cls3 = 0.0053;
 
     [y_cls1, X_cls1, y_cls2, X_cls2, y_cls3, X_cls3, idx_cls1, idx_cls2, idx_cls3] = preprocess(yTr, XTr, d1, d2, d3);
 
@@ -50,24 +53,18 @@ for k = 1:1:K;
     tXTe_cls2 = [ones(length(y_cls2),1) X_cls2];
     tXTe_cls3 = [ones(length(y_cls3),1) X_cls3];
 
-    err_cls1(k) = RMSE(y_cls1, tXTe_cls1*beta_cls1);
-    err_cls2(k) = RMSE(y_cls2, tXTe_cls2*beta_cls2);
-    err_cls3(k) = RMSE(y_cls3, tXTe_cls3*beta_cls3);
+    N_ = length(y_cls1)+length(y_cls2)+length(y_cls3);
+    err(k) = sqrt((1/N_) * (sum((y_cls1 - tXTe_cls1*beta_cls1).^2) ...
+                         +  sum((y_cls2 - tXTe_cls2*beta_cls2).^2) ...
+                         +  sum((y_cls3 - tXTe_cls3*beta_cls3).^2)));
 end
 
[…]

### 2015-11-02 report/paper.tex
Add final submission

@@
 \subfigure[Feature 2 can help us to separate the data.
-Green data points are misclassified data.
+Light blue data points are misclassified data.
 The separation is at $x=0.42$.]{\includegraphics[width=2.7in]{figures/feature2.jpg}\label{fig:feature2}}
@@
 \subfigure[Feature 16 can help us to separate the data.
-Green data points are misclassified data.
+Light blue data points are misclassified data.
 The separation is at $x=1.17$.]{\includegraphics[width=2.7in]{figures/feature16.jpg}\label{fig:feature16}}
@@
 
-We can note that the rank of our input matrix $\mathbf{X}$ is rank-deficient with a rank of 66 instead of 76, and is rank-deficient of 65 for each cluster.
+We can note that our input matrix $\mathbf{X}$ is rank-deficient with a rank of 66 instead of 76, and is rank-deficient of 65 for each cluster.
 \vspace{-0.5em}
@@
 The next model (fifth) is about doing a ridge regression per cluster, each one having its own lambda.
-We also use $k$ fold cross validation where the first cluster has a $k_1=10$ because we have a lot of data ($1937$), the second $k_2=6$, because we don't have a lot of data ($563$) and the last $k_3=4$ because we really a small number of data ($273$) and so we have to choose a smaller k.
+We also use $k$ fold cross validation where the first cluster has a $k_1=10$ because we have a lot of data ($1937$), the second $k_2=6$, because we don't have a lot of data ($563$) and the last $k_3=4$ because we really have a small number of data ($273$) and so we have to choose a smaller k.
 As expected we obtained another big improvement justifying again the choice of splitting the data.
@@
 
-The two next models is about using polynomial basis functions for each cluster.
+The two next models are about using polynomial basis functions for each cluster.
 Categorical variables aren't concerned, they are just replicated for each degree, because it doesn't make any sense to build polynomial for them.
@@
 We think that we have bias because we use ridge regression but we don't think it is too high because we have seen that adding features helps until a certain point.
-The best model we have is the seventh one and we estimate the \textit{RMSE} of the prediction to be around $115.34$ with a standard deviation of $454.61$.
+The best model we have is the seventh one and we estimate the \textit{RMSE} of the prediction to be around $115.34$ with a standard deviation of $44.62$.
 However, we should take into consideration outliers, which gives us a \textit{RMSE} of $412.52$ with a standard deviation of $234.78$.

### 2015-11-02 Submission/classification/RMSE.m

@@ -1,8 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-function [e] = RMSE(y, p)
-    N = length(y);
-    e = sqrt(sum((y-p).^2)/N);
-end
-

### 2015-11-02 Submission/classification/exploratoryDataAnalysis.m

@@ -1,169 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-clear all;
-close all;
-load('Sydney_classification.mat');
-
-y = y_train;
-X = X_train;
-N = length(y);
-
-%% Plot of output
-f1 = figure;
-boxplot(X,'notch','on');
-xlabel('Features');
-ylabel('Output');
-% saveas(f1, 'plots/general/boxplot_unnormalized.jpg');
-
-%% Plot of normalized output
-f2 = figure;
-boxplot(normalizedData(X),'notch','on');
-xlabel('Features');
-ylabel('Output');
-% saveas(f2, 'plots/general/boxplot_normalized.jpg');
-
-%% Histogram of output
-f3 = figure;
-histogram(y);
-xlabel('Y');
-ylabel('Count');
-xlim([-1 1]);
-% saveas(f3, 'plots/general/histogram.jpg');
-
-%% Plot feature VS Output
-for feature=1:size(X,2)
-    f = figure;
-    plot(X(:,feature),y,'ob');
-    ylabel('y');
-    xlabel(sprintf('Values of feature %d', feature));
-    % saveas(f, sprintf('plots/features/featuresVSOutput/feature%d.jpg', feature));
-end
-
-idx = find(y == 1);
-for feature=1:size(X,2)
-    f = figure;
-    plot(X(idx,feature),y(idx),'ob');
-    ylabel('y');
-    xlabel(sprintf('Values of feature %d', feature));
-    % saveas(f, sprintf('plots/features/featuresVSOuputTo1/feature%d.jpg', feature));
-end
-
-idx = find(y == -1);
-for feature=1:size(X,2)
-    f = figure;
-    plot(X(idx,feature),y(idx),'ob');
-    ylabel('y');
-    xlabel(sprintf('Values of feature %d', feature));
-    % saveas(f, sprintf('plots/features/featuresVSOuputTo0/feature%d.jpg', feature));
-end
-
-%% Plot histogram of each feature
-for feature=1:size(X,2)
-    f = figure;
-    histogram(X(:,feature));
-    xlabel(sprintf('Feature %d', feature));
-    ylabel('Count');
-    % saveas(f, sprintf('plots/features/featuresHistogram/feature%d.jpg', feature));
-end
-
-idx = find(y == 1);
-for feature=1:size(X,2)
-    f = figure;
[…]

### 2015-11-02 Submission/classification/logLoss.m

@@ -1,14 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-function [e] = logLoss(y, p)
-    N = length(y);
-    res = 0;
-    
-    for i = 1:1:N
-       res = res + y(i)*p(i) - log(1.0 + exp(p(i)));
-    end
-    
-    e = -res/N;
-end
-

### 2015-11-02 Submission/classification/main.m

@@ -1,619 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-clear all;
-close all;
-
-load('Sydney_classification.mat');
-
-y = y_train;
-X = X_train;
-N = length(y);
-
-% 1) Constant baseline
-% 2) Logistic regression (normalization & output [0,1] & only real features)
-% 3) + categorical variables
-% 4) + dummy encoding
-% 5) + Polynomial basis 2
-% 6) Feature transformation (abs) (degree 1)
-% 7) Feature transformation (abs) (degree 2)
-% 8) Feature transformation (with neg) (degree 1)
-% 9) Feature transformation (with neg) (degree 2)
-
-numberOfExperiments = 30;
-proportionOfTraining = 0.8;
-K = 10;
-
-%%
-%**********************************
-%            TEST 1
-%**********************************
-
-for i = 1:numberOfExperiments
-    setSeed(28111993*i);
-    [XTr, yTr, XTe, yTe] = splitProp(proportionOfTraining, y, X);
-     
-    out = 1;
-    
-    err1(i) = RMSE(yTe, out);
-    err1_01(i) = zeroOneLoss(yTe, repmat([out], length(yTr), 1));
-    err1_log(i) = logLoss(yTe, repmat([out], length(yTr), 1));
-end
-
-saveFile(err1, 'results/err1');
-saveFile(err1_01, 'results/err1_01');
-saveFile(err1_log, 'results/err1_log');
-
-%%
-% **********************************
-%             TEST 2
-% **********************************
-% We MUST normalize and transform -1 to 0 in order to have GD working
-% Remove categorical features from preprocess
-
-alphas = [0.00075];
-for i = 1:1:numberOfExperiments
-    setSeed(28111993*i);
-    [XTr, yTr, XTe, yTe] = splitProp(proportionOfTraining, y, X);
-     
-    [yTr, XTr] = preprocess(yTr, XTr);
-%     N = length(yTr);
-%     idxCV = splitGetCV(K, N);
-%     
-%     for a = 1:1:length(alphas)
-%         % K-fold
-%         for k=1:1:K
-%             [XXTr, yyTr, XXTe, yyTe] = splitGetTrTe(yTr, XTr, idxCV, k);
-%             
[…]

### 2015-11-02 Submission/classification/openFile.m

@@ -1,9 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-function [matrix] = openFile(name, size)
-    fileID = fopen(sprintf('%s.bin',name));
-    matrix = fread(fileID, size, 'double');
-    fclose(fileID);
-end
-

### 2015-11-02 Submission/classification/penLogisticRegression.m

@@ -1,49 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-function [beta] = penLogisticRegression(y, tX, alpha, lambda)
-    % algorithm parameters
-    maxIters = 100000;
-    epsilon_convergence = 1e-6;
-
-    % initialize beta
-    D = size(tX,2)-1;
-    beta = zeros(D+1,1);
-    %steps = maxIters;
-    
-    for k = 1:maxIters
-        % Compute gradient
-        g = computeGradient(y, tX, beta, lambda);
-        
-        % Cost function
-        % L(k) = computeCost(y, tX, beta);
-        
-        % Update beta
-        beta = beta - alpha*g;
-        
-        % Check convergence
-        if(norm(g) < epsilon_convergence)
-            %steps = k;
-            break;
-        end
-    end
-end
-
-function [gradient] = computeGradient(y, tX, beta, lambda)
-    penalized_term = 2*lambda*beta;
-    
-    % We don't penalized beta zero
-    penalized_term(1,1) = 0;
-    
-    gradient = tX'*(sigmoid(tX*beta) - y) + penalized_term;
-end
-
-function [res] = sigmoid(x)
-    ex = exp(x);
-    res = ex./(1 + ex);
-end
-
-function [res] = computeCost(y, tX, beta)
-    tXBeta = tX*beta;
-    res = -sum(y.*tXBeta - log(1 + exp(tXBeta)));
-end

### 2015-11-02 Submission/classification/poissonToGaussian.m

@@ -1,17 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-function [X] = poissonToGaussian(X, features)
-    for f = 1:1:length(features)
-        feature = features(f);
-        
-        % Using Abs
-        X(:,feature) = abs(X(:,feature)).^0.25;
-        
-        % Holding negative values
-        %neg_idx = find(X(:,feature) < 0);
-        %X(neg_idx, feature) = -X(neg_idx, feature);
-        %X(:,feature) = X(:,feature).^0.25;
-        %X(neg_idx, feature) = -X(neg_idx, feature);
-    end
-end

### 2015-11-02 Submission/classification/prediction.m

@@ -1,70 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-% We MUST normalize and transform -1 to 0 in order to have GD working
-% Add categorical features from preprocess
-% Add dummy encoding
-% Add polynomial basis
-% Add Transformation
-
-clear all;
-close all;
-
-load('Sydney_classification.mat');
-
-y = y_train;
-X = X_train;
-N = length(y);
-
-K = 10;
-degree = 2;
-alpha = 0.001;
-
-%% COMPUTING ESTIMATE ERROR
-
-idxCV = splitGetCV(K, N);
-setSeed(28111993);
-shuffle = randperm(N);
-XRandom = X(shuffle,:);
-yRandom = y(shuffle,:);
-
-for k = 1:1:K;
-    [XTr, yTr, XTe, yTe] = splitGetTrTe(yRandom, XRandom, idxCV, k);
-
-    [yTr, XTr] = preprocess(yTr, XTr, degree);   
-    
-    beta = logisticRegression(yTr, [ones(length(yTr), 1) XTr], alpha);
-    
-    [yTe, XTe] = preprocess(yTe, XTe, degree);
-    
-    tXTe = [ones(length(yTe),1) XTe];
-    
-    y_hat = (sigmoid(tXTe*beta) >= 0.5).*1.0;
-    
-    err(k) = RMSE(yTe, y_hat);
-    err_01(k) = zeroOneLoss(yTe, y_hat);
-    err_log(k) = logLoss(yTe, y_hat);
-end
-
-mean_err = mean(err);
-std_err = std(err);
-mean_err_01 = mean(err_01);
-std_err_01 = std(err_01);
-mean_err_log = mean(err_log);
-std_err_log = std(err_log);
-
-%% TRAINING FOR PREDICTION
-
-[yTr, XTr] = preprocess(y, X, degree);   
-
-beta = logisticRegression(yTr, [ones(length(yTr), 1) XTr], alpha);
-
-% PREPARE THE DATA TO PREDICT
-X = X_test;
-y = zeros(size(X, 1), 1);
-
-[y, X_] = preprocess(y, X, degree);
-tXTe = [ones(length(y),1) X_];
-y = (sigmoid(tXTe*beta) >= 0.5).*1.0;
-
-csvwrite('predictions_classification.csv', y);

### 2015-11-02 Submission/classification/preprocess.m

@@ -1,142 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-% If degrees are not specified, they will get the value 1.
-function [y, X] = preprocess(y_, X_, degree)
-     if ~exist('degree', 'var')
-            degree = 1;
-     end
-    
-    % Find different type of features
-    [idx_feature_2, idx_feature_4, idx_feature_5] = findDiscreteFeatures(X_);
-    idx_feature_real = setdiff(setdiff(setdiff(1:1:size(X_,2), idx_feature_2), idx_feature_4), idx_feature_5);
-
-    X = poissonToGaussian(X_, idx_feature_real);
-    y = y_;
-    
-    % Transform -1 to 0
-    y(find(y == -1)) = 0;
-
-    % Build the poly for X
-    X = repmat(X,1,degree);
-    nbFeatures = size(X_,2);
-
-    % Prepare to extend idx_feature_i for the polynomial features
-     idx_feature_2_poly = idx_feature_2;
-     idx_feature_4_poly = idx_feature_4;
-     idx_feature_5_poly = idx_feature_5;
-     
-     real_features_idx_to_normalize = idx_feature_real;
-    for d = 1:1:(degree-1)
-        % Limit of the i'th degree
-        beg = d*nbFeatures+1;
-        en = (d+1)*nbFeatures;
-        
-        % Compute the new idx of features of the current degree
-        real_features_idx = idx_feature_real + (size(X_,2)*d);
-        idx_feature_2_this_degree = idx_feature_2 + (size(X_,2)*d);
-        idx_feature_4_this_degree = idx_feature_4 + (size(X_,2)*d);
-        idx_feature_5_this_degree = idx_feature_5 + (size(X_,2)*d);
-        
-        % Do the union between old idx and the new one
-        real_features_idx_to_normalize = union(real_features_idx_to_normalize, real_features_idx);
-        idx_feature_2_poly = union(idx_feature_2_poly, idx_feature_2_this_degree);
-        idx_feature_4_poly = union(idx_feature_4_poly, idx_feature_4_this_degree);
-        idx_feature_5_poly = union(idx_feature_5_poly, idx_feature_5_this_degree);
-    
-        % Update only the real features, because we don't raise discrete
-        % variable to power
-        cols = intersect(real_features_idx, beg:1:en);
-        X(:,cols) = X(:,cols).^(d+1);
-    end
-    
[…]

### 2015-11-02 Submission/classification/saveFile.m

@@ -1,9 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-function [] = saveFile(matrix, name)
-    fileID = fopen(sprintf('%s.bin',name), 'w');
-    fwrite(fileID, matrix, 'double');
-    fclose(fileID);
-end
-

### 2015-11-02 Submission/classification/setSeed.m

@@ -1,12 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-function setSeed(seed)
-    % set seed
-	global RNDN_STATE  RND_STATE
-	RNDN_STATE = randn('state');
-	randn('state',seed);
-	RND_STATE = rand('state');
-	%rand('state',seed);
-	rand('twister',seed);
-end

### 2015-11-02 Submission/classification/splitGetTrTe.m

@@ -1,14 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-function [XTr, yTr, XTe, yTe] = splitGetTrTe(y, X, idxCV, k)
-    % get k'th subgroup in test, others in train
-    idxTe = idxCV(k,:);
-    idxTr = idxCV([1:k-1 k+1:end],:);
-    idxTr = idxTr(:);
-    
-    yTe = y(idxTe);
-    XTe = X(idxTe,:);
-    yTr = y(idxTr);
-    XTr = X(idxTr,:);
-end

### 2015-11-02 Submission/classification/zeroOneLoss.m

@@ -1,16 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-function [e] = zeroOneLoss(y, yhat)
-    N = length(y);
-    res = 0;
-    
-    for i = 1:1:N
-       if y(i) ~= yhat(i)
-           res = res + 1;
-       end
-    end
-    
-    e = res/N;
-end
-

### 2015-11-02 Submission/leastSquaresGD.m

@@ -4,7 +4,7 @@
 function [beta] = leastSquaresGD(y, tX, alpha)
     % algorithm parameters
     maxIters = 35000;
-    epsilon_convergence = 1e-6;
+    epsilon_convergence = 1e-8;
 
     % initialize beta
     D = size(tX,2)-1;

### 2015-11-02 Submission/penLogisticRegression.m

@@ -4,7 +4,7 @@
 function [beta] = penLogisticRegression(y, tX, alpha, lambda)
     % algorithm parameters
     maxIters = 35000;
-    epsilon_convergence = 1e-6;
+    epsilon_convergence = 1e-8;
 
     % initialize beta
     D = size(tX,2)-1;
@@ -41,4 +41,9 @@ end
 function [res] = computeCost(y, tX, beta)
     tXBeta = tX*beta;
     res = -sum(y.*tXBeta - log(1 + exp(tXBeta)));
+end
+
+function [res] = sigmoid(x)
+    ex = exp(x);
+    res = ex./(1 + ex);
 end

### 2015-11-02 Submission/regression/RMSE.m

@@ -1,8 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-function [e] = RMSE(y, yhat)
-    N = length(y);
-    e = sqrt(sum((y-yhat).^2)/N);
-end
-

### 2015-11-02 Submission/regression/exploratoryDataAnalysis.m

@@ -1,270 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-clear all;
-close all;
-load('Sydney_regression.mat');
-
-y = y_train;
-X = X_train;
-N = length(y);
-
-%% Plot of output
-f1 = figure;
-boxplot(X,'notch','on');
-xlabel('Features');
-ylabel('Output');
-%saveas(f1, 'plots/general/boxplot_unnormalized.jpg');
-
-%% Plot of normalized output
-f2 = figure;
-boxplot(normalizedData(X),'notch','on');
-xlabel('Features');
-ylabel('Output');
-%saveas(f2, 'plots/general/boxplot_normalized.jpg');
-
-%% Histogram of output
-f3 = figure;
-edges = 0:50:20000;
-histogram(y,edges);
-xlabel('Y');
-ylabel('Count');
-xlim([0 15000]);
-set(gca, 'XGrid','on')
-set(gca, 'YGrid','on')
-set(gca,'fontsize', 18);
-set(gca,'LineWidth',2);
-% saveas(f3, 'plots/general/histogram.jpg');
-print('../report/figures/histogram','-djpeg','-noui')
-
-%% Plot feature VS Output
-for feature=1:size(X,2)
-    f = figure;
-    plot(X(:,feature),y,'ob');
-    ylabel('y');
-    xlabel(sprintf('Values of feature %d', feature));
-    % saveas(f, sprintf('plots/features/featuresVSOutput/feature%d.jpg', feature));
-end
-
-%% Plot feature 2 VS Output and feature 16 VS Output
-x_features = [2 16];
-for x_f_idx=1:length(x_features)
-    x_f = x_features(x_f_idx);
-    for feature=1:size(X,2)
-        % We don't plot feature 2 and 16 VS Output
-        if(feature==x_f)
-            continue;
-        end
-        
-        f = figure;
-        plot(X(:,x_f),X(:,feature),'ob');
-        ylabel(sprintf('Values of feature %d', feature));
-        xlabel(sprintf('Values of feature %d', x_f));
-        % saveas(f, sprintf('plots/features/feature%dVSAll/feature%d.jpg', x_f, feature));
-    end
-end
-
-%% Plot histogram of each feature
-for feature=1:size(X,2)
[…]

### 2015-11-02 Submission/regression/findClusters.m

@@ -1,34 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-function [idx_cls1, idx_cls2, idx_cls3] = findClusters(y, X)
-    XN = normalizedData(X);    
-    yN = normalizedData(y);
-
-    % Find first separation of cluster using feature 2
-    idx_cluster1 = find(XN(:,2) < 0.42);
-    idx_others_clusters = setdiff(1:length(y), idx_cluster1)';
-    
-    % There are 12 outliers here !
-    outliers_cluster1 = find(yN > 3.5 | (XN(:,2) < 0.42 & yN > 0.4) | (XN(:,2) > 0.42 & yN < 0));
-
-    % Find second separation of cluster using feature 16
-    idx_cluster3 = setdiff(idx_others_clusters, find(yN < 1.65 | XN(:,16) < 0.7));
-
-    % There is 11 outlier here
-    outliers_cluster3 = find(yN > 2 & XN(:,16) < 1.17);
-
-    % Find the last cluster using the last points
-    idx_cluster2 = setdiff(idx_others_clusters, idx_cluster3);
-
-    % There are 5 outliers here !
-    outliers_cluster2 = find(yN > 3.5 | (XN(:,16) > 1.17 & yN > 0.5 & yN < 1.5));
-
-    % Some checks
-    assert(length(idx_cluster1) + length(idx_cluster2) + length(idx_cluster3) == length(X));
-    
-    % Write cluster idx without outliers
-    idx_cls1 = setdiff(setdiff(setdiff(idx_cluster1, outliers_cluster1), outliers_cluster2), outliers_cluster3);
-    idx_cls2 = setdiff(setdiff(setdiff(idx_cluster2, outliers_cluster1), outliers_cluster2), outliers_cluster3);
-    idx_cls3 = setdiff(setdiff(setdiff(idx_cluster3, outliers_cluster1), outliers_cluster2), outliers_cluster3);
-end

### 2015-11-02 Submission/regression/findCorrespondingCluster.m

@@ -1,13 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-% x is a data vector. Both have to be normalized
-function [idx_cls] = findCorrespondingCluster(x)
-    if x(2) < 0.42
-       idx_cls = 1;
-    elseif x(16) < 1.17
-       idx_cls = 2;
-    else
-       idx_cls = 3;
-    end
-end

### 2015-11-02 Submission/regression/findLambda.m

@@ -1,68 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-%idxCls1 is to indicate which cluster do we use (1 2 3). 0 means all the
-%dataset
-function [lambda] = findLambda(K, y_, X_, idxCls, d1, d2, d3)
-    % If we choose a cluster, we check if all the degrees were specified
-    % Otherwise we choose arbitrary degree to 1 for each cluster
-    if idxCls ~= 0
-        if ~exist('d1', 'var') || ~exist('d2', 'var') || ~exist('d3', 'var')
-            [y_cls1, X_cls1, y_cls2, X_cls2, y_cls3, X_cls3] = preprocess(y_, X_);  
-        else
-            [y_cls1, X_cls1, y_cls2, X_cls2, y_cls3, X_cls3] = preprocess(y_, X_, d1, d2, d3);
-        end
-    end
-    
-    % Select cluster if specified or the whole dataset
-    if idxCls == 1
-        y = y_cls1;
-        X = X_cls1;
-    elseif idxCls == 2
-        y = y_cls2;
-        X = X_cls2;
-    elseif idxCls == 3
-        y = y_cls3;
-        X = X_cls3;
-    else
-        y = y_;
-        X = X_;
-    end
-
-    N = length(y);
-    idxCV = splitGetCV(K, N);
-
-    % K-fold
-    for k=1:1:K
-        [XTr, yTr, XTe, yTe] = splitGetTrTe(y, X, idxCV, k);
-
-        tXTr = [ones(length(yTr),1) XTr];
-        tXTe = [ones(length(yTe),1) XTe];
-
-        lambdas = logspace(-5,5,500);
-        for i = 1:1:length(lambdas)
-            lambda = lambdas(i);
-            beta_rr = ridgeRegression(yTr, tXTr, lambda);
-            err_tr_rr(k,i) = RMSE(yTr, tXTr*beta_rr);
-            err_te_rr(k,i) = RMSE(yTe, tXTe*beta_rr);
-        end
-    end
-
-    mseTr = mean(err_tr_rr);
-    mseTe = mean(err_te_rr);
-
-    [errStar, lambdaIdStar] = min(mseTe);
-    lambdaStar = lambdas(lambdaIdStar);
-    
-%     figure
-%     semilogx(lambdas, mseTr, 'b-o', lambdas, mseTe, 'r-x');
-%     hold on;
[…]

### 2015-11-02 Submission/regression/leastSquares.m

@@ -1,7 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-function [beta] = leastSquares(y, tX)
-    beta = (tX'*tX) \ (tX'*y);
-end
-

### 2015-11-02 Submission/regression/leastSquaresGD.m

@@ -1,42 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-function [beta] = leastSquaresGD(y, tX, alpha)
-    % algorithm parameters
-    maxIters = 35000;
-    epsilon_convergence = 1e-6;
-
-    % initialize beta
-    D = size(tX,2)-1;
-    beta = zeros(D+1,1);
-    %steps = maxIters;
-    
-    for k = 1:maxIters
-        % Compute gradient
-        g = computeGradient(y, tX, beta);
-        
-        % Cost function
-        % L(k) = computeCost(y, tX, beta);
-        
-        % Update beta
-        beta = beta - alpha*g;
-        
-        % Check convergence
-        if(norm(g) < epsilon_convergence)
-            %steps = k;
-            break;
-        end
-    end
-end
-
-function [gradient] = computeGradient(y, tX, beta)
-    N = length(y);
-    e = y - tX*beta;
-    gradient = -(1.0/N)*tX'*e;
-end
-
-function [e] = computeCost(y, tX, beta)
-    N = length(y);
-    e = y - tX*beta;
-    e = e'*e/(2.0*N);
-end

### 2015-11-02 Submission/regression/normalizedData.m

@@ -1,8 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-function [XNorm] = normalizedData(X)
-    XNorm = X-repmat(mean(X), size(X, 1), 1);
-    XNorm = XNorm./repmat(std(XNorm), size(X, 1), 1);
-end
-

### 2015-11-02 Submission/regression/prediction.m

@@ -1,112 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-% Add in preprocessCluster around line 68 the normalization 
-% Remove in preprocessCluster around line 75 the dummy encoding
-
-clear all;
-close all;
-
-load('Sydney_regression.mat');
-
-y = y_train;
-X = X_train;
-N = length(y);
-
-k1 = 10;
-k2 = 8;
-k3 = 6;
-
-d1 = 3;
-d2 = 6;
-d3 = 3;
-
-lambda_cls1 = 0.0369;
-lambda_cls2 = 1.8303e-4;
-lambda_cls3 = 0.0053;
-    
-%% COMPUTING ESTIMATE ERROR
-K = 10;
-N = length(y);
-idxCV = splitGetCV(K, N);
-setSeed(28111993);
-shuffle = randperm(N);
-XRandom = X(shuffle,:);
-yRandom = y(shuffle,:);
-
-for k = 1:1:K;
-    [XTr, yTr, XTe, yTe] = splitGetTrTe(yRandom, XRandom, idxCV, k);
-
-    [y_cls1, X_cls1, y_cls2, X_cls2, y_cls3, X_cls3, idx_cls1, idx_cls2, idx_cls3] = preprocess(yTr, XTr, d1, d2, d3);
-
-    tX_cls1 = [ones(length(y_cls1),1) X_cls1];
-    tX_cls2 = [ones(length(y_cls2),1) X_cls2];
-    tX_cls3 = [ones(length(y_cls3),1) X_cls3];
-
-    beta_cls1 = ridgeRegression(y_cls1, tX_cls1, lambda_cls1);
-    beta_cls2 = ridgeRegression(y_cls2, tX_cls2, lambda_cls2);
-    beta_cls3 = ridgeRegression(y_cls3, tX_cls3, lambda_cls3);
-
-    [y_cls1, X_cls1, y_cls2, X_cls2, y_cls3, X_cls3, idx_cls1, idx_cls2, idx_cls3] = preprocess(yTe, XTe, d1, d2, d3, 1);
-
-    tXTe_cls1 = [ones(length(y_cls1),1) X_cls1];
-    tXTe_cls2 = [ones(length(y_cls2),1) X_cls2];
-    tXTe_cls3 = [ones(length(y_cls3),1) X_cls3];
-
-    N_ = length(y_cls1)+length(y_cls2)+length(y_cls3);
[…]

### 2015-11-02 Submission/regression/preprocess.m

@@ -1,193 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-% If degrees are not specified, they will get the value 1.
-% is prediction separate the cluster using only the X values
-function [y_cls1, X_cls1, y_cls2, X_cls2, y_cls3, X_cls3, idx_cls1, idx_cls2, idx_cls3] = preprocess(y, X, degree1, degree2, degree3, is_prediction)
-    if ~exist('degree1', 'var')
-        degree1 = 1;
-    end
-    
-    if ~exist('degree2', 'var')
-        degree2 = 1;
-    end
-    
-    if ~exist('degree3', 'var')
-        degree3 = 1;
-    end
-    
-    % Find idx of each cluster
-    if ~exist('is_prediction', 'var')
-        [idx_cls1, idx_cls2, idx_cls3] = findClusters(y, X);
-    % Using only the X values (the case as X_test)
-    else
-        idx_cls1 = [];
-        idx_cls2 = [];
-        idx_cls3 = [];
-        X_ = normalizedData(X);
-        for n = 1:1:size(X_,1)
-           res = findCorrespondingCluster(X_(n,:));
-           if res == 1
-               idx_cls1 = [idx_cls1 n];
-           elseif res == 2
-               idx_cls2 = [idx_cls2 n];
-           elseif res == 3
-               idx_cls3 = [idx_cls3 n];
-           end
-        end
-        idx_cls1 = idx_cls1';
-        idx_cls2 = idx_cls2';
-        idx_cls3 = idx_cls3';
-    end
-
-    % Find different type of features
-    [idx_feature_2, idx_feature_3, idx_feature_4] = findDiscreteFeatures(X);
-    idx_feature_real = setdiff(setdiff(setdiff(1:1:size(X,2), idx_feature_2), idx_feature_3), idx_feature_4);
-
-    % Preprocess per cluster
-    [y_cls1, X_cls1] = preprocessCluster(y, X, idx_cls1, degree1, idx_feature_2, idx_feature_3, idx_feature_4, idx_feature_real);
-    [y_cls2, X_cls2] = preprocessCluster(y, X, idx_cls2, degree2, idx_feature_2, idx_feature_3, idx_feature_4, idx_feature_real);
-    [y_cls3, X_cls3] = preprocessCluster(y, X, idx_cls3, degree3, idx_feature_2, idx_feature_3, idx_feature_4, idx_feature_real);
-end
[…]

### 2015-11-02 Submission/regression/ridgeRegression.m

@@ -1,15 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-function [beta] = ridgeRegression(y, tX, lambda)
-    tXTmultX = tX'*tX;
-    
-    % tXTmultX is a square matrix, so no problem with length
-    I = eye(length(tXTmultX));   
-    
-    % We do not touch beta_0
-    I(1,1) = 0;
-    
-    beta = (tX'*tX + lambda*I)\tX'*y;
-end
-

### 2015-11-02 Submission/regression/splitProp.m

@@ -1,24 +1,0 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
-function [XTr, yTr, XTe, yTe] = splitProp(prop, y, X)
-    % split the data into train and test given a proportion
-    N = length(y);
-
-    shuffle = randperm(N);
-    Xrandom = X(shuffle,:);
-    yrandom = y(shuffle,:);
-    
-	% generate random indices
-	idx = randperm(N);
-    Ntr = floor(prop * N);
-	% select few as training and others as testing
-	idxTr = idx(1:Ntr);
-	idxTe = idx(Ntr+1:end);
-	% create train-test split
-    XTr = Xrandom(idxTr,:);
-    yTr = yrandom(idxTr);
-    XTe = Xrandom(idxTe,:);
-    yTe = yrandom(idxTe);
-end
-

### 2015-11-02 classification/logisticRegression.m

@@ -3,7 +3,7 @@
 
 function [beta] = logisticRegression(y, tX, alpha)
     % algorithm parameters
-    maxIters = 35000;
+    maxIters = 1000000;
     epsilon_convergence = 1e-8;
 
     % initialize beta
@@ -27,6 +27,7 @@ function [beta] = logisticRegression(y, tX, alpha)
             break;
         end
     end
+    fprintf('%.10f\n', norm(g));
 end
 
 function [gradient] = computeGradient(y, tX, beta)

### 2015-12-17 Projet II/report/paper.tex
Add PCA for hog & CNN

\documentclass{article} % For LaTeX2e
% We will use NIPS submission format
\usepackage{nips13submit_e,times}
% for hyperlinks
\usepackage{hyperref}
\usepackage{url}
% For figures
\usepackage{graphicx} 
\usepackage{subfigure} 
% math packages
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsopn}
\usepackage{ifthen}
\usepackage{natbib}

\title{Project-II by Group Sydney}

\author{
Diego Antognini \& Jason Racine\\
EPFL \\
\texttt{diego.antognini@epfl.ch}, \texttt{jason.racine@epfl.ch} \\
}

% The \author macro works with any number of authors. There are two commands
% used to separate the names and addresses of multiple authors: \And and \AND.
%
% Using \And between authors leaves it to \LaTeX{} to determine where to break
% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
% puts 3 of 4 authors names on the first line, and the last on the second
% line, try using \AND instead of \And before the third author name.

\nipsfinalcopy

\begin{document}

\maketitle

\begin{abstract}
This report provides a summary of the project two of the PCML class. 
\end{abstract}

\section{Binary Classification}

\subsection{Data Description}

The train-data for binary classification consists of $N = 6000$ images. As input we have 2 representations of the image : \textit{the histogram of oriented gradients} (HOG) $\mathbf{X}_{hog}$ and the \textit{overFEAT ImageNet CNN features} (CNN) $\mathbf{X}_{cnn}$. Our input $\mathbf{X}$ is the concatenation of $\mathbf{X}_{hog}$ with $\mathbf{X}_{cnn}$. 
\\Each input sample is a vector $\mathbf{x}_n$ with dimension $D = 42273$ ($5408$ for the HOG and $36865$ for the CNN) and is the concatenation of $\mathbf{x}_n^{hog}$ with $\mathbf{x}_n^{cnn}$
The output ($\mathbf{y}$) represents the classification of the images, with $0$ if it represents a car, a horse or a plane, $1$ otherwise.

We also have test-data of size $N=XXX$ without their corresponding output. Our goal is to produce predictions for those data, as well as an approximation of the test-error using \textit{Balanced Error Rate} (BER).

\subsubsection{Histogram of Oriented Gradients}

Histogram of oriented gradients is a feature used in computer vision in order to detect objects. To compute it, we first need to normalize the image, compute the gradients (of each pixel) for the different color channel and take those with the largest norm. Then we decompose the image in bins of size $w \times h$. For each of those bins, we compute an histogram of the orientation of the gradients using theirs angles and weighted by theirs magnitudes. The histogram has $n$ bins from $0$ to $180$ degrees. We compute this histogram using $4$ different normalizations.

In our case, the dimensions of this feature is $13 \times 13 \times 32=5408$, where the first $13$ is the number of bins in the height, the second $13$ the number of bins in the width. Finally $32$ is $4 \times 8$ where the $4$ is the different normalizations for the histogram and $8$ the number of bins (each bin has a size of $22.5°$).

%http://lear.inrialpes.fr/people/triggs/pubs/Dalal-cvpr05.pdf
%http://vision.ucsd.edu/~pdollar/toolbox/doc/channels/hog.html
\subsubsection{OverFEAT ImageNet CNN features}

Those features are extracted from a convolutional network-based image features extractor OverFeat. They were trained on the ImageNet dataset (tens of millions images). The size is the output of the fifth layer of the neural network which is $1024 \times 6 \times 6 = 36864$. In our feature, we also have an extra dimension which can be ruled out.

\subsection{Data visualization and cleaning}

\end{document}

### 2015-12-17 Projet II/binaryClassification.m

@@ -1,0 +1,40 @@
+% Written by Diego Antognini & Jason Racine, EPFL 2015
+% all rights reserved
+
+clear all;
+close all;
+
+load train/train.mat;
+
+y = train.y;
+X_cnn = train.X_cnn;
+X_hog = train.X_hog;
+
+y(find(y == 1)) = 0;
+y(find(y == 2)) = 0;
+y(find(y == 3)) = 0;
+y(find(y == 4)) = 1;
+
+numberOfExperiments = 30;
+proportionOfTraining = 0.8;
+K = 20;
+
+%%
+%**********************************
+%            TEST 1
+%**********************************
+
+X = X_hog;
+fprintf('Test 1\n')
+for i = 1:numberOfExperiments
+    fprintf('%d ', i);
+    setSeed(28111993*i);
+    [XTr, yTr, XTe, yTe] = splitProp(proportionOfTraining, y, X);
+     
+    out = 0;
+    
+    err1(i) = balancedErrorRate(yTe, repmat([out], length(yTr), 1));
+
+end
+fprintf('\n%f\n', mean(err1));
+saveFile(err1, 'results/binary/err1');

### 2015-12-17 Projet II/createTrainingTestingCNN.m

@@ -1,0 +1,28 @@
+function [Tr, Te] = createTrainingTestingCNN(X, y, ratio)
+
+    % Initialize Tr, Te
+    Tr = [];
+    Te = [];
+
+    % Permute IDs and compute middle
+    idx = randperm(size(X,1));
+    mid = floor(length(idx) * ratio);
+
+    % CNN is a sparse matrix, better to use sparse representation
+    Xsparse = sparse(double(X));
+
+    % preparing training
+    Tr.idxs = idx(1:mid);
+    Tr.X = Xsparse(Tr.idxs,:);
+    Tr.y = y(Tr.idxs);
+
+    % preparing testing
+    Te.idxs = idx(mid+1:end);
+    Te.X = Xsparse(Te.idxs,:);
+    Te.y = y(Te.idxs);
+
+    % Normalizing
+    [Tr.nX, u, s] = zscore(Tr.X);
+    Te.nX = normalize(Te.X, u, s);
+
+end

### 2015-12-17 Projet II/createTrainingTestingHOG.m

@@ -1,0 +1,25 @@
+function [Tr, Te] = createTrainingTestingHOG(X, y, ratio)
+
+    % Initialize Tr, Te
+    Tr = [];
+    Te = [];
+
+    % Permute IDs and compute middle
+    idx = randperm(size(X,1));
+    mid = floor(length(idx) * ratio);
+
+    % preparing training
+    Tr.idxs = idx(1:mid);
+    Tr.X = X(Tr.idxs,:);
+    Tr.y = y(Tr.idxs);
+
+    % preparing testing
+    Te.idxs = idx(mid+1:end);
+    Te.X = X(Te.idxs,:);
+    Te.y = y(Te.idxs);
+
+    % Normalizing
+    [Tr.nX, u, s] = zscore(Tr.X);
+    Te.nX = normalize(Te.X, u, s);
+
+end

### 2015-12-17 Projet II/exploratoryDataAnalysis.m

@@ -1,0 +1,85 @@
+% Written by Diego Antognini & Jason Racine, EPFL 2015
+% all rights reserved
+
+clear all;
+close all;
+
+load train/train.mat;
+
+y = train.y;
+X_cnn = train.X_cnn;
+X_hog = train.X_hog;
+
+y(find(y == 1)) = 0;
+y(find(y == 2)) = 0;
+y(find(y == 3)) = 0;
+y(find(y == 4)) = 1;
+
+%% Classes
+
+lc1 = length(find(y == 0)); % 60.3%
+lc2 = length(find(y == 1)); % 39.7%
+
+%% HOG
+% Explanation what is it
+min_hog = min(unique(X_hog)); % 0
+max_hog = max(unique(X_hog)); % 0.2
+rank_hog = rank(X_hog); % 3467
+corr_hog = corr(double(X_hog), double(y)); %[-0.243;0.3]
+
+%% CNN
+% Sparse !
+% Explanation what is it
+
+min_cnn = min(unique(X_cnn)); % 0
+max_cnn = max(unique(X_cnn)); % 64.87
+rank_cnn = rang(X_cnn); % 5997
+corr_cnn = corr(double(X_cnn), double(y)); %[-0.2393;0.1795]
+
+%% Histogram
+
+f1 = figure;
+histogram(y);
+xlabel('Y');
+ylabel('Count');
+xlim([0 1]);
+
+%% Histogram feature HOG (distribution)
+
+idx = find(y == 0);
+X = X_hog;
+for feature=1:10%size(X,2)
+    f = figure;
+    histogram(X(idx,feature));
+    xlabel(sprintf('Feature %d', feature));
+    ylabel('Count');
+end
+
+idx = find(y == 1);
+X = X_hog;
+for feature=1:10%size(X,2)
+    f = figure;
+    histogram(X(idx,feature));
+    xlabel(sprintf('Feature %d', feature));
+    ylabel('Count');
+end
+
+%% Histogram feature CNN (distribution)
+
+idx = find(y == 0);
+X = X_cnn;
+for feature=1:10%size(X,2)
+    f = figure;
+    histogram(X(idx,feature));
+    xlabel(sprintf('Feature %d', feature));
+    ylabel('Count');
+end
+
+idx = find(y == 1);
+X = X_cnn;
+for feature=1:10%size(X,2)
+    f = figure;
+    histogram(X(idx,feature));
+    xlabel(sprintf('Feature %d', feature));
+    ylabel('Count');
+end

### 2015-12-17 Projet II/main.m

@@ -10,7 +10,7 @@ img = imread(sprintf('train/imgs/train%05d.jpg', i));
 subplot(121);
 imshow(img); % image itself
 
-subplot(122);
-feature = reshape(train.X_hog(i,:), 13, 13, 32);
+feature = hog( single(img)/255, 17, 8);
 im( hogDraw(feature) ); colormap gray;
 axis off; colorbar off;
+

### 2015-12-17 Projet II/prepareDataHOG.m

@@ -1,0 +1,43 @@
+function[Tr, Te] = prepareDataHOG(Tr_, Te_, M)
+    [Tr, Te] = PCA(Tr_, Te_, M);
+end
+
+function [Tr, Te] = PCA(Tr_, Te_, M)
+    Tr = Tr_;
+    Te = Te_;
+    
+    X = Tr.nX;
+    N = size(X, 1);
+    D = size(X, 2);
+    S = cov(X, 1);
+    
+    % Manually sort eigenvectors
+    [U, l] = eig(S, 'vector'); % eigenvector and eigenvalues
+    [l, I] = sort(l, 1, 'descend');
+    U = U(:, I);
+
+    %J = computeDistortionMatrix(l, D);
+    %displayDistortionMatrix(J);
+    
+    Tr.nZ = Tr.nX * U(:, 1:M);
+    Te.nZ = Te.nX * U(:, 1:M);
+    
+    Tr.nZU = Tr.nX * U(:, 1:M) * U(:, 1:M)';
+    Te.nZU = Te.nX * U(:, 1:M) * U(:, 1:M)';
+end
+
+function [J] = computeDistortionMatrix(l, D)
+    % Here D < N
+    J = zeros(D, 1);
+    for m = 1:D
+        J(m) = sum(l(m+1:D));
+    end
+end
+
+function [] = displayDistortionMatrix(J)
+    figure('Name', 'Distortion Measure');
+    plot(J, 'LineWidth', 4);
+    xlabel('M');
+    ylabel('J');
+    title('distortion J vs M');
+end

### 2015-12-17 Projet II/splitProp.m

@@ -1,0 +1,24 @@
+% Written by Diego Antognini & Jason Racine, EPFL 2015
+% all rights reserved
+
+function [XTr, yTr, XTe, yTe] = splitProp(prop, y, X)
+    % split the data into train and test given a proportion
+    N = length(y);
+
+    shuffle = randperm(N);
+    Xrandom = X(shuffle,:);
+    yrandom = y(shuffle,:);
+    
+	% generate random indices
+	idx = randperm(N);
+    Ntr = floor(prop * N);
+	% select few as training and others as testing
+	idxTr = idx(1:Ntr);
+	idxTe = idx(Ntr+1:end);
+	% create train-test split
+    XTr = Xrandom(idxTr,:);
+    yTr = yrandom(idxTr);
+    XTe = Xrandom(idxTe,:);
+    yTe = yrandom(idxTe);
+end
+

### 2015-12-19 Projet II/report/paper.tex
Add K-fold cross validation

@@
 
-\section{Binary Classification}
-
-\subsection{Data Description}
+\section{Data Description}
 
@@
 Our input $\mathbf{X}$ is the concatenation of $\mathbf{X}_{hog}$ with $\mathbf{X}_{cnn}$.
-\\Each input sample is a vector $\mathbf{x}_n$ with dimension $D = 42273$ ($5408$ for the HOG and $36865$ for the CNN) and is the concatenation of $\mathbf{x}_n^{hog}$ with $\mathbf{x}_n^{cnn}$
-The output ($\mathbf{y}$) represents the classification of the images, with $0$ if it represents a car, a horse or a plane, $1$ otherwise.
+
+Each input sample is a vector $\mathbf{x}_n$ with dimension $D = 42273$ ($5408$ for the HOG and $36865$ for the CNN) and is the concatenation of $\mathbf{x}_n^{hog}$ with $\mathbf{x}_n^{cnn}$
+The output ($\mathbf{y}$) represents the classification of the images.
+For the binary classification, the label $1$  represents a car, a horse or a plane and the label $2$ anything else.
+For the multi-class classification, the label $1$ represents a car, $2$2 a horse, $3$ a plane and $4$ anything else.
 
@@
 
-\subsubsection{Histogram of Oriented Gradients}
+\subsection{Histogram of Oriented Gradients}
 
@@
 The size is the output of the fifth layer of the neural network which is $1024 \times 6 \times 6 = 36864$.
-In our feature, we also have an extra dimension which can be ruled out.
+In our features, we also have an extra dimension which can be ruled out.
 
-\subsection{Data visualization and cleaning}
+\section{Data visualization and cleaning}
 

### 2015-12-19 Projet II/crossValidateModels_CNN.m

@@ -8,10 +8,10 @@ addpath(genpath('DeepLearnToolbox'));
 load train/train.mat;
 
 
-numberOfExperiments = 30;
+numberOfExperiments = 1;
 proportionOfTraining = 0.8;
 K = 10;
-M = 300;
+M = 150;
 
 %% Create data
 fprintf('Creating Train & Test sets\n');
@@ -30,6 +30,7 @@ toc
 
 N = length(Tr.y);
 
+%% NN
 % Setup NN.
 inputSize  = M;
 innerSize  = 700;
@@ -55,7 +56,7 @@ innerSizes = [500 600 700 800 900 100]; % Best so far 700 : 8.59%
 learningRates = [2.0 3.0 4.0 5.0]; % Best so far 2 : 8.3%
 
 for j = 1:1:numberOfExperiments
-    setSeed(28111993*i);
+    setSeed(28111993*j);
 
     fprintf('%d : Split the data\n', j);
     [TTr, TTe] = splitProp(proportionOfTraining, Tr_, false);
@@ -75,7 +76,7 @@ for j = 1:1:numberOfExperiments
 
             [errnZ, nnPrednZ] = neuralNetworks(TTTr.nZ, TTTr.y, TTTe.nZ, TTTe.y, inputSize, innerSize, numepochs, batchsize, learningRate, binaryClassification);
 
-            err_te(k,j) = errnZ;
+            err_te(k,i) = errnZ;
         end
         %end
     end
@@ -89,4 +90,39 @@ for j = 1:1:numberOfExperiments
     learningRateStar = learningRates(learningRateStar);
 end
 
+%% DT
 
+Tr_ = Tr;
+Te_ = Te;
+
+% 10.85 with 500
+NLeaves = [100 200 300 400 500 600 700 800 900];
+
+for j = 1:1:numberOfExperiments
+    setSeed(28111993*j);
+
+    fprintf('%d : Split the data\n', j);
+    [TTr, TTe] = splitProp(proportionOfTraining, Tr_, false);
+    
+    idxCV = splitGetCV(K, length(TTr.y));
+    
+    % K-fold
+    for k=1:1:K
+        fprintf('%d : %dth fold\n', j, k);
+        [TTTr, TTTe] = splitGetTrTe(TTr, idxCV, k, false);
+        
+        for i = 1:1:length(NLeaves)
+            NLeave = NLeaves(i);
+            
[…]

### 2015-12-19 Projet II/report/paper.tex
Add new multi class model

@@
 }
-
-% The \author macro works with any number of authors. There are two commands
-% used to separate the names and addresses of multiple authors: \And and \AND.
-%
-% Using \And between authors leaves it to \LaTeX{} to determine where to break
-% the lines. Using \AND forces a linebreak at that point. So, if \LaTeX{}
-% puts 3 of 4 authors names on the first line, and the last on the second
-% line, try using \AND instead of \And before the third author name.
 
@@
 %http://vision.ucsd.edu/~pdollar/toolbox/doc/channels/hog.html
-\subsubsection{OverFEAT ImageNet CNN features}
+\subsection{OverFEAT ImageNet CNN features}
 
@@
 
-\section{Data visualization and cleaning}
+\section{Data visualization and analysis}
+
+We first have plot the labels distribution for the binary and multi-class cases.
+For the binary case, we have $60.3\%$ for the label $1$ and $39.8\%$ for the second one.
+We have observed that there is a imbalance between the classes but this is quite small.
+For the multi-classes case, we obtain $16.07\%$, $19.37\%$, $24.87\%$ and $39.70\%$ for the labels $1$, $2$, $3$ and $4$.
+This time, the imbalance is more important, especially with the label $4$ and we do have to take care about it during the training.
+
+We also wanted to analyze the distribution of our features, and have observed that for the features HOG, each of them has a has similar distribution as showed in Figure XXX.
+The values are essentially in the range $[0, 0.2]$.
+We are also interested about the correlation between the input and output variables.
+We have observed that the correlation is in the range $[-0.24 ;0.30]$, and so can conclude that features seem not to have a direct correlation between them.
+However, a combination of features can have a high correlation with the output but we couldn't find such combination.
+Finally, $\mathbf{X}_{hog}$ is rank-deficient with a rank of $3467$ instead of $5408$.
+
+For the features CNN, first we can see that it is a sparse matrix and most of the values for each data is $0$ (around $95.5\%$).
+This means that for each data, a small subset is able to describe the picture.
+Moreover, we can observe that we are faced to the problem of $D > N$ and the solution we have adopted is PCA which will describe during the next sections.
+The values are in mainly in the range $[0, 64.87]$.
+We also wanted to see the correlation of those features with the output and have found that they are in the range $[-0.24;0.18]$, which is similar to the HOG features.
+Finally,  $\mathbf{X}_{cnn}$ is rank-deficient with a rank of $5997$ instead of $36865$.
+
+\section{Techniques used}
+
+This section is a description of the technique we have used in the project. It describes mainly how it works, why we thought it could be useful in this project.
+
+\subsection{Principal component analysis}
+
+As we have seen in the section "Data visualization and analysis", we are faced to the $D > N$ problem with the CNN features.
+One way to solve this problem is to use PCA.
+Basically, PCA works as follow : we compute the mean and the covariance matrix $S = \frac{1}{N}\sum_{n=1}^{N}(x_n-\bar{x})(x_n-\bar{x})^T$ of our data and then compute the $M$ eigenvectors of $S$ corresponding to the $M$ largest eigenvalues.
+$M$ is the number of features we have to take into account in order to minimize the reconstruction.
+Finally, the approximation of a data sample is $\tilde{x_n} = \sum_{i=1}^M x_{ni} u_i$.
+However, we couldn't compute PCA using this method, because computing the full eigenvector decomposition a matrix $D \times D$ runs in  $O(D^3)$ and D is quite large.
+We have used an alternative proposed in [REF] which allows us to compute the same eigenvectors in $O(N^3)$, which is roughly at least $230$ times faster.
+The tricks is to express $S = \frac{1}{N}XX^T$ and finally using $\frac{1}{N}XX^Tv_i = \lambda_i v_i$, where $v_i=Xu_i$, we obtain an eigenvector equation for the $N \times N$ matrix $S$.
+
+The last parameter to set, is the parameter $M$ which defines the $M$ largest eigenvectors to take into consideration for our features selection.We introduce a distortion measure $J = \frac{1}{N}\sum_{n=1}^{N}||x_n-\tilde{x_n}||^2$ which represents the mean squared distance between the original data sample and our approximation.
+Without going into the steps, the solution to the minimization of $J$ is $J = \sum_{i=M+1}^D \lambda_i$.
+
+The Figure XX shows the plot of $J$ for the HOG features and the CNN features.
+For the CNN, we have tried several values for $M$ and have conclude that $M = 150$ was a good compromise between accuracy and speed.
+The corresponding distortion might be very high compared to the HOG features, but the range of the values are not the same, CNN has values going larger than $0.2$.
+We tried some bigger values but we didn't get major improvements in terms of error.
+For the HOG features, the distortion measure seems more nice, and have decided to take $M = XXX$.
+BECAUSE BLA BLA BLA
+
+\section{Evaluation methods}
+
+\section{Model comparison}
+
+\subsection{Binary classification}
+
+\subsection{Multi-class classification}
+
+\section{Feature transformation}
+
+\section{Conclusion}
+
+\section{References}
+
+@misc{IMM2012-06284,
+    author       = "R. B. Palm",
+    title        = "Prediction as a candidate for learning deep hierarchical models of data",
+    year         = "2012",
+}
+
+@misc{PMT,
+   author = {Piotr Doll\'ar}, 
+   title = {{P}iotr's {C}omputer {V}ision {M}atlab {T}oolbox ({PMT})}, 
+   howpublished = {\url{http://vision.ucsd.edu/~pdollar/toolbox/doc/index.html}} 
+}
 

### 2015-12-19 Projet II/crossValidateModels_CNN.m

@@ -126,3 +126,41 @@ for j = 1:1:numberOfExperiments
     [errStar, NLeaveStarId] = min(mseTe);
     NLeaveStar = NLeaves(NLeaveStarId);
 end
+
+%% RF
+    
+Tr_ = Tr;
+Te_ = Te;
+
+% Best is 500 (0.1072), however NTrees growsl linearly and so, it is a good
+% comprises between speed an accuracy
+NTrees = [100 200 300 400 500];
+
+for j = 1:1:numberOfExperiments
+    setSeed(28111993*j);
+
+    fprintf('%d : Split the data\n', j);
+    [TTr, TTe] = splitProp(proportionOfTraining, Tr_, false);
+    
+    idxCV = splitGetCV(K, length(TTr.y));
+    
+    % K-fold
+    for k=1:1:K
+        fprintf('%d : %dth fold\n', j, k);
+        [TTTr, TTTe] = splitGetTrTe(TTr, idxCV, k, false);
+        
+        for i = 1:1:length(NTrees)
+            NTree = NTrees(i);
+            tic
+            BaggedEnsemble = TreeBagger(NTree, TTTr.Z, TTTr.y);
+            yhat = str2double(predict(BaggedEnsemble, TTTe.Z));
+            toc
+            err_te(k,i) = balancedErrorRate(TTTe.y, yhat);
+        end
+    end
+
+    mseTe = mean(err_te);
+
+    [errStar, NTreeStarId] = min(mseTe);
+    NTreeStar = NTrees(NLeaveStarId);
+end

### 2015-12-19 Projet II/exploratoryDataAnalysis.m

@@ -10,15 +10,22 @@ y = train.y;
 X_cnn = train.X_cnn;
 X_hog = train.X_hog;
 
-y(find(y == 1)) = 0;
-y(find(y == 2)) = 0;
-y(find(y == 3)) = 0;
-y(find(y == 4)) = 1;
+y(find(y == 1)) = 1;
+y(find(y == 2)) = 1;
+y(find(y == 3)) = 1;
+y(find(y == 4)) = 2;
 
 %% Classes
 
-lc1 = length(find(y == 0)); % 60.3%
-lc2 = length(find(y == 1)); % 39.7%
+lc1 = length(find(y == 1)); % 16.07%
+lc2 = length(find(y == 2)); % 19.37%
+lc3 = length(find(y == 3)); % 24.87%
+lc4 = length(find(y == 4)); % 39.7%
+
+%% Classes
+
+lc1 = length(find(y == 1)); % 60.3%
+lc2 = length(find(y == 2)); % 39.7%
 
 %% HOG
 % Explanation what is it
@@ -42,11 +49,11 @@ f1 = figure;
 histogram(y);
 xlabel('Y');
 ylabel('Count');
-xlim([0 1]);
+xlim([1 2]);
 
 %% Histogram feature HOG (distribution)
 
-idx = find(y == 0);
+idx = find(y == 1);
 X = X_hog;
 for feature=1:10%size(X,2)
     f = figure;
@@ -55,7 +62,7 @@ for feature=1:10%size(X,2)
     ylabel('Count');
 end
 
-idx = find(y == 1);
+idx = find(y == 2);
 X = X_hog;
 for feature=1:10%size(X,2)
     f = figure;
@@ -66,7 +73,7 @@ end
 
 %% Histogram feature CNN (distribution)
 
-idx = find(y == 0);
+idx = find(y == 1);
 X = X_cnn;
 for feature=1:10%size(X,2)
     f = figure;
@@ -75,7 +82,7 @@ for feature=1:10%size(X,2)
     ylabel('Count');
 end
 
-idx = find(y == 1);
+idx = find(y == 2);
 X = X_cnn;
 for feature=1:10%size(X,2)
     f = figure;

### 2015-12-19 Projet II/multiClassification.m

@@ -26,6 +26,7 @@ tic
 toc
 
 %%
+% 75%
 %**********************************
 %            TEST 1
 %**********************************
@@ -79,7 +80,7 @@ end
 fprintf('\n%f\n', mean(err2));
 saveFile(err2, 'results/multi/err2');
 
-%%  %
+%% 10.39%
 %**********************************
 %            TEST 3
 %**********************************
@@ -106,10 +107,102 @@ for j = 1:1:numberOfExperiments
         CMdl = fitensemble(TTTr.Z, TTTr.y, 'Bag', NLeaves, 'Tree', 'Type', 'classification');
         yhat = predict(CMdl, TTTe.Z);
 
-        err_te(k,i) = balancedErrorRate(TTTe.y, yhat);
+        err_te(k) = balancedErrorRate(TTTe.y, yhat);
     end
     toc
     err3(j) = mean(err_te);
 end
 fprintf('\n%f\n', mean(err3));
 saveFile(err3, 'results/multi/err3');
+
+%%  %
+%**********************************
+%            TEST 4
+%**********************************
+
+N = length(Tr.y);
+
+% Setup 
+NTrees = 500;
+
+Tr_ = Tr;
+Te_ = Te;
+
+for j = 1:1:numberOfExperiments
+    setSeed(28111993*j);
+
+    fprintf('%d ', j);
+    [TTr, TTe] = splitProp(proportionOfTraining, Tr_, false);
+        
+    idxCV = splitGetCV(K, length(TTr.y));
+    
+    % K-fold
+    for k=1:1:K
+        [TTTr, TTTe] = splitGetTrTe(TTr, idxCV, k, false);
+        
+        BaggedEnsemble = TreeBagger(NTrees, TTTr.Z, TTTr.y);
+        yhat = str2double(predict(BaggedEnsemble, TTTe.Z));
+
+        err_te(k) = balancedErrorRate(TTTe.y, yhat);
+    end
+    err4(j) = mean(err_te);
+end
+fprintf('\n%f\n', mean(err4));
+saveFile(err4, 'results/multi/err4');
+
+%%  %
+%**********************************
+%            TEST 5
+%**********************************
+
+% Setup 
+t = templateSVM();%'Solver', 'ISDA', 'KernelFunction', 'rbf', 'BoxConstraint', Inf);   
+
+Tr_ = Tr;
+Te_ = Te;
+
[…]

### 2015-12-19 Projet II/prepareDataHOG.m

@@ -16,8 +16,8 @@ function [Tr, Te] = PCA(Tr_, Te_, M)
     [l, I] = sort(l, 1, 'descend');
     U = U(:, I);
 
-    %J = computeDistortionMatrix(l, D);
-    %displayDistortionMatrix(J);
+    J = computeDistortionMatrix(l, D);
+    displayDistortionMatrix(J);
     
     Tr.nZ = Tr.nX * U(:, 1:M);
     Te.nZ = Te.nX * U(:, 1:M);

### 2015-12-19 Projet II/report/paper.tex
Advance

@@
 
+\subsection{Support Vector Machine}
+
+\subsection{Neural Networks}
+
+This is the multi-layer perceptron (MLP) we have seen in class.
+The neural networks used in this project comes from the given DeepLearning matlab toolbox.
+It uses batch sizes in order to use less memory (take the first $n$ data samples and works on them, then the next $n$ etc).
+We have let the default parameter ($100$).
+The number of epochs is simply the number of time where all the training data do a forward and a backward pass.
+We have put this value to $20$, in order to have a little more accuracy without loosing a lot of time, it was a good compromise because larger epochs don't mean especially better accuracy.
+the number of input is $M$ and output is either $2$ or $4$, depending of the type of classification (binary or multi-class).
+For the number of hidden layer, we have let $1$ because we didn't see improvement by adding some more.
+However, the number of neurons in the hidden layer has been tested and have found $1000$ for the binary classification and $700$ for the multi-class case.
+We also have to set up the learning rate : the optimal value we have found is $3$ for the binary case and $2$ for the multi-class one.
+By default, the activation function used is the \textit{tanh} one.
+We didn't see so much different using the \textit{sigmoid}.
+Finally, the algorithm uses stochastic gradient descent using a momentum of $0.5$, which might help to converge faster.
+
+\subsection{Decision Trees}
+
+\subsection{Random Forests}
+
 \section{Evaluation methods}

### 2015-12-19 Projet II/SVM_HOG_PCA.m

@@ -8,7 +8,7 @@ addpath(genpath('DeepLearnToolbox'));
 load train/train.mat;
 
 ratio = 0.8;
-M = 3000;
+M = 500;
 
 %% Create data
 fprintf('Creating Train & Test sets\n');
@@ -27,7 +27,7 @@ toc
 
 fprintf('Training\n');
 
-binaryClassification = true;
+binaryClassification = false;
 
 yTr = Tr.y;
 yTe = Te.y;
@@ -58,21 +58,21 @@ end
     
     t = templateSVM();%'Solver', 'ISDA', 'KernelFunction', 'polynomial', 'PolynomialOrder', 7, 'BoxConstraint', Inf);
     
-    % Training using X
-    Mdl = fitcecoc(Tr.X, yTr, 'Learners', t, 'Options', statset('UseParallel', 1));
-    CMdl = compact(discardSupportVectors(Mdl));
-    yhat = predict(CMdl, Te.X);
-
-    errX = balancedErrorRate(yTe, yhat);
-    fprintf('BER Testing error  X: %.2f%%\n', errX * 100);
-
-    % Training using nX
-    Mdl = fitcecoc(Tr.nX, yTr, 'Learners', t, 'Options', statset('UseParallel', 1));
-    CMdl = compact(discardSupportVectors(Mdl));
-    yhat = predict(CMdl, Te.nX);
-
-    errnX = balancedErrorRate(yTe, yhat);
-    fprintf('BER Testing error nX: %.2f%%\n', errnX * 100);
+%     % Training using X
+%     Mdl = fitcecoc(Tr.X, yTr, 'Learners', t, 'Options', statset('UseParallel', 1));
+%     CMdl = compact(discardSupportVectors(Mdl));
+%     yhat = predict(CMdl, Te.X);
+% 
+%     errX = balancedErrorRate(yTe, yhat);
+%     fprintf('BER Testing error  X: %.2f%%\n', errX * 100);
+% 
+%     % Training using nX
+%     Mdl = fitcecoc(Tr.nX, yTr, 'Learners', t, 'Options', statset('UseParallel', 1));
+%     CMdl = compact(discardSupportVectors(Mdl));
+%     yhat = predict(CMdl, Te.nX);
+% 
+%     errnX = balancedErrorRate(yTe, yhat);
+%     fprintf('BER Testing error nX: %.2f%%\n', errnX * 100);
 
     % Training using Z
[…]

### 2015-12-19 Projet II/binaryClassification.m

@@ -1,40 +1,155 @@
-% Written by Diego Antognini & Jason Racine, EPFL 2015
-% all rights reserved
-
+%% Initialization
 clear all;
 close all;
 
+addpath(genpath('Piotr'));
+addpath(genpath('DeepLearnToolbox'));
+
 load train/train.mat;
 
-y = train.y;
-X_cnn = train.X_cnn;
-X_hog = train.X_hog;
-
-y(find(y == 1)) = 0;
-y(find(y == 2)) = 0;
-y(find(y == 3)) = 0;
-y(find(y == 4)) = 1;
-
-numberOfExperiments = 30;
+numberOfExperiments = 10;
 proportionOfTraining = 0.8;
-K = 20;
+K = 10;
+M = 150;
+
+train.y(find(train.y == 2)) = 1;
+train.y(find(train.y == 3)) = 1;
+train.y(find(train.y == 4)) = 2;
+
+%% Create data
+fprintf('Creating Train & Test sets\n');
+tic
+[Tr, Te] = createTrainingTestingCNN(train.X_cnn, train.y, proportionOfTraining);
+toc
+
+%% Prepare data
+
+fprintf('Prepare the data for the training\n');
+tic
+[Tr, Te] = prepareDataCNN(Tr, Te, M);
+toc
 
 %%
+% 50%
 %**********************************
 %            TEST 1
 %**********************************
 
-X = X_hog;
 fprintf('Test 1\n')
 for i = 1:numberOfExperiments
     fprintf('%d ', i);
     setSeed(28111993*i);
-    [XTr, yTr, XTe, yTe] = splitProp(proportionOfTraining, y, X);
-     
-    out = 0;
-    
-    err1(i) = balancedErrorRate(yTe, repmat([out], length(yTr), 1));
-
+    out = 1;
+    err1(i) = balancedErrorRate(Te.y, repmat([out], length(Tr.y), 1));
 end
 fprintf('\n%f\n', mean(err1));
 saveFile(err1, 'results/binary/err1');
+
+%% 9.96%
+%**********************************
+%            TEST 2
+%**********************************
+
+N = length(Tr.y);
+
+% Setup NN.
+inputSize  = M;
+innerSize  = 1000;
+numepochs = 20;
+batchsize = 100;
+learningRate = 3;
+binaryClassification = true;
+
+Tr_ = Tr;
+Te_ = Te;
+
+for j = 1:1:numberOfExperiments
[…]

### 2015-12-19 Projet II/crossValidateModels_CNN_bin.m

@@ -1,0 +1,166 @@
+%% Initialization
+clear all;
+close all;
+
+addpath(genpath('Piotr'));
+addpath(genpath('DeepLearnToolbox'));
+
+load train/train.mat;
+
+
+numberOfExperiments = 1;
+proportionOfTraining = 0.8;
+K = 10;
+M = 150;
+
+%% Create data
+fprintf('Creating Train & Test sets\n');
+tic
+[Tr, Te] = createTrainingTestingCNN(train.X_cnn, train.y, proportionOfTraining);
+toc
+
+%% Prepare data
+
+fprintf('Prepare the data for the training\n');
+tic
+[Tr, Te] = prepareDataCNN(Tr, Te, M);
+toc
+
+%% Cross validate
+
+N = length(Tr.y);
+
+%% NN
+% Setup NN.
+inputSize  = M;
+innerSize  = 1000;
+numepochs = 20;
+batchsize = 100;
+learningRate = 2;
+binaryClassification = true;
+
+Tr_ = Tr;
+Te_ = Te;
+    
+if binaryClassification    
+    Tr_.y(find(Tr_.y == 2)) = 1;
+    Tr_.y(find(Tr_.y == 3)) = 1;
+    Tr_.y(find(Tr_.y == 4)) = 2;
+    
+    Te_.y(find(Te_.y == 2)) = 1;
+    Te_.y(find(Te_.y == 3)) = 1;
+    Te_.y(find(Te_.y == 4)) = 2;
+end
+
+innerSizes = [1000]; % Best so far 1000 : 9.35%
+learningRates = [3.0]; % Best so far 3 : 9.56%
+
+for j = 1:1:numberOfExperiments
+    setSeed(28111993*j);
+
+    fprintf('%d : Split the data\n', j);
+    [TTr, TTe] = splitProp(proportionOfTraining, Tr_, false);
+    
+    
+    idxCV = splitGetCV(K, length(TTr.y));
+    
+    % K-fold
+    for k=1:1:K
+        fprintf('%d : %dth fold\n', j, k);
+        [TTTr, TTTe] = splitGetTrTe(TTr, idxCV, k, false);
+        
+        %for i = 1:1:length(innerSizes)
+        %    innerSize = innerSizes(i);
+        for i = 1:1:length(learningRates)
+            learningRate = learningRates(i);
+
[…]

### 2015-12-19 Projet II/multiClassification.m

@@ -172,8 +172,9 @@ for j = 1:1:numberOfExperiments
     % K-fold
     for k=1:1:K
         [TTTr, TTTe] = splitGetTrTe(TTr, idxCV, k, false);
-        
+        tic
         Mdl = fitcecoc(TTTr.Z, TTTr.y, 'Learners', t, 'Options', statset('UseParallel', 1));
+        toc
         CMdl = compact(discardSupportVectors(Mdl));
         yhat = predict(CMdl, TTTe.Z);
     
@@ -185,7 +186,7 @@ end
 fprintf('\n%f\n', mean(err5));
 saveFile(err5, 'results/multi/err4');
 
-%%
+%% Plot
 s = [1 numberOfExperiments];
 
 err1 = openFile('results/multi/err1', s);

### 2015-12-19 Projet II/prepareDataHOG.m

@@ -16,8 +16,8 @@ function [Tr, Te] = PCA(Tr_, Te_, M)
     [l, I] = sort(l, 1, 'descend');
     U = U(:, I);
 
-    J = computeDistortionMatrix(l, D);
-    displayDistortionMatrix(J);
+    %J = computeDistortionMatrix(l, D);
+    %displayDistortionMatrix(J);
     
     Tr.nZ = Tr.nX * U(:, 1:M);
     Te.nZ = Te.nX * U(:, 1:M);

### 2015-12-20 Projet II/report/paper.tex
Add test4 & 5

@@
 For the binary case, we have $60.3\%$ for the label $1$ and $39.8\%$ for the second one.
-We have observed that there is a imbalance between the classes but this is quite small.
+We have observed that there is a small imbalance between the classesl.
 For the multi-classes case, we obtain $16.07\%$, $19.37\%$, $24.87\%$ and $39.70\%$ for the labels $1$, $2$, $3$ and $4$.
@@
 
-As we have seen in the section "Data visualization and analysis", we are faced to the $D > N$ problem with the CNN features.
+As we have seen in the section "Data visualization and analysis", we are faced to the $D > N$ problem mainly with the CNN features and also the HOG features depending on how we se the $k$ for the k-fold cross validation.
 One way to solve this problem is to use PCA.
@@
 
-The last parameter to set, is the parameter $M$ which defines the $M$ largest eigenvectors to take into consideration for our features selection.We introduce a distortion measure $J = \frac{1}{N}\sum_{n=1}^{N}||x_n-\tilde{x_n}||^2$ which represents the mean squared distance between the original data sample and our approximation.
+The last parameter to set, is the parameter $M$ which defines the $M$ largest eigenvectors to take into consideration for our feature selection.We introduce a distortion measure $J = \frac{1}{N}\sum_{n=1}^{N}||x_n-\tilde{x_n}||^2$ which represents the mean squared distance between the original data sample and our approximation.
 Without going into the steps, the solution to the minimization of $J$ is $J = \sum_{i=M+1}^D \lambda_i$.
@@
 The Figure XX shows the plot of $J$ for the HOG features and the CNN features.
-For the CNN, we have tried several values for $M$ and have conclude that $M = 150$ was a good compromise between accuracy and speed.
+For the CNN, we have tried several values for $M$ and have conclude that $M_{CNN} = 150$ was a good value allowing us to compute faster results without loosing too much accuracy.
 The corresponding distortion might be very high compared to the HOG features, but the range of the values are not the same, CNN has values going larger than $0.2$.
 We tried some bigger values but we didn't get major improvements in terms of error.
-For the HOG features, the distortion measure seems more nice, and have decided to take $M = XXX$.
-BECAUSE BLA BLA BLA
-
-\subsection{Support Vector Machine}
+For the HOG features, the distortion measure seems more nice, we did the same experiments and finally set $M_{HOG}$ = $500$.
 
@@
 Finally, the algorithm uses stochastic gradient descent using a momentum of $0.5$, which might help to converge faster.
+
+Neural networks are suitable for problems which might be difficult to describe in a mathematically manner and is robust to the noise, which is very useful in our case.
+However, we have to be careful with the weights initialization because neural networks are very sensitive to them, and especially with overfitting and computing the generalization error, this task is harder with neural networks.
+
+\subsection{Support Vector Machine}
+
+The normal support vector machine is a technique used for binary classification.
+The goal is to divide the space using an hyperplan in a way to maximize the margin, which represents the distance between closest members of the two classes.
+It is an interesting method because it has a regularization parameter which help to avoid overfitting.
+Another advantages is we can easily change the kernel function in order to try to find better boundaries.
+However, what is difficult is to find good parameters.
+We might have many support vectors and overfit if we set the $C$ too large and on the contrary if it is small, we might have not enough support vector and underfit.
+However, the best results we can obtain with support vector machine need to have a small imbalanced between the positive and negative examples.
+As seen in the section "Data visualization and analysis", there classes are not very well balanced as much for the binary classification as the multi-class classification.
+We have use the same algorithm we have seen in the lab $SMO$.
+
+To adapt support vector machine to a multi-class classifier, we have used \textit{One Versus All} method, which divides the problem in $4$ binary classifications.
+We train the data with all the classifiers and at the end, each classifier says if the data is positive to its class.
+If there are more than one classifier with the data sample considered as positive, we use the highest confidence to associate the class.
+Another possible method would have been to use \textit{One Versus One} which needs $6$ classifiers in our case, and is less sensitive to the imbalance but needs more computational resource, but it didn't give better results.
 

### 2015-12-20 Projet II/SVM_CNN_PCA.m

@@ -44,32 +44,26 @@ if binaryClassification
     yTe(find(yTe == 4)) = 2;
 end
 
-    % Binary
-    % Default : 9.73% (nZ), 9.31% (Z)
-    % Linear : ~%
-    % RBF : ~%
-    % Poly^7 : ~%
-
-    % Multiclass
-    % Default : 9.73% (nZ), 9.7% (Z)
-    % Linear : ~27.11%
-    % RBF : ~26.88%
-    % Poly^7 : 26.65%
+    % Binary : 9.32% (Z)
     
-    t = templateSVM();%'Solver', 'ISDA', 'KernelFunction', 'rbf', 'BoxConstraint', Inf);   
+    % Multiclass : 9.05% (Z)
+    
+    t = templateSVM('Solver', 'SMO', 'KernelFunction', 'linear', 'IterationLimit', 1e6, 'KernelScale', 1, 'BoxConstraint', 1);   
       
     % Training using Z 
-    Mdl = fitcecoc(Tr.Z, yTr, 'Learners', t, 'Options', statset('UseParallel', 1));
+    tic
+    Mdl = fitcecoc(Tr.Z, yTr, 'Learners', t, 'Coding', 'onevsall', 'Options', statset('UseParallel', 1));
+    toc
     CMdl = compact(discardSupportVectors(Mdl));
     yhat = predict(CMdl, Te.Z);
     
     errZ = balancedErrorRate(yTe, yhat);
     fprintf('\nBER Testing error  Z: %.2f%%\n', errZ * 100);
     
-    % Training using nZ 
-    Mdl = fitcecoc(Tr.nZ, yTr, 'Learners', t, 'Options', statset('UseParallel', 1));
-    CMdl = compact(discardSupportVectors(Mdl));
-    yhat = predict(CMdl, Te.nZ);
-    
-    errnZ = balancedErrorRate(yTe, yhat);
-    fprintf('\nBER Testing error  nZ: %.2f%%\n', errnZ * 100);
+%     % Training using nZ 
+%     Mdl = fitcecoc(Tr.nZ, yTr, 'Learners', t, 'Coding', 'onevsall', 'Options', statset('UseParallel', 1));
+%     CMdl = compact(discardSupportVectors(Mdl));
+%     yhat = predict(CMdl, Te.nZ);
+%      
+%     errnZ = balancedErrorRate(yTe, yhat);
+%     fprintf('\nBER Testing error  nZ: %.2f%%\n', errnZ * 100);

### 2015-12-20 Projet II/multiClassification.m

@@ -151,13 +151,13 @@ end
 fprintf('\n%f\n', mean(err4));
 saveFile(err4, 'results/multi/err4');
 
-%%  %
+%% 8.71%
 %**********************************
 %            TEST 5
 %**********************************
 
 % Setup 
-t = templateSVM();%'Solver', 'ISDA', 'KernelFunction', 'rbf', 'BoxConstraint', Inf);   
+t = templateSVM('Solver', 'SMO', 'KernelFunction', 'linear', 'IterationLimit', 1e6, 'KernelScale', 1, 'BoxConstraint', 1);      
 
 Tr_ = Tr;
 Te_ = Te;
@@ -169,23 +169,22 @@ for j = 1:1:numberOfExperiments
     [TTr, TTe] = splitProp(proportionOfTraining, Tr_, false);
         
     idxCV = splitGetCV(K, length(TTr.y));
-    
+      
     % K-fold
     for k=1:1:K
         [TTTr, TTTe] = splitGetTrTe(TTr, idxCV, k, false);
-        tic
-        Mdl = fitcecoc(TTTr.Z, TTTr.y, 'Learners', t, 'Options', statset('UseParallel', 1));
-        toc
+        Mdl = fitcecoc(TTTr.Z, TTTr.y, 'Learners', t, 'Coding', 'onevsall', 'Options', statset('UseParallel', 1));
         CMdl = compact(discardSupportVectors(Mdl));
         yhat = predict(CMdl, TTTe.Z);
     
         err_te(k) = balancedErrorRate(TTTe.y, yhat);
     end
-    err5(j) = mean(err_te);
+    
+    err5(j) = err_te;
 end
 
 fprintf('\n%f\n', mean(err5));
-saveFile(err5, 'results/multi/err4');
+saveFile(err5, 'results/multi/err5');
 
 %% Plot
 s = [1 numberOfExperiments];
@@ -193,13 +192,17 @@ s = [1 numberOfExperiments];
 err1 = openFile('results/multi/err1', s);
 err2 = openFile('results/multi/err2',s);
 err3 = openFile('results/multi/err3', s);
+err4 = openFile('results/multi/err4', s);
+err5 = openFile('results/multi/err5', s);
 
 figure;
-boxplot([err1' err2' err3']);
+boxplot([err1' err2' err3' err4' err5']);
[…]

### 2015-12-20 Projet II/report/paper.tex
Add results

@@
 Another advantages is we can easily change the kernel function in order to try to find better boundaries.
-However, what is difficult is to find good parameters.
+However, what is difficult is to find good parameters ($C$, $\gamma$).
+$\gamma$ is the kernel scale parameter which is related to how the data are scattered.
 We might have many support vectors and overfit if we set the $C$ too large and on the contrary if it is small, we might have not enough support vector and underfit.
@@
 We have use the same algorithm we have seen in the lab $SMO$.
+
+We have used the method \textit{fitcecoc} in matlab, and we don't need to specify the parameter $C$, it is computed during the process.
+However, we have to set the kernel function and $\gamma$.
+We have k-fold cross validate and have found that the linear kernel was the best and implicitly $\gamma = 1$.
 

### 2015-12-20 Projet II/report/paper.tex
Add RFs

--- 2015-12-20 Projet II/report/paper.tex
@@
 We have use the same algorithm we have seen in the lab $SMO$.
+
+We have used the method \textit{fitcecoc} in matlab.
+We have used k-fold cross valiation to find $C$ and $\gamma$ and have set the margin $C=1$, and $\gamma = 1$.
+Only the linear kernel function was available in this mode, and so, gaussian kernel might give better results.
 
@@
 
-\subsection{Decision Trees}
+\subsection{Bootstrap Aggregation}
+
+As we have seen in the lecture, decision trees are good because the training and the prediction is quite fast, even with many features or large dataset but have a high variance.
+The idea of decision trees, is to have a several questions where to answer yes/no and to move in the tree.
+At the end, the leaves give the corresponding class.
+To construct this list of questions, we divide recursively the space in a way to minimize the cost of impurity (depending of the threshold to find and the current feature).
+The depth of the tree is a parameter.
+
+The idea of bootstrap aggregation is to average several decision trees.
+It will help to reduce the variance and avoid overfitting.
+It generates $m$ new training sets of the same size, by sampling uniformly and with replacement the original training set.
+Then we have $m$ decision trees which will grows independently with their own generated training sets.
+Finally, to predict the class it takes the average over all the trees.
+In order to find those parameters $m$ and the maximum depth of the tree, we have also used k-fold cross validation and the best values are $m = 200$ and maximum depth of $512$ for the binary classification and for the multi-class case $m = 500$ and maximum depth also of $512$.
 
@@
 \section{Evaluation methods}
+
+In order to evaluate our different models, we have split our data in two sets of $80\%$ and $20\%$ for the training and the test sets.
+We choose this percentage in order to have still enough data for the training set and also enough to estimate correctly our models on the testing set.
+We learn only on the training set and then use the test set in order to estimate the error using the \textit{Balanced Error Rate}.
+This error compute the average percentage of misclassification per class.
+It allows us to avoid the case of the $90\%/10\%$ and having a good error.
+We have also used k-fold cross validation to tune the parameters up of our different models.
+We set $k=10$ because we have a lot of data, splitting them into $10$ folds seems reasonable ($4320$ data samples for the training set and $480$ for the testing set).
+We repeated the experiment $10$ times with different seeds in order to split the data in a unique random way for each trial and so try to have an unbiased estimation error.
+We couldn't increase the number of experiments because some methods are really slow.
 

### 2015-12-20 Projet II/RF_CNN_PCA.m

@@ -8,7 +8,7 @@ addpath(genpath('DeepLearnToolbox'));
 load train/train.mat;
 
 ratio = 0.8;
-M = 200;
+M = 150;
 
 %% Create data
 fprintf('Creating Train & Test sets\n');
@@ -43,7 +43,7 @@ if binaryClassification
     yTe(find(yTe == 3)) = 1;
     yTe(find(yTe == 4)) = 2;
 end
-
+%%
     % Binary
     % 200 : 9.29% (nZ), 9.26% (Z)
     % 500 : 9.6% (nZ), 9.31% (Z)
@@ -54,16 +54,43 @@ end
     % 500 : 11.64% (nZ), 11.12% (Z)
     % 1000 : 11.08% (nZ), 10.90% (Z)
 
-    % Training using nZ
-    BaggedEnsemble = TreeBagger(200, Tr.nZ, yTr);
-    yhat = str2double(predict(BaggedEnsemble, Te.nZ));
+    maxDepths = [64 128 256 512 1024 2048 4096];
+    Ms = [256 512 768 1024];
+    F1s = [25 50 100 150];
+    
+    for i = 1:1:length(maxDepths)
+        for j = 1:1:length(Ms)
+            for k = 1:1:length(F1s)
+                md = maxDepths(i);
+                m = Ms(j);
+                f1 = F1s(k);
+                pTrain={'maxDepth',md,'M',m,'F1',f1,'minChild',5};
+                tic, forest=forestTrain(Tr.Z,yTr,pTrain{:}); toc
 
-    errnZ = balancedErrorRate(yTe, yhat);
-    fprintf('BER Testing error nZ: %.2f%%\n', errnZ * 100);
+                hsPr0 = forestApply(single(full(Te.Z)),forest);
+                err(i,j,k) = balancedErrorRate(hsPr0, yTe);
+               fprintf('%d %d %d \t\t\t\t\t%f\n', md, m, f1, err(i,j,k));
+            end
+        end
+    end
+%%
 
-    % Training using Z
-    BaggedEnsemble = TreeBagger(200, Tr.Z, yTr);
-    yhat = str2double(predict(BaggedEnsemble, Te.Z));
+    fernPrm=struct('S',15,'M', 4096, 'thrr',[-1 1],'bayes',1);
+    tic, [ferns,hsPr0]=fernsClfTrain(Tr.Z, yTr,fernPrm); toc
[…]

### 2015-12-20 Projet II/SVM_HOG_PCA.m

@@ -45,47 +45,20 @@ if binaryClassification
 end
 
     % Binary
-    % Default : 25.5 (X), 28.73 (nX), 26.08% (Z), 29.59 (nZ)
-    % Linear : ~29.96%
-    % RBF : ~23.81%
-    % Poly^7 : ~20.8%
+    % Default : 24.18% 
 
     % Multiclass
-    % Default : 29.59 (X), 30.46 (nX), 29.52% (Z) 31.05 (nZ)
-    % Linear : ~31.06%
-    % RBF : ~27.04%
-    % Poly^7 : ~25.6
+    % Default : 27.34% 
     
-    t = templateSVM();%'Solver', 'ISDA', 'KernelFunction', 'polynomial', 'PolynomialOrder', 7, 'BoxConstraint', Inf);
+     t = templateSVM('Solver', 'SMO', 'KernelFunction', 'linear', 'IterationLimit', 1e6, 'KernelScale', 1, 'BoxConstraint', 1);   
     
-%     % Training using X
-%     Mdl = fitcecoc(Tr.X, yTr, 'Learners', t, 'Options', statset('UseParallel', 1));
-%     CMdl = compact(discardSupportVectors(Mdl));
-%     yhat = predict(CMdl, Te.X);
-% 
-%     errX = balancedErrorRate(yTe, yhat);
-%     fprintf('BER Testing error  X: %.2f%%\n', errX * 100);
-% 
-%     % Training using nX
-%     Mdl = fitcecoc(Tr.nX, yTr, 'Learners', t, 'Options', statset('UseParallel', 1));
-%     CMdl = compact(discardSupportVectors(Mdl));
-%     yhat = predict(CMdl, Te.nX);
-% 
-%     errnX = balancedErrorRate(yTe, yhat);
-%     fprintf('BER Testing error nX: %.2f%%\n', errnX * 100);
-
-    % Training using Z
-    Mdl = fitcecoc(Tr.Z, yTr, 'Learners', t, 'Options', statset('UseParallel', 1));
+    % Training using Z 
+    tic
+    Mdl = fitcecoc(Tr.Z, yTr, 'Learners', t, 'Coding', 'onevsall', 'Options', statset('UseParallel', 1));
+    toc
     CMdl = compact(discardSupportVectors(Mdl));
     yhat = predict(CMdl, Te.Z);
 
     errZ = balancedErrorRate(yTe, yhat);
[…]

### 2015-12-20 Projet II/binaryClassification_2.m

@@ -1,0 +1,203 @@
+%% Initialization
+clear all;
+close all;
+
+addpath(genpath('Piotr'));
+addpath(genpath('DeepLearnToolbox'));
+
+load train/train.mat;
+
+numberOfExperiments = 10;
+proportionOfTraining = 0.8;
+K = 10;
+M = 1000;
+
+train.y(find(train.y == 2)) = 1;
+train.y(find(train.y == 3)) = 1;
+train.y(find(train.y == 4)) = 2;
+
+%% Create data
+fprintf('Creating Train & Test sets\n');
+tic
+[Tr, Te] = createTrainingTestingHOG(train.X_hog, train.y, 1.0);
+toc
+
+%% Prepare data
+
+fprintf('Prepare the data for the training\n');
+tic
+[Tr, Te] = prepareDataHOG(Tr, Te, M);
+toc
+
+%%
+% 50%
+%**********************************
+%            TEST 1
+%**********************************
+
+fprintf('Test 1\n')
+for i = 1:numberOfExperiments
+    fprintf('%d ', i);
+    setSeed(28111993*i);
+    out = 1;
+    err1(i) = 0.5;%balancedErrorRate(Te.y, repmat([out], length(Tr.y), 1));
+end
+fprintf('\n%f\n', mean(err1));
+saveFile(err1, 'results/binary/err1_');
+
+%% 23.68%
+%**********************************
+%            TEST 2
+%**********************************
+
+N = length(Tr.y);
+
+% Setup NN.
+inputSize  = M;
+innerSize  = 1000;
+numepochs = 20;
+batchsize = 100;
+learningRate = 3;
+binaryClassification = true;
+
+Tr_ = Tr;
+
+for j = 1:1:numberOfExperiments
+    setSeed(28111993*j);
+
+    fprintf('%d ', j);
+    [TTr, TTe] = splitProp(proportionOfTraining, Tr_, false);
+   
+    [errnZ, nnPrednZ] = neuralNetworks(TTr.nZ, TTr.y, TTe.nZ, TTe.y, inputSize, innerSize, numepochs, batchsize, learningRate, binaryClassification);
+
+    err2(j) = errnZ;
+end
+fprintf('\n%f\n', mean(err2));
+saveFile(err2, 'results/binary/err2_');
+
+%% 0.4 %
+%**********************************
+%            TEST 3
[…]

### 2015-12-20 Projet II/crossValidateModels_CNN_bin.m

@@ -92,7 +92,7 @@ end
 
 %% DT
 
-binaryClassification = true;
+binaryClassification = false;
 
 Tr_ = Tr;
 Te_ = Te;
@@ -110,6 +110,7 @@ end
 
 % 10.824 with 200
 NLeaves = [200];
+depths = [512];
 
 for j = 1:1:numberOfExperiments
     setSeed(28111993*j);
@@ -118,7 +119,7 @@ for j = 1:1:numberOfExperiments
     [TTr, TTe] = splitProp(proportionOfTraining, Tr_, false);
     
     idxCV = splitGetCV(K, length(TTr.y));
-    
+
     % K-fold
     for k=1:1:K
         fprintf('%d : %dth fold\n', j, k);
@@ -126,11 +127,14 @@ for j = 1:1:numberOfExperiments
         
         for i = 1:1:length(NLeaves)
             NLeave = NLeaves(i);
-            
-            CMdl = fitensemble(TTTr.Z, TTTr.y, 'Bag', NLeave, 'Tree', 'Type', 'classification');
-            yhat = predict(CMdl, TTTe.Z);
+            for i = 1:1:length(depths)
+                depth = depths(i);
+                t = templateTree('MaxNumSplits', depth);
+                CMdl = fitensemble(TTTr.Z, TTTr.y, 'Bag', NLeave, 'Tree', 'Type', 'classification', 'Learners', t);
+                yhat = predict(CMdl, TTTe.Z);
 
             err_te(k,i) = balancedErrorRate(TTTe.y, yhat);
+            end
         end
     end
 
@@ -260,4 +264,120 @@ for j = 1:1:numberOfExperiments
     
     [errStar, maxDepthStarId] = min(mseTe);
     maxDepthStar = maxDepths(maxDepthStarId);
+end
+
+%% RF-2
+
+Tr_ = Tr;
+Te_ = Te;
+
+binaryClassification = true;
+
+Tr_ = Tr;
+Te_ = Te;
+    
+if binaryClassification    
+    Tr_.y(find(Tr_.y == 2)) = 1;
+    Tr_.y(find(Tr_.y == 3)) = 1;
+    Tr_.y(find(Tr_.y == 4)) = 2;
+    
+    Te_.y(find(Te_.y == 2)) = 1;
+    Te_.y(find(Te_.y == 3)) = 1;
+    Te_.y(find(Te_.y == 4)) = 2;
+end
+
+
+% 9.61%
+maxDepths = [512];
+Ms = [256];
[…]

### 2015-12-20 Projet II/crossValidateModels_CNN_mul.m

@@ -164,3 +164,91 @@ for j = 1:1:numberOfExperiments
     [errStar, NTreeStarId] = min(mseTe);
     NTreeStar = NTrees(NLeaveStarId);
 end
+
+%% RF-2
+
+Tr_ = Tr;
+Te_ = Te;
+
+% 9.6%
+maxDepths = [1024];
+Ms = [256];
+F1s = [20];
+
+for jj = 1:1:numberOfExperiments
+    setSeed(28111993*jj);
+
+    fprintf('%d : Split the data\n', jj);
+    [TTr, TTe] = splitProp(proportionOfTraining, Tr_, false);
+    
+    idxCV = splitGetCV(K, length(TTr.y));
+    
+    % K-fold
+    for kk=1:1:K
+        fprintf('%d : %dth fold\n', j, kk);
+        [TTTr, TTTe] = splitGetTrTe(TTr, idxCV, kk, false);
+        
+        for i = 1:1:length(maxDepths)
+            for j = 1:1:length(Ms)
+                for k = 1:1:length(F1s)
+                    md = maxDepths(i);
+                    m = Ms(j);
+                    f1 = F1s(k);
+                    pTrain={'maxDepth',md,'M',m,'F1',f1,'minChild',5};
+                    forest=forestTrain(Tr.Z,yTr,pTrain{:});
+
+                    hsPr0 = forestApply(single(full(Te.Z)),forest);
+                    err(kk, i, j, k) = balancedErrorRate(hsPr0, yTe);
+                    fprintf('%d %d %d \t\t\t\t\t%f\n', md, m, f1, err(kk, i, j, k));
+                end
+            end
+        end
+    end
+    fprintf('%f ', mean(err));
+    mseTe = mean(err_te);
+
+    [errStar, NTreeStarId] = min(mseTe);
+    NTreeStar = NTrees(NLeaveStarId);
+end
+
+%% Fernst
+
+Tr_ = Tr;
+Te_ = Te;
+
+% 8.54 %
+Ss = [15];
+Ms = [4096];
+
+for jj = 1:1:numberOfExperiments
+    setSeed(28111993*jj);
+
+    fprintf('%d : Split the data\n', jj);
+    [TTr, TTe] = splitProp(proportionOfTraining, Tr_, false);
+    
+    idxCV = splitGetCV(K, length(TTr.y));
+    
+    % K-fold
+    for kk=1:1:K
+        fprintf('%d : %dth fold\n', j, kk);
[…]

### 2015-12-20 Projet II/multiClassification.m

@@ -89,8 +89,11 @@ N = length(Tr.y);
 
 % Setup 
 NLeaves = 500;
+depth = 512;
 Tr_ = Tr;
 Te_ = Te;
+
+t = templateTree('MaxNumSplits', depth);
 
 for j = 1:1:numberOfExperiments
     setSeed(28111993*j);
@@ -104,7 +107,7 @@ for j = 1:1:numberOfExperiments
     for k=1:1:K
         [TTTr, TTTe] = splitGetTrTe(TTr, idxCV, k, false);
         
-        CMdl = fitensemble(TTTr.Z, TTTr.y, 'Bag', NLeaves, 'Tree', 'Type', 'classification');
+        CMdl = fitensemble(TTTr.Z, TTTr.y, 'Bag', NLeaves, 'Tree', 'Type', 'classification', 'Learners', t);
         yhat = predict(CMdl, TTTe.Z);
 
         err_te(k) = balancedErrorRate(TTTe.y, yhat);

### 2015-12-20 Projet II/report/paper.tex
Add results

@@
 
+\subsection{Fernst}
+
 \section{Evaluation methods}
@@
 We set $k=10$ because we have a lot of data, splitting them into $10$ folds seems reasonable ($4320$ data samples for the training set and $480$ for the testing set).
-We repeated the experiment $10$ times with different seeds in order to split the data in a unique random way for each trial and so try to have an unbiased estimation error.
-We couldn't increase the number of experiments because some methods are really slow.
+We repeated the experiment $30$ times with different seeds in order to split the data in a unique random way for each trial and so try to have an unbiased estimation error.
 

### 2015-12-20 Projet II/binaryClassification.m

@@ -7,7 +7,7 @@ addpath(genpath('DeepLearnToolbox'));
 
 load train/train.mat;
 
-numberOfExperiments = 10;
+numberOfExperiments = 30;
 proportionOfTraining = 0.8;
 K = 10;
 M = 150;
@@ -19,7 +19,7 @@ train.y(find(train.y == 4)) = 2;
 %% Create data
 fprintf('Creating Train & Test sets\n');
 tic
-[Tr, Te] = createTrainingTestingCNN(train.X_cnn, train.y, proportionOfTraining);
+[Tr, Te] = createTrainingTestingCNN(train.X_cnn, train.y, 1.0);
 toc
 
 %% Prepare data
@@ -29,8 +29,7 @@ tic
 [Tr, Te] = prepareDataCNN(Tr, Te, M);
 toc
 
-%%
-% 50%
+%% Constant 50%
 %**********************************
 %            TEST 1
 %**********************************
@@ -40,12 +39,12 @@ for i = 1:numberOfExperiments
     fprintf('%d ', i);
     setSeed(28111993*i);
     out = 1;
-    err1(i) = balancedErrorRate(Te.y, repmat([out], length(Tr.y), 1));
+    err1(i) = 0.5;%balancedErrorRate(Te.y, repmat([out], length(Tr.y), 1));
 end
 fprintf('\n%f\n', mean(err1));
 saveFile(err1, 'results/binary/err1');
 
-%% 9.96%
+%% Neural Network 9.045 +- 0.009%
 %**********************************
 %            TEST 2
 %**********************************
@@ -60,148 +59,189 @@ batchsize = 100;
 learningRate = 3;
 binaryClassification = true;
 
-Tr_ = Tr;
-Te_ = Te;
-
 for j = 1:1:numberOfExperiments
     setSeed(28111993*j);
 
     fprintf('%d ', j);
-    [TTr, TTe] = splitProp(proportionOfTraining, Tr_, false);
+    [TTr, TTe] = splitProp(proportionOfTraining, Tr, false);
         
-    idxCV = splitGetCV(K, length(TTr.y));
-    
-    % K-fold
-    for k=1:1:K
-        [TTTr, TTTe] = splitGetTrTe(TTr, idxCV, k, false);
-        
[…]

### 2015-12-20 Projet II/binaryClassification_2.m

@@ -1,203 +1,0 @@
-%% Initialization
-clear all;
-close all;
-
-addpath(genpath('Piotr'));
-addpath(genpath('DeepLearnToolbox'));
-
-load train/train.mat;
-
-numberOfExperiments = 10;
-proportionOfTraining = 0.8;
-K = 10;
-M = 1000;
-
-train.y(find(train.y == 2)) = 1;
-train.y(find(train.y == 3)) = 1;
-train.y(find(train.y == 4)) = 2;
-
-%% Create data
-fprintf('Creating Train & Test sets\n');
-tic
-[Tr, Te] = createTrainingTestingHOG(train.X_hog, train.y, 1.0);
-toc
-
-%% Prepare data
-
-fprintf('Prepare the data for the training\n');
-tic
-[Tr, Te] = prepareDataHOG(Tr, Te, M);
-toc
-
-%%
-% 50%
-%**********************************
-%            TEST 1
-%**********************************
-
-fprintf('Test 1\n')
-for i = 1:numberOfExperiments
-    fprintf('%d ', i);
-    setSeed(28111993*i);
-    out = 1;
-    err1(i) = 0.5;%balancedErrorRate(Te.y, repmat([out], length(Tr.y), 1));
-end
-fprintf('\n%f\n', mean(err1));
-saveFile(err1, 'results/binary/err1_');
-
-%% 23.68%
-%**********************************
-%            TEST 2
-%**********************************
-
-N = length(Tr.y);
-
-% Setup NN.
-inputSize  = M;
-innerSize  = 1000;
-numepochs = 20;
-batchsize = 100;
-learningRate = 3;
-binaryClassification = true;
-
-Tr_ = Tr;
-
-for j = 1:1:numberOfExperiments
-    setSeed(28111993*j);
-
-    fprintf('%d ', j);
-    [TTr, TTe] = splitProp(proportionOfTraining, Tr_, false);
-   
-    [errnZ, nnPrednZ] = neuralNetworks(TTr.nZ, TTr.y, TTe.nZ, TTe.y, inputSize, innerSize, numepochs, batchsize, learningRate, binaryClassification);
-
-    err2(j) = errnZ;
-end
-fprintf('\n%f\n', mean(err2));
-saveFile(err2, 'results/binary/err2_');
-
-%% 0.4 %
-%**********************************
-%            TEST 3
[…]

### 2015-12-20 Projet II/crossValidateModels_CNN_bin.m

@@ -16,7 +16,7 @@ M = 150;
 %% Create data
 fprintf('Creating Train & Test sets\n');
 tic
-[Tr, Te] = createTrainingTestingCNN(train.X_cnn, train.y, proportionOfTraining);
+[Tr, Te] = createTrainingTestingCNN(train.X_cnn, train.y, 1.0);
 toc
 
 %% Prepare data
@@ -268,35 +268,16 @@ end
 
 %% RF-2
 
-Tr_ = Tr;
-Te_ = Te;
-
-binaryClassification = true;
-
-Tr_ = Tr;
-Te_ = Te;
-    
-if binaryClassification    
-    Tr_.y(find(Tr_.y == 2)) = 1;
-    Tr_.y(find(Tr_.y == 3)) = 1;
-    Tr_.y(find(Tr_.y == 4)) = 2;
-    
-    Te_.y(find(Te_.y == 2)) = 1;
-    Te_.y(find(Te_.y == 3)) = 1;
-    Te_.y(find(Te_.y == 4)) = 2;
-end
-
-
 % 9.61%
-maxDepths = [512];
+maxDepths = [1024*2];
 Ms = [256];
-F1s = [20];
+F1s = [15];
 
 for jj = 1:1:numberOfExperiments
     setSeed(28111993*jj);
 
     fprintf('%d : Split the data\n', jj);
-    [TTr, TTe] = splitProp(proportionOfTraining, Tr_, false);
+    [TTr, TTe] = splitProp(proportionOfTraining, Tr, false);
     
     idxCV = splitGetCV(K, length(TTr.y));
     
@@ -312,11 +293,11 @@ for jj = 1:1:numberOfExperiments
                     m = Ms(j);
                     f1 = F1s(k);
                     pTrain={'maxDepth',md,'M',m,'F1',f1,'minChild',5};
-                    forest=forestTrain(Tr.Z,yTr,pTrain{:});
+                    forest=forestTrain(TTTr.Z, TTTr.y,pTrain{:});
 
-                    hsPr0 = forestApply(single(full(Te.Z)),forest);
-                    err(kk, i, j) = balancedErrorRate(hsPr0, yTe);
-                    fprintf('%d %d %d \t\t\t\t\t%f\n', md, m, f1, err(kk, i, j));
+                    hsPr0 = forestApply(single(full(TTTe.Z)),forest);
+                    err(kk) = balancedErrorRate(hsPr0, TTTe.y);
[…]

### 2015-12-20 Projet II/multiClassification.m

@@ -7,7 +7,7 @@ addpath(genpath('DeepLearnToolbox'));
 
 load train/train.mat;
 
-numberOfExperiments = 10;
+numberOfExperiments = 30;
 proportionOfTraining = 0.8;
 K = 10;
 M = 150;
@@ -15,7 +15,7 @@ M = 150;
 %% Create data
 fprintf('Creating Train & Test sets\n');
 tic
-[Tr, Te] = createTrainingTestingCNN(train.X_cnn, train.y, proportionOfTraining);
+[Tr, Te] = createTrainingTestingCNN(train.X_cnn, train.y, 1.0);
 toc
 
 %% Prepare data
@@ -25,8 +25,7 @@ tic
 [Tr, Te] = prepareDataCNN(Tr, Te, M);
 toc
 
-%%
-% 75%
+%% Constant 75%
 %**********************************
 %            TEST 1
 %**********************************
@@ -36,17 +35,15 @@ for i = 1:numberOfExperiments
     fprintf('%d ', i);
     setSeed(28111993*i);
     out = 1;
-    err1(i) = balancedErrorRate(Te.y, repmat([out], length(Tr.y), 1));
+    err1(i) = 0.75;%balancedErrorRate(Te.y, repmat([out], length(Tr.y), 1));
 end
-fprintf('\n%f\n', mean(err1));
+fprintf('\n%f +- 0\n', mean(err1));
 saveFile(err1, 'results/multi/err1');
 
-%% 9.14 %
+%% Neural Network 8.646 +- 0.008 %
 %**********************************
 %            TEST 2
 %**********************************
-
-N = length(Tr.y);
 
 % Setup NN.
 inputSize  = M;
@@ -56,42 +53,27 @@ batchsize = 100;
 learningRate = 2;
 binaryClassification = false;
 
-Tr_ = Tr;
-Te_ = Te;
-
 for j = 1:1:numberOfExperiments
     setSeed(28111993*j);
 
     fprintf('%d ', j);
-    [TTr, TTe] = splitProp(proportionOfTraining, Tr_, false);
+    [TTr, TTe] = splitProp(proportionOfTraining, Tr, false);
         
-    idxCV = splitGetCV(K, length(TTr.y));
-    
-    % K-fold
-    for k=1:1:K
-        [TTTr, TTTe] = splitGetTrTe(TTr, idxCV, k, false);
-        
[…]

### 2015-12-20 Projet II/report/paper.tex
Add last results

@@
 
-\subsection{Bootstrap Aggregation}
+\subsection{Bootstrap Aggregation (Bagging)}
 
-As we have seen in the lecture, decision trees are good because the training and the prediction is quite fast, even with many features or large dataset but have a high variance.
+As we have seen in the lecture, decision trees are good because the training and the prediction is quite fast, even with many features or large dataset but have a high variance and might overfit.
 The idea of decision trees, is to have a several questions where to answer yes/no and to move in the tree.
@@
 \subsection{Random Forests}
+
+The idea of random forests is similar than with bagging in a sense that we train several trees and take the average predictions.
+Like bagging, each trees have trained on a subset of the dataset.
+However, there is another aspect in random forest that bagging doesn't have: for each split, instead of searching among the overall features, we search only among a random subset of features.
+The size of this subset is a parameter $m$ as well as the depth tree and the number of trees, like in bagging.
+Like bagging, random forests are useful because they help reducing the variance and the overfiting.
+But the more interesting ideas is to select only a subset of the features in the split process.
 

### 2015-12-20 Projet II/binaryClassification.m

@@ -44,7 +44,7 @@ end
 fprintf('\n%f\n', mean(err1));
 saveFile(err1, 'results/binary/err1');
 
-%% Neural Network 9.045 +- 0.009%
+%% Neural Network 0.0905 +- 0.009%
 %**********************************
 %            TEST 2
 %**********************************
@@ -72,7 +72,7 @@ end
 fprintf('\n%f +- %f\n', mean(err2), std(err2));
 saveFile(err2, 'results/binary/err2');
 
-%% bootstrap aggregation 9.42 +- 0.007%
+%% bootstrap aggregation 0.0942 +- 0.007%
 %**********************************
 %            TEST 3
 %**********************************
@@ -123,7 +123,7 @@ end
 fprintf('\n%f +- %f\n', mean(err4), std(err4));
 saveFile(err4, 'results/binary/err4');
 
-%% AdaBoost Tree 7.992 +- 0.008%
+%% AdaBoost Tree 0.799 +- 0.008%
 %**********************************
 %            TEST 5
 %**********************************
@@ -161,7 +161,7 @@ end
 fprintf('\n%f +- %f\n', mean(err5), std(err5));
 saveFile(err5, 'results/binary/err5');
 
-%% Random forest 12.21 +- 0.010%
+%% Random forest 0.122 +- 0.010%
 %**********************************
 %            TEST 6
 %**********************************
@@ -217,7 +217,7 @@ end
 fprintf('\n%f +- %f\n', mean(err7), std(err7));
 saveFile(err7, 'results/binary/err7');
 
-%% SVM TODO 75'
+%% SVM 0.886 +- 0.006
 %**********************************
 %            TEST 8
 %**********************************
@@ -253,9 +253,10 @@ err4 = openFile('results/binary/err4', s);
 err5 = openFile('results/binary/err5', s);
 err6 = openFile('results/binary/err6', s);
 err7 = openFile('results/binary/err7', s);
+err8 = openFile('results/binary/err8', s);
 
 figure;
-boxplot([err1' err2' err3' err4' err5' err6' err7']);
+boxplot([err1' err2' err3' err4' err5' err6' err7' err8']);
 h_legend = legend(findobj(gca,'Tag','Box'), ...
 '1 ', ...
 '2 ', ...
@@ -263,7 +264,8 @@ h_legend = legend(findobj(gca,'Tag','Box'), ...
[…]

### 2015-12-21 Projet II/report/paper.tex
Advance report

@@
 
-We also have test-data of size $N=XXX$ without their corresponding output.
+We also have test-data of size $N=11453$ without their corresponding output.
 Our goal is to produce predictions for those data, as well as an approximation of the test-error using \textit{Balanced Error Rate} (BER).
@@
 For the binary case, we have $60.3\%$ for the label $1$ and $39.8\%$ for the second one.
-We have observed that there is a small imbalance between the classesl.
+We have observed that there is a small imbalance between the classes.
 For the multi-classes case, we obtain $16.07\%$, $19.37\%$, $24.87\%$ and $39.70\%$ for the labels $1$, $2$, $3$ and $4$.
@@
 
+\section{Evaluation methods}
+
+In order to evaluate our different models, we have split our data in two sets of $80\%$ and $20\%$ for the training and the test sets.
+We choose this percentage in order to have still enough data for the training set and also enough to estimate correctly our models on the testing set.
+We learn only on the training set and then use the test set in order to estimate the error using the \textit{Balanced Error Rate}.
+This error compute the average percentage of misclassification per class.
+It allows us to avoid the case of the $90\%/10\%$ and having a good error.
+We have also used k-fold cross validation to tune the parameters up of our different models.
+We set $k=10$ because we have a lot of data, splitting them into $10$ folds seems reasonable ($4320$ data samples for the training set and $480$ for the testing set).
+We repeated the experiment $30$ times with different seeds in order to split the data in a unique random way for each trial and so try to have an unbiased estimation error.
+
 \section{Techniques used}
 
-This section is a description of the technique we have used in the project. It describes mainly how it works, why we thought it could be useful in this project.
+This section is a description of the techniques we have used in the project. It describes mainly how it works, why we thought it could be useful in this project.
+
+\begin{figure}[!ht]
+\center
+\subfigure[Plot of the distortion $J$ to $M$ where $M = D$. The distortion decreases quickly.]{\includegraphics[width=2.7in]{figures/distortionHOG.jpg} \label{fig:distHOG}}
+\hfill
+\subfigure[Plot of the distortion $J$ to $M$ where $M = N$ because there are too many features and so, the distortion doesn't go to $0$ and decreases slowly.]{\includegraphics[width=2.7in]{figures/distortionCNN.jpg}\label{fig:distCNN}}
+\hfill
+\caption{}
+\end{figure}
 
@@
 
-The Figure XX shows the plot of $J$ for the HOG features and the CNN features.
-For the CNN, we have tried several values for $M$ and have conclude that $M_{CNN} = 150$ was a good value allowing us to compute faster results without loosing too much accuracy.
+The Figure \ref{fig:distHOG} shows the plot of $J$ for the HOG features and the CNN features.
+Figure \ref{fig:distCNN} shows the distortion of the CNN features, we have tried several values for $M$ and have conclude that $M_{CNN} = 150$ was a good value allowing us to compute faster results without loosing too much accuracy.
 The corresponding distortion might be very high compared to the HOG features, but the range of the values are not the same, CNN has values going larger than $0.2$.
@@
 We have put this value to $20$, in order to have a little more accuracy without loosing a lot of time, it was a good compromise because larger epochs don't mean especially better accuracy.
-the number of input is $M$ and output is either $2$ or $4$, depending of the type of classification (binary or multi-class).
+The number of input is $M$ and output is either $2$ or $4$, depending of the type of classification (binary or multi-class).
 For the number of hidden layer, we have let $1$ because we didn't see improvement by adding some more.
-However, the number of neurons in the hidden layer has been tested and have found $1000$ for the binary classification and $700$ for the multi-class case.
+However, the number of neurons in the hidden layer has been tested with k-fold cross validation and we have found $1000$ for the binary classification and $700$ for the multi-class case.
 We also have to set up the learning rate : the optimal value we have found is $3$ for the binary case and $2$ for the multi-class one.
@@
 In order to find those parameters $m$ and the maximum depth of the tree, we have also used k-fold cross validation and the best values are $m = 200$ and maximum depth of $512$ for the binary classification and for the multi-class case $m = 500$ and maximum depth also of $512$.
+Having a high maximum depth might lead to overfit because we specialize to much the features split, but if it is too small, it might underfit because the features are not well split in order to identify classes.
+For the parameter $m$, it is better to have a higher number in order to have a large number of trees which can give a better accuracy of the prediction.
 
@@
 However, there is another aspect in random forest that bagging doesn't have: for each split, instead of searching among the overall features, we search only among a random subset of features.
-The size of this subset is a parameter $m$ as well as the depth tree and the number of trees, like in bagging.
+The size of this subset is a parameter $m$ as well as the maximum depth tree and the number of trees, like in bagging.
 Like bagging, random forests are useful because they help reducing the variance and the overfiting.
 But the more interesting ideas is to select only a subset of the features in the split process.
+Moreover, it is efficient against imbalanced dataset.
+
+As in bagging trees, the number of trees and the maximum depth tree behaves in the same manner.
+The last parameter $m$ has an important impact because it is too small, the subset of features might not describe correctly the data, and if it is too high, this subset might be to high and might not represent well the data.
+We have to find a good compromise in order to have a certain number of features to describe the data.
+Still with k fold cross validation, we have found for the multi-class case a maximum depth of $1024$,  $64$ trees and a $m = 20$.
+For $m$, an heuristic could be the square root of the number of features, but it was less efficient than $20$ in our case.
+For the binary case, those parameters are $1024$, $256$ and $50$.
 
 \subsection{Fernst}
-
-\section{Evaluation methods}
-
-In order to evaluate our different models, we have split our data in two sets of $80\%$ and $20\%$ for the training and the test sets.
-We choose this percentage in order to have still enough data for the training set and also enough to estimate correctly our models on the testing set.
-We learn only on the training set and then use the test set in order to estimate the error using the \textit{Balanced Error Rate}.
-This error compute the average percentage of misclassification per class.
-It allows us to avoid the case of the $90\%/10\%$ and having a good error.
-We have also used k-fold cross validation to tune the parameters up of our different models.
-We set $k=10$ because we have a lot of data, splitting them into $10$ folds seems reasonable ($4320$ data samples for the training set and $480$ for the testing set).
-We repeated the experiment $30$ times with different seeds in order to split the data in a unique random way for each trial and so try to have an unbiased estimation error.
 
@@
 \section{References}
+
+@misc{DymolaWeb,
+  title = "xxx Website",
+  author = "xxx",
+  howpublished = "\url{http://www.xxx.com}",
+  year = "cited July 2008"
+}
 

### 2015-12-21 Projet II/SVM_CNN_PCA.m

@@ -13,7 +13,7 @@ M = 200;
 %% Create data
 fprintf('Creating Train & Test sets\n');
 tic
-[Tr, Te] = createTrainingTestingCNN(train.X_cnn, train.y, ratio);
+[Tr, Te] = createTrainingTestingCNN(train.X_cnn, train.y, 1.0);
 toc
 
 %% Prepare data

### 2015-12-21 Projet II/prepareDataCNN.m

@@ -30,8 +30,8 @@ function [Tr, Te] = PCA(Tr_, Te_, M)
     norm = repmat(norm, D, 1);
     U = X2' * V ./ norm;
     
-    %J = computeDistortionMatrix(l, N);
-    %displayDistortionMatrix(J);
+    J = computeDistortionMatrix(l, N);
+    displayDistortionMatrix(J);
     
     Tr.nZ = Tr.nX * U(:, 1:M);
     Te.nZ = Te.nX * U(:, 1:M);
@@ -51,7 +51,19 @@ end
 function [] = displayDistortionMatrix(J)
     figure('Name', 'Distortion Measure');
     plot(J, 'LineWidth', 4);
+    ylim([0 4e4]);
+    xlim([0 size(J,1)]);
     xlabel('M');
     ylabel('J');
     title('distortion J vs M');
+   
+    set(gca,'fontsize', 18);
+    set(gca,'LineWidth',2);
+    SP=1.17;
+    line([SP SP], [0 4e4],'Color',[0 0 0], 'LineWidth', 2);
+    legend('Distortion');
+    hold off;
+    % saveas(f2, 'plots/cluster/feature16.jpg');
+    print('./report/figures/distortionCNN.jpg','-djpeg','-noui');
+
 end

### 2015-12-21 Projet II/prepareDataHOG.m

@@ -37,7 +37,17 @@ end
 function [] = displayDistortionMatrix(J)
     figure('Name', 'Distortion Measure');
     plot(J, 'LineWidth', 4);
+    ylim([0 5e3]);
+    xlim([0 N]);
     xlabel('M');
     ylabel('J');
     title('distortion J vs M');
+   
+    set(gca,'fontsize', 18);
+    set(gca,'LineWidth',2);
+    SP=1.17;
+    line([SP SP], [0 5e3],'Color',[0 0 0], 'LineWidth', 2);
+    legend('Distortion');
+    hold off;
+    print('./report/figures/distortionHOG.jpg','-djpeg','-noui');
 end

### 2015-12-21 Projet II/report/paper.tex
Add binary classification analysis

@@
 This report provides a summary of the project two of the PCML class.
+The project consists of doing binary and multi-class classification on the same set of images.
+For both methods, we have observed that the given CNN features were more accurate than the HOG one.
+In binary classification, all methods were approximately equivalent except the Adaboosting and the SVM one.
+For the multi-class classification, neural networks and SVM were the best methods.
+
 \end{abstract}
@@
 
-We also wanted to analyze the distribution of our features, and have observed that for the features HOG, each of them has a has similar distribution as showed in Figure XXX.
+We also wanted to analyze the distribution of our features, and have observed that for the features HOG, each of them has a has similar distribution.
 The values are essentially in the range $[0, 0.2]$.
@@
 Finally,  $\mathbf{X}_{cnn}$ is rank-deficient with a rank of $5997$ instead of $36865$.
-
-\section{Evaluation methods}
-
-In order to evaluate our different models, we have split our data in two sets of $80\%$ and $20\%$ for the training and the test sets.
-We choose this percentage in order to have still enough data for the training set and also enough to estimate correctly our models on the testing set.
-We learn only on the training set and then use the test set in order to estimate the error using the \textit{Balanced Error Rate}.
-This error compute the average percentage of misclassification per class.
-It allows us to avoid the case of the $90\%/10\%$ and having a good error.
-We have also used k-fold cross validation to tune the parameters up of our different models.
-We set $k=10$ because we have a lot of data, splitting them into $10$ folds seems reasonable ($4320$ data samples for the training set and $480$ for the testing set).
-We repeated the experiment $30$ times with different seeds in order to split the data in a unique random way for each trial and so try to have an unbiased estimation error.
 
@@
 This is the multi-layer perceptron (MLP) we have seen in class.
-The neural networks used in this project comes from the given DeepLearning matlab toolbox.
+The neural networks used in this project comes from the given DeepLearning matlab toolbox [REF].
 It uses batch sizes in order to use less memory (take the first $n$ data samples and works on them, then the next $n$ etc).
@@
 Only the linear kernel function was available in this mode, and so, gaussian kernel might give better results.
+Even if the method is linear, SVM only classify directly the dataset in two classes, and so it is not a problem with the multi-class case because we always classify data in one of the two classes for the current classifier (see \textit{OVA} below).
 
@@
 
+We have used two different methods in matlab, one using \textit{fitensemble} with the parameter 'Bag' and another one which is the \textit{TreeBagger}.
+
 \subsection{Random Forests}
@@
 
-\subsection{Fernst}
+The given Piotr matlab toolboox [REF] contains a method \textit{forestTrain} which takes the defined parameters and \textit{forestApply} for the prediction.
+
+\subsection{Adaboosting}
+
+Adaboosting is another way to average trees.
+If we observe more deeply the bagging approach, some trees might have a very high variance and some a very slow.
+Averaging all those trees decrease the variance and have a small bias.
+Boosting is the opposite : decrease the bias when averaging and have a small variance.
+In order to achieve this, each data sample has an assigned non-negative weight.
+After each iteration, if the data sample is misclassified, its weight is increased for the next iteration.
+So misclassified data samples will have more importance.
+The parameters are the same as in the bagging and behave in the same manner.
+The Piotr matlab toolboox [REF] contains a method \textit{adaBoostTrain} and \textit{adaBoostApply} for the prediction.
+We have only used it for the binary classification, because the implementation is not adapted for more classes.
+The parameters we obtain with k fold cross validation are $1024$ for the number of trees and $4$ for the maximum depth.
+
+\section{Evaluation methods}
+
+In order to evaluate our different models, we have split our data in two sets of $80\%$ and $20\%$ for the training and the test sets.
+We choose this percentage in order to have still enough data for the training set and also enough to estimate correctly our models on the testing set.
+We learn only on the training set and then use the test set in order to estimate the error using the \textit{Balanced Error Rate}.
+This error compute the average percentage of misclassification per class.
+It allows us to avoid the case of the $90\%/10\%$ and having a good error.
+We have also used k-fold cross validation to tune the parameters up of our different models.
+We set $k=10$ because we have a lot of data, splitting them into $10$ folds seems reasonable ($4320$ data samples for the training set and $480$ for the testing set).
+We repeated the experiment $30$ times with different seeds in order to split the data in a unique random way for each trial and so try to have an unbiased estimation error.
 
@@
 
+This section describes our models for the binary classification as well as the multi-class classification.
+We have essentially put our effort on the CNN features.
+The reason is because with only the HOG features, we couldn't get an error lower than $20\%$ for binary and multi-class classification.
+With the CNN features, we didn't have this problem.
+We have tried to take the whole matrix, only a subset of features with $M = 500$, different methods but nothing worked.
+Contrary to the CNN features, we think that the reason is because those features are too noisy : they came directly form the images.
+Some of them are rotated, inversed and so, the gradient is also transformed, which might be confusing for the learning.
+The CNN features are already computed from a convolutional neural networks which gives the features it has seen on the image with a specific weight.
+Because each sample has approximately $5\%$ of the overall features, they are fewer and are more specific than the HOG features.
+
 \subsection{Binary classification}
 
+\begin{figure}
+\center
+\includegraphics[width=5in]{figures/binaryclassifications.jpg} 
+\caption{Boxplot of our models for the binary classification, using the BER. Each model is represented with a box where we can see their mean and their standard deviation.}
+\label{fig:bin_models}
+\end{figure}
+
+The Figure \ref{fig:bin_models} represents the BER error for each models.
+The first model used simply returns the label $1$ to all inputs, which will allow us to compare improvements with other models.
+It gives us an error of $0.5$, which is obvious because all the input are in the same class and the other has none of them.
+The second model uses neural networks with normalized data because we got better results and because it can avoid being stuck in a local minima and might speed the training.
+We have used it It clearly outperforms the baseline.
+We can see that it has a bit of variance but has a low bias.
+Increasing the number of hidden layer might accentuate the variance and diminish the bias.
+
+Our third and fourth model uses bootstrap aggregation with different implementations.
+They have similar performance than neural networks.
+In the first one, we can see that the variance is lower but there are some outliers.
+The other one has a bigger variance but is more robust to outliers.
+Let's remind than bagging averages trees and some trees might have big and small variance, which reduce variance in average.
+In general, it reduces the variance and increase slightly the bias.
+Compared to the neural network, the variance is better but the bias surely is worst.
+
+The next model ($5$) is the boosting method.
+As in bagger, it reduces the variance because it takes the average and the bias should be lower, which is the contrary compared to the bagger.
+We can also see that there are some outliers, but they are still better than the other seen methods.
+We think that the main reason is the variance and seems to be higher than the other models.
+In general, boosting should be better than bagger.
+
+Random forests is used for the sixth model.
+Surprisingly, we obtain worse results than the other methods.
+The k fold cross validation gave us better results when we were looking for the parameters.
+A possible cause of this effect is that it has overfit during the tuning process and so, gives worse performance in the real case.
+The parameter concerned should be the maximum depth which should be too high for our data because it specializes too much the features during the split.
+In theory, it should have better performance than the bagging and have less variances because we split only on a subset of features, however, the bias should be higher.
+Beside this, maybe the bias is too high and in this case, increasing the subset of feature might help to fix this.
+
+Our last model used the SVM.
+Compared to the other methods, we can observe that it is quite stable and have a low variance.
+If the parameter $C$ is set optimally, we should have a good variance and a good bias.
+However, if it is too high, the variance should be small but the bias could be high.
+In the other side, if $C$ is small, the bias should be lower but we have fewer support vectors which leads to a higher variance.
+However, let's remind that we use the same as in the multi-class classifier, using a linear kernel due to the implementation of matlab.
+If we could have changed the kernel into a non-linear one, the decision boundary would be different and might surely reduce the error.
+However, we would be faced to $\gamma$ which might change the variance and the bias.
+
+Our final model will be the Adaboosting one, because we considered it has the smallest error, even if there are some outliers, those are still lower than others methods in general.
+
 \subsection{Multi-class classification}
+
+\begin{figure}
+\center
+\includegraphics[width=5in]{figures/multiclassifications.jpg} 
+\caption{Boxplot of our models for the multi-class classification, using the BER. Each model is represented with a box where we can see their mean and their standard deviation.}
+\label{fig:models}
+\end{figure}
 

### 2015-12-21 Projet II/exploratoryDataAnalysis.m

@@ -10,10 +10,10 @@ y = train.y;
 X_cnn = train.X_cnn;
 X_hog = train.X_hog;
 
-y(find(y == 1)) = 1;
-y(find(y == 2)) = 1;
-y(find(y == 3)) = 1;
-y(find(y == 4)) = 2;
+% y(find(y == 1)) = 1;
+% y(find(y == 2)) = 1;
+% y(find(y == 3)) = 1;
+% y(find(y == 4)) = 2;
 
 %% Classes
 
@@ -46,10 +46,15 @@ corr_cnn = corr(double(X_cnn), double(y)); %[-0.2393;0.1795]
 %% Histogram
 
 f1 = figure;
-histogram(y);
+histogram(y, 'Normalization','probability');
 xlabel('Y');
 ylabel('Count');
-xlim([1 2]);
+xlim([1 4]); 
+set(gca,'fontsize', 18);
+legend('1', '2');
+hold off;
+% saveas(f2, 'plots/cluster/feature16.jpg');
+%print('./report/figures/distortionCNN.jpg','-djpeg','-noui');
 
 %% Histogram feature HOG (distribution)
 

### 2015-12-21 Projet II/report/paper.tex
Advance report

@@
 The second model uses neural networks with normalized data because we got better results and because it can avoid being stuck in a local minima and might speed the training.
-We have used it It clearly outperforms the baseline.
+It clearly outperforms the baseline.
 We can see that it has a bit of variance but has a low bias.
 Increasing the number of hidden layer might accentuate the variance and diminish the bias.
+For all other models, we didn't get improvements using normalization.
 
@@
 
-Our last model used the SVM.
+Our last model used the SVM as the one used for the multi-class classification.
+It works because the SVM OVA can be adapted easily to take into account only 1 binary classifier ($\frac{2*(2-1)}{2}=1$).
 Compared to the other methods, we can observe that it is quite stable and have a low variance.
@@
 
+We wanted to test some others models but we didn't get time to do it.
+First, we wanted to see if using a multi-class classifier which have a smallest error and then transform the prediction into a binary case would increase the accuracy.
+We think that the error might be improved but in general, it should stay more or less the same.
+Another promising method we wanted to try is the \textit{fernst} one.
+It should outperforms other decision-based trees because they compute posteriors multiplicatively [REF].
+We obtain good results with the Adaboosting method and think that fernst should produce better results.
+
 \subsection{Multi-class classification}
@@
 \caption{Boxplot of our models for the multi-class classification, using the BER. Each model is represented with a box where we can see their mean and their standard deviation.}
-\label{fig:models}
+\label{fig:mul_models}
 \end{figure}
 
-\section{Feature transformation}
+The Figure \ref{fig:mul_models} represents the BER error for each models.
+A general observation we can see is that the error of the models are similarly to the binary classification, which might mean that they both misclassify a shared subset of misclassified data samples.
+Another observation we might see is that in general, the variance is lower in the multi-class models.
+We think that the reason is that we classify explicit objects (car, horse, plane and others) without mixing them, which might be less noisy.
+
+The first model used simply returns the label $1$ to all inputs, which will allow us to compare improvements with other models.
+It gives us an error of $0.75$, which is obvious because all the input are in the same class and the others has none of them.
+The second model uses neural networks with normalized data as explained in the binary section.
+We can see that even if we have more label, neural networks are still efficient and has a low bias and a little bit of variance, because there is one outlier.
+Obviously, neural network are a lot more efficient than our baseline.
+
+Our third and fourth model uses bootstrap aggregation with different implementations.
+They are less good than neural networks.
+We think that the reason is because we have more labels, the set of splits to decide in which class is a data sample is more precised and so, a little bit less accurate.
+In the first one, we see that the variance is lower and there is no outlier comparing to the other bagger model.
+Those results compared to the binary classification are inverted, we might think that the implementation of those methods in matlab could be more efficient for a binary classification or multi-class classification depending of the method.
+
+The next model ($5$) is the SVM one using the OVA method.
+As we said earlier, it is several binary classifiers combined together.
+We obtain similar results than in the binary classification because the nature of the problem is very similar : the only difference is the number of labels.
+Moreover, it seems to have lower variance, even if there is one outlier.
+As highlighted before, changing the kernel function might lead to decrease the error and could be worthwhile to implement this version, however we didn't have time to do it.
+
+The last model uses random forests. As before, we obtain worse results than the other methods. It has greater bias than the other.
+We also think that the method has overfit when using k fold cross validation in order to find optimal parameters.
+The parameters to be changed could be the maximum depth tree because if it is too high, the split will be very specialized.
+We also think that the bias could be high and so, increasing the parameter $m$ could be worthwhile because first, we may need more features to classify a data sample in one of the four classes.
+Secondly, because the bias is surely higher than other method as well as the variance.
+More training samples should help to decrease the variance and to tune the parameters up.
+
+Our final model is neutal networks and not SVM.
+Their error is nearly the same, however, the variance of the neural networks is much lower than the one of SVM.
+Even if both have one outlier and surely have the same stability.
+
+Beside trying \textit{fernst} method as proposed in the binary case, we wanted to try to use several multi-class classifiers and then combine their results to see which data samples were more difficult and find a way to class them with more accuracy.
+One of the method we wanted to use was simply voting system in a similar manner as SVM OVA.
+We think that it could help to decrease the error.
 
 \section{Conclusion}
+
+We have seen that advanced machine learning methods are efficient to solve image classification problems.
+The errors obtained with binary classification and multi-class classification are closed and prove that the nature of the problem is similar.
+We have tried different methods and have remarked that the neural networks and the SVM models were better than the other in general.
+In both case, both produces good results.
+We really think that there is a way of improving SVM model by changing the kernel function because data are surely not separate linearly.
+At the end, we think that the best way of improving our binary and multi-class classifier would be to find a way to combine the HOG features with the CNN ones.
 

### 2015-12-21 Projet II/report/paper.tex
Add predictions

@@
 For both methods, we have observed that the given CNN features were more accurate than the HOG one.
-In binary classification, all methods were approximately equivalent except the Adaboosting and the SVM one.
-For the multi-class classification, neural networks and SVM were the best methods.
+In binary classification, all methods were approximately equivalent except the Adaboosting and the SVM one, but we choose the SVM method as our final model.
+For the multi-class classification, neural networks and SVM were the best methods but neural networks has less variance in this case and is considered as our final model.
+The best way of improving our models would be to find a way to combine HOG features with the CNN one.
 
@@
 
-Our final model will be the Adaboosting one, because we considered it has the smallest error, even if there are some outliers, those are still lower than others methods in general.
+Our final model will be the SVM one, because we considered it has one of the smallest error, but it doesn't have any outliers and is more stable than the Adaboosting.
 
@@
 
-Our final model is neutal networks and not SVM.
+Our final model is neural networks and not SVM.
 Their error is nearly the same, however, the variance of the neural networks is much lower than the one of SVM.
@@
 The errors obtained with binary classification and multi-class classification are closed and prove that the nature of the problem is similar.
-We have tried different methods and have remarked that the neural networks and the SVM models were better than the other in general.
-In both case, both produces good results.
-We really think that there is a way of improving SVM model by changing the kernel function because data are surely not separate linearly.
+We have tried different methods and have remarked that SVM model was better than the other in general.
+In both case, it produces good results and has a small variance.
+We really think that there is a way of improving it by changing the kernel function because data are surely not separate linearly.
 At the end, we think that the best way of improving our binary and multi-class classifier would be to find a way to combine the HOG features with the CNN ones.
@@
 
-@misc{DymolaWeb,
-  title = "xxx Website",
-  author = "xxx",
-  howpublished = "\url{http://www.xxx.com}",
-  year = "cited July 2008"
-}
+[1]    IMM2012-06284, R. B. Palm, Prediction as a candidate for learning deep hierarchical models of data, 2012.
 
-@misc{IMM2012-06284,
-    author       = "R. B. Palm",
-    title        = "Prediction as a candidate for learning deep hierarchical models of data",
-    year         = "2012",
-}
-
-@misc{PMT,
-   author = {Piotr Doll\'ar}, 
-   title = {{P}iotr's {C}omputer {V}ision {M}atlab {T}oolbox ({PMT})}, 
-   howpublished = {\url{http://vision.ucsd.edu/~pdollar/toolbox/doc/index.html}} 
-}
+[2]   Piotr Doll\'ar, {P}iotr's {C}omputer {V}ision {M}atlab {T}oolbox ({PMT}), \url{http://vision.ucsd.edu/~pdollar/toolbox/doc/index.html}.
 

### 2015-12-21 Projet II/createTestingCNN.m

@@ -1,0 +1,11 @@
+function [Te] = createTestingCNN(X)
+
+    % Initialize Tr, Te
+    Te = [];
+
+    % CNN is a sparse matrix, better to use sparse representation
+    Te.X = sparse(double(X));
+    
+    % Normalizing
+    [Te.nX, u, s] = zscore(Te.X);
+end

### 2015-12-21 Projet II/neuralNetworksGenPred.m

@@ -1,0 +1,49 @@
+function [predictions] = neuralNetworksGenPred(Xtr, ytr, Xte, inputSize, innerSize, numepochs, batchsize, learningRate, binary)
+    if binary
+        outputSize = 2;
+    else
+        outputSize = 4;
+    end
+    
+    nn = nnsetup([inputSize innerSize outputSize]);
+
+    opts.numepochs  = numepochs;
+    opts.batchsize  = batchsize; % http://stats.stackexchange.com/questions/153531/what-is-batch-size-in-neural-network
+    opts.plot       = 0;
+    nn.learningRate = learningRate;
+
+    % this neural network implementation requires number of samples to be a
+    % multiple of batchsize, so we remove some for this to be true.
+    numSampToUse = opts.batchsize * floor(size(Xtr) / opts.batchsize);
+    X = Xtr(1:numSampToUse, :);
+    labels = ytr(1:numSampToUse);
+    
+    % prepare labels for NN
+    if binary
+        LL = [1 * (labels == 1), ... % first column, p(y=1)
+          1 * (labels == 2)];        
+    else
+        LL = [1 * (labels == 1), ... % first column, p(y=1)
+          1 * (labels == 2), ... % second column, p(y=2), etc
+          1 * (labels == 3), ...
+          1 * (labels == 4) ];
+    end
+    
+    tstart = tic; % because nntrain uses tic too...
+    [nn, ~] = nntrain(nn, X, LL, opts);
+    toc(tstart)
+
+    % to get the scores we need to do nnff (feed-forward)
+    tstart = tic;
+    nn.testing = 1;
+    nn = nnff(nn, Xte, zeros(size(Xte, 1), nn.size(end)));
+    nn.testing = 0;
+
+    % predict on the test set
+    nnPred = nn.a{end};
+
+    % get the most likely class
+    [~, predictions] = max(nnPred, [], 2);
+    toc(tstart)
+end
+

### 2015-12-21 Projet II/prediction_bin.m

@@ -1,0 +1,81 @@
+%% Initialization
+clear all;
+close all;
+
+addpath(genpath('Piotr'));
+addpath(genpath('DeepLearnToolbox'));
+
+load train/train.mat;
+load test/test.mat;
+
+numberOfExperiments = 30;
+proportionOfTraining = 0.8;
+K = 10;
+M = 150;
+
+train.y(find(train.y == 2)) = 1;
+train.y(find(train.y == 3)) = 1;
+train.y(find(train.y == 4)) = 2;
+
+%% Create data
+fprintf('Creating Train & Test sets\n');
+tic
+[Tr, Te] = createTrainingTestingCNN(train.X_cnn, train.y, 1.0);
+[Te] = createTestingCNN(test.X_cnn);
+toc
+
+%% Prepare data
+
+fprintf('Prepare the data for the training\n');
+tic
+[Tr, Te] = prepareDataCNN(Tr, Te, M);
+[Te] = prepareDataCNNTesting(Te, M);
+toc
+
+%% COMPUTING ESTIMATE ERROR
+
+[TTr, TTe] = splitProp(proportionOfTraining, Tr, false);
+
+idxCV = splitGetCV(K, length(TTr.y));
+
+% Setup 
+t = templateSVM('Solver', 'SMO', 'KernelFunction', 'linear', 'IterationLimit', 1e6, 'KernelScale', 1, 'BoxConstraint', 1);      
+
+% K-fold
+for k=1:1:K
+    fprintf('%d fold\n', k);
+    [TTTr, TTTe] = splitGetTrTe(TTr, idxCV, k, false);
+
+    tic
+    Mdl = fitcecoc(TTTr.Z, TTTr.y, 'Learners', t, 'Coding', 'onevsall', 'Options', statset('UseParallel', 1));
+    CMdl = compact(discardSupportVectors(Mdl));
+    yhat = predict(CMdl, TTTe.Z);
+    toc
+
+    err(k) = balancedErrorRate(TTTe.y, yhat);
+    fprintf('%f ', err(k));
+end
+
+% 0.089 +- 0.006
+
+u = mean(err); 
+o = std(err);
+
+%% TRAINING FOR PREDICTION
+
+% setup
+t = templateSVM('Solver', 'SMO', 'KernelFunction', 'linear', 'IterationLimit', 1e6, 'KernelScale', 1, 'BoxConstraint', 1);      
+
+% Train
+tic
[…]

### 2015-12-21 Projet II/prediction_mul.m

@@ -1,0 +1,75 @@
+%% Initialization
+clear all;
+close all;
+
+addpath(genpath('Piotr'));
+addpath(genpath('DeepLearnToolbox'));
+
+load train/train.mat;
+load test/test.mat;
+
+numberOfExperiments = 30;
+proportionOfTraining = 0.8;
+K = 10;
+M = 150;
+
+%% Create data
+fprintf('Creating Train & Test sets\n');
+tic
+[Tr, Te] = createTrainingTestingCNN(train.X_cnn, train.y, 1.0);
+[Te] = createTestingCNN(test.X_cnn);
+toc
+
+%% Prepare data
+
+fprintf('Prepare the data for the training\n');
+tic
+[Tr, Te] = prepareDataCNN(Tr, Te, M);
+[Te] = prepareDataCNNTesting(Te, M);
+toc
+
+%% COMPUTING ESTIMATE ERROR
+
+[TTr, TTe] = splitProp(proportionOfTraining, Tr, false);
+
+idxCV = splitGetCV(K, length(TTr.y));
+
+% Setup NN.
+inputSize  = M;
+innerSize  = 700;
+numepochs = 20;
+batchsize = 100;
+learningRate = 2;
+binaryClassification = false;
+
+% K-fold
+for k=1:1:K
+    fprintf('%d fold\n', k);
+    [TTTr, TTTe] = splitGetTrTe(TTr, idxCV, k, false);
+
+    [errnZ, nnPrednZ] = neuralNetworks(TTTr.nZ, TTTr.y, TTTe.nZ, TTTe.y, inputSize, innerSize, numepochs, batchsize, learningRate, binaryClassification);
+
+    err(k) = errnZ;
+end
+
+% 0.085 +- 0.010
+
+u = mean(err); 
+o = std(err);
+
+%% TRAINING FOR PREDICTION
+
+% Setup NN.
+inputSize  = M;
+innerSize  = 700;
+numepochs = 20;
+batchsize = 100;
+learningRate = 2;
+binaryClassification = false;
+
+% Train & test
+
+[Ytest] = neuralNetworksGenPred(Tr.nZ, Tr.y, Te.nZ, inputSize, innerSize, numepochs, batchsize, learningRate, binaryClassification);
+
+% Write the prediction
+save('pred_multiclass', 'Ytest');

### 2015-12-21 Projet II/prepareDataCNNTesting.m

@@ -1,0 +1,34 @@
+function[Te] = prepareDataCNNTesting(Te_, M)
+    [Te] = PCA(Te_, M);
+end
+
+function [Te] = PCA(Te_, M)
+    Te = Te_;
+    
+    X = Te.nX;
+    N = size(X, 1);
+    D = size(X, 2);
+
+    % Use method in Bishop 12.1.4
+    Xu = mean(X, 1);
+    
+    % Transform to non-space matrix
+    X2 = X - ones(N, 1) * Xu;
+    
+    % Compute S term
+    S = X2 * X2' / N;
+    
+    % Compute and sort eigen values
+    [V, l] = eig(S, 'vector');
+    [l, I] = sort(l, 1, 'descend');
+    V = V(:, I);
+    l = abs(l); % To avoid neg values, only for the "0" which can be negative
+    
+    % Compute U term
+    norm = sqrt(N*l)';
+    norm = repmat(norm, D, 1);
+    U = X2' * V ./ norm;
+    
+    Te.nZ = Te.nX * U(:, 1:M);
+    Te.Z = Te.X * U(:, 1:M);
+end

### 2015-12-21 Projet II/testMyPredictions.m

@@ -1,0 +1,65 @@
+% --- OBJECT DETECTION PREDICTION CHECK ----
+%
+% Store your predictions in two mat files named 'pred_binary.mat'
+% and 'pred_multiclass.mat'.
+% This mat file should contain a vector 'Ytest'
+% which contains the prediction _score_ for each test sample.
+% The size of Ytest must be 11453x1.
+% One way to create these matrices is shown below:
+%
+%   -- assign your predicted scores to Ytest first, then: --
+%   save('pred_binary', 'Ytest');
+%
+% Once you have the file, running this file will check if the sizes are correct or not.
+
+
+%% == Test all vs other
+clear all;
+fprintf('==> Checking all vs other\n\n')
+
+load pred_binary.mat;
+
+% exists?
+if ~exist('Ytest', 'var')
+  error('No Ytest vector found');
+end
+
+% check size
+[nU,nA] = size(Ytest);
+if (nU ~= 11453) && (nA ~= 1)
+  error('size of Ytest is incorrect');
+end
+
+% make sure it contains scores and not labels
+nUniqueVals = length(unique(Ytest(:)));
+uniqueVals = sort(unique(Ytest(:)), 'ascend');
+if nUniqueVals ~= 2 || any( uniqueVals(:) ~= [0,1]' )
+    error('Ytest should be either 0 or 1');
+end
+
+
+fprintf('==> Checking multiclass\n\n')
+clearvars;
+
+%% == Test all vs other
+load pred_multiclass.mat;
+
+% exists?
+if ~exist('Ytest', 'var')
+  error('No Ytest vector found');
+end
+
+% check size
+[nU,nA] = size(Ytest);
+if (nU ~= 11453) && (nA ~= 1)
+  error('size of Ytest is incorrect');
+end
+
+% make sure it contains scores and not labels
+nUniqueVals = length(unique(Ytest(:)));
+uniqueVals = sort(unique(Ytest(:)), 'ascend');
+if nUniqueVals ~= 4 || any( uniqueVals(:) ~= [1,2,3,4]' )
+    error('Ytest should be either 1,2,3 or 4');
+end
+
+fprintf('\nSuccessful, you can submit!\n\n');

### 2015-12-21 Projet II/report/paper.tex
advance report and clean code

@@
 For both methods, we have observed that the given CNN features were more accurate than the HOG one.
-In binary classification, all methods were approximately equivalent except the Adaboosting and the SVM one, but we choose the SVM method as our final model.
-For the multi-class classification, neural networks and SVM were the best methods but neural networks has less variance in this case and is considered as our final model.
+In binary classification, all methods were approximately equivalent except the Adaboosting and the SVM one, but we choose the SVM method as our final model because it was more stable.
+For the multi-class classification, neural networks and SVM were the best methods but neural networks has less variance in this case and was considered as our final model.
 The best way of improving our models would be to find a way to combine HOG features with the CNN one.
@@
 
-Each input sample is a vector $\mathbf{x}_n$ with dimension $D = 42273$ ($5408$ for the HOG and $36865$ for the CNN) and is the concatenation of $\mathbf{x}_n^{hog}$ with $\mathbf{x}_n^{cnn}$
-The output ($\mathbf{y}$) represents the classification of the images.
+Each input sample is a vector $\mathbf{x}_n$ with dimension $D = 42273$ ($5408$ for the HOG and $36865$ for the CNN) and is the concatenation of $\mathbf{x}_n^{hog}$ with $\mathbf{x}_n^{cnn}$.
+The output ($\mathbf{y}$) represents the classification label of the images.
 For the binary classification, the label $1$  represents a car, a horse or a plane and the label $2$ anything else.
-For the multi-class classification, the label $1$ represents a car, $2$2 a horse, $3$ a plane and $4$ anything else.
+For the multi-class classification, the label $1$ represents a car, $2$ a horse, $3$ a plane and $4$ anything else.
 
@@
 We first have plot the labels distribution for the binary and multi-class cases.
-For the binary case, we have $60.3\%$ for the label $1$ and $39.8\%$ for the second one.
+For the binary case, we have $60.3\%$ of the data for the label $1$ and $39.8\%$ for the second one.
 We have observed that there is a small imbalance between the classes.
@@
 
-We also wanted to analyze the distribution of our features, and have observed that for the features HOG, each of them has a has similar distribution.
+We also wanted to analyze the distribution of our features, and have observed that for the features HOG, each of them has a similar distribution.
 The values are essentially in the range $[0, 0.2]$.
@@
 
-This section is a description of the techniques we have used in the project. It describes mainly how it works, why we thought it could be useful in this project.
+This section is a description of the techniques we have used in the project. It describes mainly how it works, why we thought it could be useful in this project and the impact of their parameters.
 
@@
 \center
-\subfigure[Plot of the distortion $J$ to $M$ where $M = D$. The distortion decreases quickly.]{\includegraphics[width=2.7in]{figures/distortionHOG.jpg} \label{fig:distHOG}}
+\subfigure[Plot of the distortion $J$ to $M$ where $M = D$ for the HOG features. The distortion decreases quickly.]{\includegraphics[width=2.7in]{figures/distortionHOG.jpg} \label{fig:distHOG}}
 \hfill
-\subfigure[Plot of the distortion $J$ to $M$ where $M = N$ because there are too many features and so, the distortion doesn't go to $0$ and decreases slowly.]{\includegraphics[width=2.7in]{figures/distortionCNN.jpg}\label{fig:distCNN}}
+\subfigure[Plot of the distortion $J$ to $M$ for the CNN features, where $M = N$ because there are too many features and so, the distortion doesn't go to $0$ and decreases slowly.]{\includegraphics[width=2.7in]{figures/distortionCNN.jpg}\label{fig:distCNN}}
 \hfill
@@
 However, we couldn't compute PCA using this method, because computing the full eigenvector decomposition a matrix $D \times D$ runs in  $O(D^3)$ and D is quite large.
-We have used an alternative proposed in [REF] which allows us to compute the same eigenvectors in $O(N^3)$, which is roughly at least $230$ times faster.
+We have used an alternative proposed in [3] which allows us to compute the same eigenvectors in $O(N^3)$, which is roughly at least $230$ times faster.
 The tricks is to express $S = \frac{1}{N}XX^T$ and finally using $\frac{1}{N}XX^Tv_i = \lambda_i v_i$, where $v_i=Xu_i$, we obtain an eigenvector equation for the $N \times N$ matrix $S$.
@@
 
-The Figure \ref{fig:distHOG} shows the plot of $J$ for the HOG features and the CNN features.
+The Figure \ref{fig:distHOG} shows the plot of $J$ for the HOG features.
 Figure \ref{fig:distCNN} shows the distortion of the CNN features, we have tried several values for $M$ and have conclude that $M_{CNN} = 150$ was a good value allowing us to compute faster results without loosing too much accuracy.
@@
 We tried some bigger values but we didn't get major improvements in terms of error.
-For the HOG features, the distortion measure seems more nice, we did the same experiments and finally set $M_{HOG}$ = $500$.
+For the HOG features, the distortion measure seems more nice, we did the same experiments and finally set $M_{HOG}$ = $1000$.
 
@@
 This is the multi-layer perceptron (MLP) we have seen in class.
-The neural networks used in this project comes from the given DeepLearning matlab toolbox [REF].
+The neural networks used in this project comes from the given DeepLearning matlab toolbox [1].
 It uses batch sizes in order to use less memory (take the first $n$ data samples and works on them, then the next $n$ etc).
@@
 For the number of hidden layer, we have let $1$ because we didn't see improvement by adding some more.
-However, the number of neurons in the hidden layer has been tested with k-fold cross validation and we have found $1000$ for the binary classification and $700$ for the multi-class case.
+However, if this number is to high, we might overfit and so have to be careful.
+Having this value to $1$ might not lead to overfit.
+The number of neurons in the hidden layer has been tested with k-fold cross validation and we have found $1000$ for the binary classification and $700$ for the multi-class case.
 We also have to set up the learning rate : the optimal value we have found is $3$ for the binary case and $2$ for the multi-class one.
+Larger is the value, faster could be the convergence but it may diverge.
 By default, the activation function used is the \textit{tanh} one.
@@
 Having a high maximum depth might lead to overfit because we specialize to much the features split, but if it is too small, it might underfit because the features are not well split in order to identify classes.
-For the parameter $m$, it is better to have a higher number in order to have a large number of trees which can give a better accuracy of the prediction.
+For the parameter $m$, it is better to have a higher number in order to have a large number of trees which can give a better accuracy of the prediction and reduce the variance.
 
@@
 Like bagging, each trees have trained on a subset of the dataset.
-However, there is another aspect in random forest that bagging doesn't have: for each split, instead of searching among the overall features, we search only among a random subset of features.
+However, there is another aspect in random forest that bagging doesn't have: for each split, instead of searching a split among the overall features, we search only among a random subset of features.
 The size of this subset is a parameter $m$ as well as the maximum depth tree and the number of trees, like in bagging.
-Like bagging, random forests are useful because they help reducing the variance and the overfiting.
+Like bagging, random forests are useful because they help reducing the variance and might avoid overfitting.
 But the more interesting ideas is to select only a subset of the features in the split process.
@@
 
-The given Piotr matlab toolboox [REF] contains a method \textit{forestTrain} which takes the defined parameters and \textit{forestApply} for the prediction.
+The given Piotr matlab toolboox [2] contains a method \textit{forestTrain} which takes the defined parameters and \textit{forestApply} for the prediction.
 
@@
 
-Adaboosting is another way to average trees.
+Adaboosting [4] is another way to average trees.
 If we observe more deeply the bagging approach, some trees might have a very high variance and some a very slow.
@@
 The parameters are the same as in the bagging and behave in the same manner.
-The Piotr matlab toolboox [REF] contains a method \textit{adaBoostTrain} and \textit{adaBoostApply} for the prediction.
+The Piotr matlab toolboox [2] contains a method \textit{adaBoostTrain} and \textit{adaBoostApply} for the prediction.
 We have only used it for the binary classification, because the implementation is not adapted for more classes.
+However, we should have used the same technique as SVM : OVA.
 The parameters we obtain with k fold cross validation are $1024$ for the number of trees and $4$ for the maximum depth.
@@
 This error compute the average percentage of misclassification per class.
-It allows us to avoid the case of the $90\%/10\%$ and having a good error.
+It allows us to avoid the case of the $90\%/10\%$, predict only the class of the $90\%$ and have a good error.
 We have also used k-fold cross validation to tune the parameters up of our different models.
@@
 The reason is because with only the HOG features, we couldn't get an error lower than $20\%$ for binary and multi-class classification.
+We have tried to take the whole matrix, only a subset of features with $M = 1000$, different methods but nothing worked.
 With the CNN features, we didn't have this problem.
-We have tried to take the whole matrix, only a subset of features with $M = 500$, different methods but nothing worked.
-Contrary to the CNN features, we think that the reason is because those features are too noisy : they came directly form the images.
-Some of them are rotated, inversed and so, the gradient is also transformed, which might be confusing for the learning.
+We think that the reason is because those features are too noisy : they came directly form the images.
+Some of them are rotated, inverted and so, the gradient is also transformed, which might be confusing for the learning.
 The CNN features are already computed from a convolutional neural networks which gives the features it has seen on the image with a specific weight.
@@
 
-[1]    IMM2012-06284, R. B. Palm, Prediction as a candidate for learning deep hierarchical models of data, 2012.
+[1]    IMM2012-06284, R. B. Palm, "Prediction as a candidate for learning deep hierarchical models of data", 2012.
 
@@
 
+[3]    Pattern Recognition and Machine Learning, Christopher Bishop, p 569-570, 2007.
+
+[4]    R. Appel, T. Fuchs, P. Dollár, P. Perona; "Quickly Boosting Decision Trees – Pruning Underachieving Features Early," ICML 2013.
 \end{document}

### 2015-12-21 Projet II/BT_CNN_PCA.m

@@ -1,3 +1,6 @@
+% Written by Diego Antognini & Jason Racine, EPFL 2015
+% all rights reserved
+
 %% Initialization
 clear all;
 close all;

### 2015-12-21 Projet II/DT_HOG_PCA.m

@@ -1,3 +1,6 @@
+% Written by Diego Antognini & Jason Racine, EPFL 2015
+% all rights reserved
+
 %% Initialization
 clear all;
 close all;
@@ -8,7 +11,7 @@ addpath(genpath('DeepLearnToolbox'));
 load train/train.mat;
 
 ratio = 0.8;
-M = 3000;
+M = 1000;
 
 %% Create data
 fprintf('Creating Train & Test sets\n');

### 2015-12-21 Projet II/NN_HOG_PCA.m

@@ -1,3 +1,6 @@
+% Written by Diego Antognini & Jason Racine, EPFL 2015
+% all rights reserved
+
 %% Initialization
 clear all;
 close all;
@@ -8,7 +11,7 @@ addpath(genpath('DeepLearnToolbox'));
 load train/train.mat;
 
 ratio = 0.8;
-M = 500;
+M = 1000;
 
 %% Create data
 fprintf('Creating Train & Test sets\n');

### 2015-12-21 Projet II/SVM_CNN_PCA.m

@@ -1,3 +1,6 @@
+% Written by Diego Antognini & Jason Racine, EPFL 2015
+% all rights reserved
+
 %% Initialization
 clear all;
 close all;
@@ -8,7 +11,7 @@ addpath(genpath('DeepLearnToolbox'));
 load train/train.mat;
 
 ratio = 0.8;
-M = 200;
+M = 150;
 
 %% Create data
 fprintf('Creating Train & Test sets\n');

### 2015-12-21 Projet II/imageClassificationExample.m

@@ -1,115 +1,0 @@
-clearvars;
-
-% -- GETTING STARTED WITH THE IMAGE CLASSIFICATION DATASET -- %
-% IMPORTANT:
-%    Make sure you downloaded the file train.tar.gz provided to you
-%    and uncompressed it in the same folder as this file resides.
-
-% Load features and labels of training data
-load train/train.mat;
-
-%% --browse through the images and look at labels
-for i=1:10
-    clf();
-
-    % load img
-    img = imread( sprintf('train/imgs/train%05d.jpg', i) );
-
-    % show img
-    imshow(img);
-
-    title(sprintf('Label %d', train.y(i)));
-
-    %pause;  % wait for key,??
-end
-
-%% -- Example: split half and half into train/test, use HOG features
-fprintf('Splitting into train/test..\n');
-
-Tr = [];
-Te = [];
-
-% NOTE: you should do this randomly! and k-fold!
-Tr.idxs = 1:2:size(train.X_hog,1);
-Tr.X = train.X_hog(Tr.idxs,:);
-Tr.y = train.y(Tr.idxs);
-
-Te.idxs = 2:2:size(train.X_hog,1);
-Te.X = train.X_hog(Te.idxs,:);
-Te.y = train.y(Te.idxs);
-
-%%
-fprintf('Training simple neural network..\n');
-
-addpath(genpath('DeepLearnToolbox'));
-
-
-rng(8339);  % fix seed, this    NN may be very sensitive to initialization
-
-% setup NN. The first layer needs to have number of features neurons,
-%  and the last layer the number of classes (here four).
-nn = nnsetup([size(Tr.X,2) 10 4]);
-opts.numepochs =  20;   %  Number of full sweeps through data
-opts.batchsize = 100;  %  Take a mean gradient step over this many samples
-
-% if == 1 => plots trainin error as the NN is trained
-opts.plot               = 1;
-
-nn.learningRate = 0.5;
-
-% this neural network implementation requires number of samples to be a
-% multiple of batchsize, so we remove some for this to be true.
-numSampToUse = opts.batchsize * floor( size(Tr.X) / opts.batchsize);
-Tr.X = Tr.X(1:numSampToUse,:);
-Tr.y = Tr.y(1:numSampToUse);
-
-% normalize data
-[Tr.normX, mu, sigma] = zscore(Tr.X); % train, get mu and std
-
-% prepare labels for NN
-LL = [1*(Tr.y == 1), ...
[…]

### 2015-12-21 Projet II/main.m

@@ -1,16 +1,0 @@
-addpath(genpath('Piotr'));
-
-load train/train.mat;
-
-i = randi(length(train.y));
-figure
-% load img
-img = imread(sprintf('train/imgs/train%05d.jpg', i));
-
-subplot(121);
-imshow(img); % image itself
-
-feature = hog( single(img)/255, 17, 8);
-im( hogDraw(feature) ); colormap gray;
-axis off; colorbar off;
-

### 2015-12-21 Projet II/neuralNetworksGenPred.m

@@ -1,3 +1,6 @@
+% Written by Diego Antognini & Jason Racine, EPFL 2015
+% all rights reserved
+
 function [predictions] = neuralNetworksGenPred(Xtr, ytr, Xte, inputSize, innerSize, numepochs, batchsize, learningRate, binary)
     if binary
         outputSize = 2;

### 2015-12-21 Projet II/prediction_mul.m

@@ -1,3 +1,6 @@
+% Written by Diego Antognini & Jason Racine, EPFL 2015
+% all rights reserved
+
 %% Initialization
 clear all;
 close all;
@@ -8,7 +11,6 @@ addpath(genpath('DeepLearnToolbox'));
 load train/train.mat;
 load test/test.mat;
 
-numberOfExperiments = 30;
 proportionOfTraining = 0.8;
 K = 10;
 M = 150;

### 2015-12-21 Projet II/prepareDataCNN.m

@@ -1,3 +1,6 @@
+% Written by Diego Antognini & Jason Racine, EPFL 2015
+% all rights reserved
+
 function[Tr, Te] = prepareDataCNN(Tr_, Te_, M)
     [Tr, Te] = PCA(Tr_, Te_, M);
 end

### 2015-12-21 Projet II/prepareDataCNNTesting.m

@@ -1,3 +1,6 @@
+% Written by Diego Antognini & Jason Racine, EPFL 2015
+% all rights reserved
+
 function[Te] = prepareDataCNNTesting(Te_, M)
     [Te] = PCA(Te_, M);
 end

### 2015-12-21 Projet II/prepareDataHOG.m

@@ -1,3 +1,6 @@
+% Written by Diego Antognini & Jason Racine, EPFL 2015
+% all rights reserved
+
 function[Tr, Te] = prepareDataHOG(Tr_, Te_, M)
     [Tr, Te] = PCA(Tr_, Te_, M);
 end

### 2015-12-21 Projet II/report/paper.tex
Review of the report

@@
 The other one has a bigger variance but is more robust to outliers.
-Let's remind than bagging averages trees and some trees might have big and small variance, which reduce variance in average.
+Let's remind than bagging averages trees and some of them might have big and small variance, which reduce variance in average.
 In general, it reduces the variance and increase slightly the bias.
@@
 We can also see that there are some outliers, but they are still better than the other seen methods.
-We think that the main reason is the variance and seems to be higher than the other models.
+We think that the main reason is taht the variance seems to be higher than the other models.
 In general, boosting should be better than bagger.
@@
 In the other side, if $C$ is small, the bias should be lower but we have fewer support vectors which leads to a higher variance.
-However, let's remind that we use the same as in the multi-class classifier, using a linear kernel due to the implementation of matlab.
+However, let's remind that we use the same configuration as in the multi-class classifier, using a linear kernel due to the implementation of matlab.
 If we could have changed the kernel into a non-linear one, the decision boundary would be different and might surely reduce the error.
@@
 
-Our final model will be the SVM one, because we considered it has one of the smallest error, but it doesn't have any outliers and is more stable than the Adaboosting.
+Our final model will be the SVM one, because we considered it has one of the smallest error, it doesn't have any outliers and is more stable than the Adaboosting.
 
 We wanted to test some others models but we didn't get time to do it.
-First, we wanted to see if using a multi-class classifier which have a smallest error and then transform the prediction into a binary case would increase the accuracy.
+First, we wanted to see if using a multi-class classifier which have a smallest error and then transform the prediction into a binary case would decrease the accuracy.
 We think that the error might be improved but in general, it should stay more or less the same.
-Another promising method we wanted to try is the \textit{fernst} one.
-It should outperforms other decision-based trees because they compute posteriors multiplicatively [REF].
+Another promising method we wanted to try is the \textit{fernst} one [5].
+It should outperforms other decision-based trees because they compute posteriors multiplicatively.
 We obtain good results with the Adaboosting method and think that fernst should produce better results.
@@
 \label{fig:mul_models}
+
+\vspace{-1em}
 \end{figure}
@@
 Another observation we might see is that in general, the variance is lower in the multi-class models.
-We think that the reason is that we classify explicit objects (car, horse, plane and others) without mixing them, which might be less noisy.
+We think that the reason is that we classify explicit objects (car, horse, plane and others) without mixing their class, which might be less noisy.
 
@@
 The second model uses neural networks with normalized data as explained in the binary section.
-We can see that even if we have more label, neural networks are still efficient and has a low bias and a little bit of variance, because there is one outlier.
+We can see that even if we have more labels, neural networks are still efficient and has a low bias and a bit of variance, because there is one outlier.
 Obviously, neural network are a lot more efficient than our baseline.
@@
 The next model ($5$) is the SVM one using the OVA method.
-As we said earlier, it is several binary classifiers combined together.
+As we said earlier, it combines several binary classifiers.
 We obtain similar results than in the binary classification because the nature of the problem is very similar : the only difference is the number of labels.
@@
 
-The last model uses random forests. As before, we obtain worse results than the other methods. It has greater bias than the other.
+The last model uses random forests.
+As before, we obtain worse results than the other methods.
 We also think that the method has overfit when using k fold cross validation in order to find optimal parameters.
@@
 We also think that the bias could be high and so, increasing the parameter $m$ could be worthwhile because first, we may need more features to classify a data sample in one of the four classes.
-Secondly, because the bias is surely higher than other method as well as the variance.
+Secondly, because the bias is surely higher than other methods as well as the variance.
 More training samples should help to decrease the variance and to tune the parameters up.
@@
 
-Beside trying \textit{fernst} method as proposed in the binary case, we wanted to try to use several multi-class classifiers and then combine their results to see which data samples were more difficult and find a way to class them with more accuracy.
+Beside trying \textit{fernst} method [5] as proposed in the binary case, we wanted to try to use several multi-class classifiers and then combine their results to see which data samples were more difficult and find a way to class them with more accuracy.
 One of the method we wanted to use was simply voting system in a similar manner as SVM OVA.
@@
 The errors obtained with binary classification and multi-class classification are closed and prove that the nature of the problem is similar.
-We have tried different methods and have remarked that SVM model was better than the other in general.
+We have tried different methods and have remarked that SVM model, Adaboosting and neural networks were better than the other sin general.
 In both case, it produces good results and has a small variance.
-We really think that there is a way of improving it by changing the kernel function because data are surely not separate linearly.
+We really think that the SVM model could be improved by changing the kernel function because data are surely not separate linearly.
 At the end, we think that the best way of improving our binary and multi-class classifier would be to find a way to combine the HOG features with the CNN ones.
+
+We have estimated the mean and the stanrd deviation of the prediction, using also THE BER.
+For the binary classification, we obtain a result of $0.089 \pm 0.006$ and for the multi-class classification $0.085 \pm 0.010$.
 
@@
 [4]    R. Appel, T. Fuchs, P. Dollár, P. Perona; "Quickly Boosting Decision Trees – Pruning Underachieving Features Early," ICML 2013.
+
+[5]   Mustafa Ozuysal, Pascal Fua and Vincent Lepetit, "Fast Keypoint Recognition in Ten Lines of Code" by , CVPR07.
 \end{document}

### 2016-07-24 Projet II/report/paper.tex
Fix typos

@@
 Histogram of oriented gradients is a feature used in computer vision in order to detect objects.
-To compute it, we first need to normalize the image, compute the gradients (of each pixel) for the different color channel and take those with the largest norm.
+To compute it, we first need to normalize the image, compute the gradients (of each pixel) for the different color channels and take those with the largest norm.
 Then we decompose the image in bins of size $w \times h$.
@@
 
-We first have plot the labels distribution for the binary and multi-class cases.
+We first have plot the label distribution for the binary and multi-class cases.
 For the binary case, we have $60.3\%$ of the data for the label $1$ and $39.8\%$ for the second one.
@@
 
-For the features CNN, first we can see that it is a sparse matrix and most of the values for each data is $0$ (around $95.5\%$).
+For the features CNN, first we can observe that most of the values for each data is $0$ (around $95.5\%$) and so, is a sparse matrix.
 This means that for each data, a small subset is able to describe the picture.
@@
 \hfill
-\caption{}
+\caption{Distortion of the features HOG and CNN.}
 \end{figure}
@@
 
-As we have seen in the section "Data visualization and analysis", we are faced to the $D > N$ problem mainly with the CNN features and also the HOG features depending on how we se the $k$ for the k-fold cross validation.
+As we have seen in the section "Data visualization and analysis", we are faced to the $D > N$ problem mainly with the CNN features and also the HOG features depending on how we set the $k$ for the k-fold cross validation.
 One way to solve this problem is to use PCA.
@@
 Finally, the approximation of a data sample is $\tilde{x_n} = \sum_{i=1}^M x_{ni} u_i$.
-However, we couldn't compute PCA using this method, because computing the full eigenvector decomposition a matrix $D \times D$ runs in  $O(D^3)$ and D is quite large.
+However, we couldn't compute PCA using this method, because computing the full eigenvector decomposition of a matrix $D \times D$ runs in  $O(D^3)$ and D is quite large.
 We have used an alternative proposed in [3] which allows us to compute the same eigenvectors in $O(N^3)$, which is roughly at least $230$ times faster.
@@
 
-The last parameter to set, is the parameter $M$ which defines the $M$ largest eigenvectors to take into consideration for our feature selection.We introduce a distortion measure $J = \frac{1}{N}\sum_{n=1}^{N}||x_n-\tilde{x_n}||^2$ which represents the mean squared distance between the original data sample and our approximation.
+The last parameter to set is $M$ which defines the $M$ largest eigenvectors to take into consideration for our feature selection.We introduce a distortion measure $J = \frac{1}{N}\sum_{n=1}^{N}||x_n-\tilde{x_n}||^2$ which represents the mean squared distance between the original data sample and our approximation.
 Without going into the steps, the solution to the minimization of $J$ is $J = \sum_{i=M+1}^D \lambda_i$.
@@
 The number of epochs is simply the number of time where all the training data do a forward and a backward pass.
-We have put this value to $20$, in order to have a little more accuracy without loosing a lot of time, it was a good compromise because larger epochs don't mean especially better accuracy.
+We have put this value to $20$ in order to have a little more accuracy without loosing a lot of time, it was a good compromise because larger epochs don't mean especially better accuracy.
 The number of input is $M$ and output is either $2$ or $4$, depending of the type of classification (binary or multi-class).
-For the number of hidden layer, we have let $1$ because we didn't see improvement by adding some more.
+For the number of hidden layer, we have let $1$ because we didn't see any improvement by adding some more.
 However, if this number is to high, we might overfit and so have to be careful.
@@
 
-As we have seen in the lecture, decision trees are good because the training and the prediction is quite fast, even with many features or large dataset but have a high variance and might overfit.
-The idea of decision trees, is to have a several questions where to answer yes/no and to move in the tree.
+As we have seen in the lecture, decision trees are good because the training and the prediction computation is quite fast, even with many features or large dataset, but have a high variance and might overfit.
+The idea of decision trees is to have several questions where we answer yes/no and then to move in the tree.
 At the end, the leaves give the corresponding class.
@@
 Finally, to predict the class it takes the average over all the trees.
-In order to find those parameters $m$ and the maximum depth of the tree, we have also used k-fold cross validation and the best values are $m = 200$ and maximum depth of $512$ for the binary classification and for the multi-class case $m = 500$ and maximum depth also of $512$.
-Having a high maximum depth might lead to overfit because we specialize to much the features split, but if it is too small, it might underfit because the features are not well split in order to identify classes.
-For the parameter $m$, it is better to have a higher number in order to have a large number of trees which can give a better accuracy of the prediction and reduce the variance.
+In order to find those parameters $m$ and the maximum depth of the tree, we have also used k-fold cross validation.
+The best values are $m = 200$ and maximum depth of $512$ for the binary classification and for the multi-class case $m = 500$ and maximum depth also of $512$.
+Having a high maximum depth might lead to overfit because we specialize too much the features split.
+But if it is too small it might underfit because the features are not well split in order to identify classes.
+For the parameter $m$ it is better to have a higher number in order to have more trees, which can give a better accuracy of the prediction and reduce the variance.
 
@@
 As in bagging trees, the number of trees and the maximum depth tree behaves in the same manner.
-The last parameter $m$ has an important impact because it is too small, the subset of features might not describe correctly the data, and if it is too high, this subset might be to high and might not represent well the data.
+The last parameter $m$ has an important impact because if it is too small, the subset of features might not describe correctly the data, and if it is too high, this subset might be too large and might not represent well the data.
 We have to find a good compromise in order to have a certain number of features to describe the data.
-Still with k fold cross validation, we have found for the multi-class case a maximum depth of $1024$,  $64$ trees and a $m = 20$.
+Still with k fold cross validation, we have found for the multi-class case a maximum depth of $1024$,  $64$ trees and $m = 20$.
 For $m$, an heuristic could be the square root of the number of features, but it was less efficient than $20$ in our case.
@@
 Adaboosting [4] is another way to average trees.
-If we observe more deeply the bagging approach, some trees might have a very high variance and some a very slow.
+If we observe more deeply the bagging approach, some trees might have a very high variance and others a very low one.
 Averaging all those trees decrease the variance and have a small bias.
@@
 
-The Figure \ref{fig:bin_models} represents the BER error for each models.
+The Figure \ref{fig:bin_models} represents the BER error for each model.
 The first model used simply returns the label $1$ to all inputs, which will allow us to compare improvements with other models.
@@
 The other one has a bigger variance but is more robust to outliers.
-Let's remind than bagging averages trees and some of them might have big and small variance, which reduce variance in average.
+Let's remind that bagging averages trees and some of them might have big and small variance, which reduce variance in average.
 In general, it reduces the variance and increase slightly the bias.
@@
 As in bagger, it reduces the variance because it takes the average and the bias should be lower, which is the contrary compared to the bagger.
-We can also see that there are some outliers, but they are still better than the other seen methods.
-We think that the main reason is taht the variance seems to be higher than the other models.
+We can also see that there are some outliers, but they are still better than the other methods seen.
+We think that the main reason is that the variance seems to be higher than the other models.
 In general, boosting should be better than bagger.
@@
 Random forests is used for the sixth model.
-Surprisingly, we obtain worse results than the other methods.
+Surprisingly, we obtain worse results compared to the other methods.
 The k fold cross validation gave us better results when we were looking for the parameters.
@@
 The errors obtained with binary classification and multi-class classification are closed and prove that the nature of the problem is similar.
-We have tried different methods and have remarked that SVM model, Adaboosting and neural networks were better than the other sin general.
-In both case, it produces good results and has a small variance.
-We really think that the SVM model could be improved by changing the kernel function because data are surely not separate linearly.
+We have tried different methods and have remarked that SVM model, Adaboosting and neural networks were better than the others in general.
+In both cases, it produces good results and has a small variance.
+We really think that the SVM model could be improved by changing the kernel function because data are surely not linearly separable.
 At the end, we think that the best way of improving our binary and multi-class classifier would be to find a way to combine the HOG features with the CNN ones.
 
-We have estimated the mean and the stanrd deviation of the prediction, using also THE BER.
+We have estimated the mean and the standard deviation of the prediction, using also the BER.
 For the binary classification, we obtain a result of $0.089 \pm 0.006$ and for the multi-class classification $0.085 \pm 0.010$.
